### 2024-04-12 README.Rmd
inital commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# msp-ms

<!-- badges: start -->
<!-- badges: end -->

The goal of msp-ms is provide a concise code-base for the normalization and data processing required for the msp-ms method.

Additionally, we provide a graphical user interface powered by shiny apps that allows for a user to utilize the method without requiring any R coding knowledge.

So how does the msps data normalization process work?

1. Takes two input files from PEAKS and combines them
3. Normalizes values and then does a reverse log2 transformation
4. Looks for outliers across replicates. Removes them. 
5. Imputes data for missing values (likely to be very low intesity).
6. Figures out the locations of the detected clevages within the library of peptide sequences used. Is it cleaved at the N or C terminus, or both?
7. Calculates the fold change and the p/q value across experimental conditions using T-tests.

## Combining Peaks file outputs

The files coming from peaks should have..... **Describe how to do this here**.

Importantly, this. data contains... what????

```{r}
### Loading the files ###
lfq_filename = "legacy/protein-peptides-lfq.csv"
#file "protein-peptides.csv" exported from PEAKS identification
id_filename = "legacy/protein-peptides-id.csv"

# Prepare the data for normalyzer analysis
prepared_data = prepare_for_normalyzer(lfq_filename,id_filename)
```

## Normalizing values.

msp-ms uses the .... package to do normalization under the hood.

First we need to extract a design matrix.

Then we can normalyze the data. 
```{r}
# Extracting design matrix
design_matrix = extract_design_matrix(prepared_data)

# normalyze the data

source("R/normalyze.R")

normalyzed_data = normalyze(prepared_data,design_matrix)
```

## Handling Outliers

Here we use a dixion test from the outliers package to detect outliers from each of our replicates.

```{r}
outliers = handle_outliers(normalyzed_data)
```

## Imputation of data

We have a lot of missing, or 0 values. For these, we need to impute them so we can do downstream statistics
Data is imputated by ....

```{r}
imputed = impute(outliers)
```

## Joining with Library

Next we need to join everything with the sequences of the peptide library
```{r}
joined_with_library = join_with_library(imputed)
```

## Calcuating clevages
Next, we need to determine where the peptide sequences are cleaved.

We check both the N and C terminus.

Sequences are presented as the 4 amino acids on both sides of a cleavage. X indicates that there was nothing on that side in the library because it was cleaved close to the edge.

```{r}
final_data = add_cleavages(joined_with_library)
```

## Conducting stats

Right now this is left to the user to do on their own. Will update after talkng to Bri.

### 2024-04-12 new_workflow.Rmd
inital commit

---
title: "updated_workflow"
output: html_document
date: "2024-03-29"
---

```{r}
#library(mspms)
source("R/prepare_for_normalyzer.R")
source("R/extract_design_matrix.R")

### Loading the files ###
lfq_filename = "legacy/protein-peptides-lfq.csv"
#file "protein-peptides.csv" exported from PEAKS identification
id_filename = "legacy/protein-peptides-id.csv"

# Prepare the data for normalyzer analysis
prepared_data = prepare_for_normalyzer(lfq_filename,id_filename)

# Extracting design matrix
design_matrix = extract_design_matrix(prepared_data)

# normalyze the data

source("R/normalyze.R")

normalyzed_data = normalyze(prepared_data,design_matrix)

source("R/handle_outliers.R")
outliers = handle_outliers(normalyzed_data)

#outliers get converted to NAs. We dont want to impute these. We want to impute the values that are not NAs
source("R/impute.R")
imputed = impute(outliers)

# Writing sequences to file
peptide_library2 = peptide_library %>% 
  mutate(library_reference_id = gsub(">","",library_reference_id))

saveRDS(peptide_library2,"peptide_library")

source("R/join_with_library.R")

joined_with_library = join_with_library(imputed)

# Doing all the clevages

## Calculating where all of the cleavages are at.

source("R/nterm_cleavage.R")
source("R/cterm_cleavage.R")

source("R/add_cleavages.R")

final_data = add_cleavages(joined_with_library)

## Now that we have the final data, we need to

```

### 2024-04-12 R/normalyze.R

@@ -1,0 +1,42 @@
+#' normalyze: this function normalyzes the data using the NormalyzerDE package. 
+#'
+#' @param prepared_data = this is the data that has been prepared
+#' @param design_matrix = this is the design matrix
+#' @param outdir = this is the directory you would like NormalyzerDE::nomralyzer to write output to. 
+#'
+#' @return
+#' files written to the outdir specified
+#' the normalyzed data is then read in as a data frame. 
+#' @export
+#'
+#' @examples
+normalyze = function(prepared_data,design_matrix,outdir = getwd()){
+  
+  # Extracting only the data
+  dataOnly = prepared_data[, design_matrix$sample]
+
+  #Extracting only the annotation columns
+  annotOnly = prepared_data[, !(colnames(prepared_data) %in% design_matrix$sample)]
+  
+  # Making a summarized Experiment object
+  sumExpObj = SummarizedExperiment::SummarizedExperiment(
+    as.matrix(dataOnly),
+    colData = design_matrix,
+    rowData = annotOnly)
+  
+
+  # Making the jobName be the date folowed by brief tag
+  jobName = paste0(Sys.Date(),"_msp-ms_normalyze_output")
+  ## Performing the normalization, this function outputs a bunch of stuff to a directory, including plots
+  NormalyzerDE::normalyzer(jobName= jobName,
+                           experimentObj = sumExpObj,
+                           outdir)
+  
+  ## Reading in the data frame we just produced that we care about ##
+  norm_data = readr::read_delim(paste0(outdir,"/",jobName,"/","median-normalized.txt")) %>% 
+    #Performing a reverse log2 transformation
+    dplyr::mutate(across(30:length(.),~2**.x))
+
+  return(norm_data)
+}
+

### 2024-04-12 mspms_shiny/about.Rmd
fixed a bunch of stuff.

---
title: "MSPMS-Shiny"
author: "Charlie Bayne"
date: "`r Sys.Date()`"
output:
  html_document:
        theme: cosmo
        code_folding: hide
        highlight: tango
        df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
ggplot2::theme_set(ggprism::theme_prism(palette = "viridis"))
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# About

This is a Shiny app designed to use the [mspms R package](https://github.com/baynec2/mspms) to analyze data from the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/)

# How to use

On the side bar, you will see two places to upload files.

1. The first file should be the LFQ file from PEAKS. This file should be a csv file.
2. The second file should be the ID file from PEAKS. This file should also be a csv file.

You will also see two tabs at the top of the page, About and Output.. 
About links to these instructions.

Output contains a rendering of the normalized data, and a button that says download. This button will allow you to download the data that has been processed as a .csv file.

Be patient with the output tab, it takes a while for the code to execute. When it is finished, the table will appear and the data will be ready for download.

### 2024-04-12 README.Rmd
fixed a bunch of stuff.
updated readme

@@
 
-The goal of msp-ms is provide a concise code-base for the normalization and data processing required for the msp-ms method.
+The goal of mspms is provide a concise code-base for the normalization and data processing required to analyze data from the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).
 
-Additionally, we provide a graphical user interface powered by shiny apps that allows for a user to utilize the method without requiring any R coding knowledge.
+Additionally, we provide a [graphical user interface powered by shiny apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to utilize the method without requiring any R coding knowledge.
+
+## Installation
+
+You can install the released version of mspms from github
+
+```{r,eval = FALSE}
+devtools::install_github("baynec2/mspms")
+```
+
+## Workflow
 
@@
 
-## Combining Peaks file outputs
+### Combining Peaks file outputs
 
@@
 ```{r}
+library(dplyr)
+library(mspms)
+
 ### Loading the files ###
@@
 
-## Normalizing values.
+### Normalizing values.
 
@@
 
-## Handling Outliers
+### Handling Outliers
 
-Here we use a dixion test from the outliers package to detect outliers from each of our replicates.
+Here we use a dixon test from the outliers package to detect outliers from each of our replicates.
 
@@
 
-## Imputation of data
+### Imputation of data
 
@@
 
-## Joining with Library
+### Joining with Library
 
@@
 
-## Calcuating clevages
+### Calcuating clevages
 Next, we need to determine where the peptide sequences are cleaved.
@@
 
-## Conducting stats
+### Conducting stats
 

### 2024-04-12 README.Rmd
updated

@@
 
-### Normalizing values.
+### Loading design matrix
+
+Before we normalyze data, we need to know what samples are in what groups. We can do that by defining a design matrix. 
+```{r}
+design_matrix = readr::read_csv("annotation.csv")
+
+head(design_matrix)
+
+```
+### Normalyzing data
 
@@
 
-First we need to extract a design matrix.
-
-Then we can normalyze the data. 
+Now we can normalyze the data.
 ```{r}
-# Extracting design matrix
-design_matrix = extract_design_matrix(prepared_data)
-
-# normalyze the data
-
-source("R/normalyze.R")
-
 normalyzed_data = normalyze(prepared_data,design_matrix)
@@
 Here we use a dixon test from the outliers package to detect outliers from each of our replicates.
+We need to know what samples are part of which groups, so we need to specify the design matrix here too.
+Make sure that the column header names are "sample" and "group" just like before.
 
 ```{r}
-outliers = handle_outliers(normalyzed_data)
+design_matrix = readr::read_csv("annotation.csv")
+outliers = handle_outliers(normalyzed_data,design_matrix)
 ```
@@
 final_data = add_cleavages(joined_with_library)
+
+head(final_data)
 ```

### 2024-04-12 R/handle_outliers.R

@@ -1,50 +1,50 @@
 #' handle_outliers
 #'
 #' @param normalyzed_data = this is the normalzed data.
+#' @param design_matrix = this is the design_matrix containing info on what sample is in what group
 #'
-#' @return a data frame in the long format where outliers according to a dixon test have been handled. 
+#' @return a data frame in the long format where outliers according to a dixon test have been handled.
 #' @export
 #'
 #' @examples
-handle_outliers = function(normalyzed_data){
-  
+handle_outliers = function(normalyzed_data,design_matrix){
+
   # Putting into the long format, adding group information, based on sample_time_replicate naming convention
-  
-  long_data = normalyzed_data %>% 
-    tidyr::pivot_longer(30:length(.)) %>% 
-    tidyr::separate(name,sep = "_",into = c("sample","time","replicate")) %>% 
-    dplyr::mutate(sample_id = paste0(sample,"_",time,"_",replicate),
-           group = paste0(sample,".",time)) %>% 
+
+
+  long_data = normalyzed_data %>%
+    tidyr::pivot_longer(30:length(.),names_to = "sample") %>%
+    dplyr::inner_join(design_matrix,by = "sample") %>%
     dplyr::filter(!is.na(value))
-  
-  
+
+
   # initalizing the data frame
   out = data.frame()
-  
+
   # Looping through each group to check for outliers.
   for(i in unique(long_data$group)){
-    
+
     #filtering data to only include the samples with the same condition and time
     f = dplyr::filter(long_data, group == i)
-    
+
     # Looping though each peptide within each group
     for(j in unique(long_data$Peptide)){
       f2 = dplyr::filter(f,Peptide == j )
       # # only doing tests if there are between 3 and 30 values
        if(length(f2$value) > 3 & length(f2$value) < 30){
-      #   
+      #
       #   #is there an outlier?
         stats = outliers::dixon.test(f2$value)$p.value
-        
+
         #what would the outlier be if there was one?
         outlier = outliers::outlier(f2$value)
-       
-        # If there is a outlier, let's change the value to 0 for now, we will change it to NA after we impute. 
+
+        # If there is a outlier, let's change the value to 0 for now, we will change it to NA after we impute.
         if(stats < 0.05){
           f2[f2$value == outlier,"value"] = 0
[…]

### 2024-04-13 README.Rmd
fix

@@
 Right now this is left to the user to do on their own. Will update after talkng to Bri.
+
+```{r}
+
+#Figuring out where the samples start
+start_of_samples = names(final_data)[which(names(final_data) == "z")+1]
+
+# Combining with design matrix
+long = final_data %>% 
+ tidyr::pivot_longer(start_of_samples:length(.),names_to = "sample") %>% 
+ inner_join(design_matrix,by = "sample")
+
+# What is the control group called?
+control_group = "DMSO_T0"
+
+# calculating log2 FC
+control_data = long %>% 
+  filter(group == control_group) %>% 
+  group_by(Peptide) %>% 
+  summarise(control_mean = mean(value,na.rm = TRUE))
+
+reference_group = "DMSO_T240"
+
+reference_data = long %>% 
+  filter(group == reference_group) %>% 
+  group_by(Peptide) %>% 
+  summarise(reference_mean = mean(value,na.rm = TRUE))
+
+log2fc = inner_join(control_data,reference_data,by = "Peptide") %>% 
+ dplyr::mutate(comparison = paste0(control_group,"_",reference_group),
+         log2fc = log2(reference_mean/control_mean)) %>% 
+  dplyr::select(Peptide,comparison,log2fc)
+
+# Doing T tests
+
+stat = long %>% 
+  filter(group %in% c(control_group,reference_group)) %>%
+  group_by(Peptide) %>%
+  rstatix::t_test(value ~ group) %>%
+  rstatix::adjust_pvalue(method = "fdr") %>% 
+  mutate(comparison = paste0(group1,"_",group2))
+
+combined = inner_join(log2fc,stat,by = c("Peptide","comparison"))
+
+```
+
+```{r}
+
+# Lets do anova to see if there are any differences among groups
+stat = long %>% 
+  group_by(Peptide) %>% 
+  rstatix::anova_test(formula = value ~ group) %>% 
+  rstatix::adjust_pvalue(method = "fdr")
+
+sig = stat %>% 
+  as.data.frame() %>% 
+  dplyr::filter(p.adj < 0.05) %>% 
+  dplyr::pull(Peptide)
+
+p1 = long %>%
+  mutate(peptide %in% stat)
+  ggplot(aes())
+
+```

### 2024-04-16 README.Rmd
major changes

--- 2024-04-12 README.Rmd
@@
   collapse = TRUE,
-  comment = "#>"
+  comment = "#>",
+  cache = TRUE
 )
@@
 6. Figures out the locations of the detected clevages within the library of peptide sequences used. Is it cleaved at the N or C terminus, or both?
+
+This package also contains tools for doing some basic statistics and data visualizations.
+
 7. Calculates the fold change and the p/q value across experimental conditions using T-tests.   
+8. Conducts ANOVA tests to determine if there are any peptides that are significantly different across all time points.  
+9. Conducts a hierarchical clustering analysis to visualize overall patterns in the data.
 
@@
 
-The files coming from peaks should have..... **Describe how to do this here**.
+All analysis performed by this package is downstream of the data generated by the PEAKS software. The first step is to combine the two files generated by PEAKS.
 
-Importantly, this. data contains... what????
+The files coming from peaks should be generated according to the instructions found [here](www/PeaksDataAnalysis_howto.pdf).
 
@@
 ### Loading the files ###
-lfq_filename = "legacy/protein-peptides-lfq.csv"
+lfq_filename = "test/protein-peptides-lfq.csv"
 #file "protein-peptides.csv" exported from PEAKS identification
-id_filename = "legacy/protein-peptides-id.csv"
+id_filename = "test/protein-peptides-id.csv"
 
@@
 prepared_data = prepare_for_normalyzer(lfq_filename,id_filename)
+
+# saving to data folder in package for tests. Ignore when using on your own data
+usethis::use_data(prepared_data,overwrite = TRUE)
 ```
@@
 ```{r}
-design_matrix = readr::read_csv("annotation.csv")
+design_matrix = readr::read_csv("test/design_matrix.csv")
 
@@
 normalyzed_data = normalyze(prepared_data,design_matrix)
+
+# saving to data folder in package for tests. Ignore when using on your own data
+usethis::use_data(normalyzed_data,overwrite = TRUE)
 ```
@@
 ```{r}
-design_matrix = readr::read_csv("annotation.csv")
+design_matrix = readr::read_csv("test/design_matrix.csv")
 outliers = handle_outliers(normalyzed_data,design_matrix)
+
+# saving to data folder in package for tests. Ignore when using on your own data
+usethis::use_data(outliers,overwrite = TRUE)
 ```
@@
 imputed = impute(outliers)
+
+# saving to data folder in package for tests. Ignore when using on your own data
+usethis::use_data(imputed,overwrite = TRUE)
 ```
@@
 joined_with_library = join_with_library(imputed)
+
+# saving to data folder in package for tests. Ignore when using on your own data
+usethis::use_data(joined_with_library,overwrite = TRUE)
 ```
 
-### Calcuating clevages
+### Calcuating clevages.
+
 Next, we need to determine where the peptide sequences are cleaved.
@@
 ```{r}
-final_data = add_cleavages(joined_with_library)
+cleavage_added_data = add_cleavages(joined_with_library)
 
-head(final_data)
+head(cleavage_added_data)
+
+# saving to data folder in package for tests. Ignore when using on your own data
+usethis::use_data(cleavage_added_data,overwrite = TRUE)
+```
+
+### polishing
+
+Someimes there is a need to polish the data a bit for downstream analysis.
+This function does that by removing combining the cterm and nterm cleavage information into one column while removing any rows that don't have any cleavage information or have cleavage information on the cterm and nterm.
+
+```{r}
+polished_data = polish(cleavage_added_data)
 ```
@@
 
-Right now this is left to the user to do on their own. Will update after talkng to Bri.
+mspms provides a number of convenience functions to conduct statistics on the data.
+
+First, we need to prepare the data for stats.
+This involves reshaping the data into the long format so it is easy to conduct statistics on and then appending the data in the design matrix that allows us to determine what sample contains what conditions/time and conduct the appropriate statistics.
+
+```{r}
+prepared_for_stats = prepare_for_stats(cleavage_added_data,design_matrix)
+
+usethis::use_data(prepared_for_stats,overwrite = TRUE)
+```
+
+Noe we can conduct the statistics.
+
+#### T tests
+
+Ttests are performed within each condition and compared to time 0.
+
+For example, for an experiment where there are two conditions, DMSO and MZB as well as 3 time points, 0, 1, and 2, the t tests would be as follows:
+
+* DMSO.TO vs DMSO.T0. 
+* DMSO.T0 vs DMSO.T1. 
+* DMSO.T0 vs DMSO.T2.
+
+* MZB.T0 vs MZB.T0. 
+* MZB.T0 vs MZB.T1. 
+* MZB.T0 vs MZB.T2.
+
+```{r}
+# Perform T test
+t_test_stats = mspms::mspms_t_tests(prepared_for_stats)
+```
+
+#### log2FC
+
+We can also calculate the log 2 fc. The comparisons here are the same as for the T tests
+```{r}
+# calclulate log2fc
+log2fc = mspms::mspms_log2fc(prepared_for_stats)
+
+head(log2fc)
+```
+
+#### log2fc_t_tests. 
+Sometimes (such as when you want to make volcano plots) it is useful to look at the log2fc and the t test statistics at the same time. This function does that.
+
+```{r}
+log2fc_t_test = mspms::log2fc_t_test(prepared_for_stats)
+
+head(log2fc_t_test)
+```
+
+Now that we have this data we can visualize these results easily using the ggplot2 package.
+
+```{r}
+library(ggplot2)
+
+p1 = log2fc_t_test %>% 
+  ggplot(aes(x = log2fc,y = -log10(p.adj)))+
+  geom_point(size =0.5)+
+  geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
+  geom_vline(xintercept = 3, linetype = "dashed",color = "red")+
+  geom_vline(xintercept = -3, linetype = "dashed",color = "red")+
+  theme_minimal()+
+  labs(x = "Log2 Fold Change",y = "-log10(p value)")+
+  facet_wrap(~comparison,scales = "free")
+
+p1
+```
+
+#### ANOVA
+
+We also might want to perfom an anova. Here, we have it set up to show the effect of time within each condition.
+
+For example, for an experiment where there are two conditions, DMSO and MZB as well as 3 time points, 0, 1, and 2, the anova would statistics for the effect of time for each peptide within DMSO or MZB.
+
+```{r}
+# Doing ANOVA
+anova_stats = mspms::mspms_anova(prepared_for_stats)
+
+head(anova_stats)
+```
+
+### Common Data Visualizations
+
+We also provide some functions that make common data visualizations easier.
+
+#### PCA
+
+We can generate a PCA plot to visualize the data. Here, the colors show the different time points while the shape shows the different conditions.
+
+```{r}
+
+mspms::plot_pca(prepared_for_stats)
+```
+
+#### Hierchical clustering
+
+We can also generate an interactive heatmap with the hierchical clustering results.
+
+The format of this readme does not allow for interactivity, so a static picture of the output is shown instead.
+
+```{r,eval = FALSE}
+
+mspms::plot_heatmap(prepared_for_stats)
+
+```
+
+![](www/plot_heatmap_output.png)

### 2024-04-16 mspms_shiny/about.Rmd
major changes

@@
 
-On the side bar, you will see two places to upload files.
+On the side bar, you will see three places to upload files.
 
-1. The first file should be the LFQ file from PEAKS. This file should be a csv file.
-2. The second file should be the ID file from PEAKS. This file should also be a csv file.
+1. The first file should be the LFQ file as exported from PEAKS. This file should be a csv file.
+
+This file should look like the following:
+
+```{r}
+DT::datatable(head(mspms::peaks_lfq,2))
+```
+
+2. The second file should be the ID file as exported from PEAKS. This file should also be a csv file.
+
+This file should look like the following:
+
+```{r}
+DT::datatable(head(mspms::peaks_id,2))
+```
+
+3. The third file should be the design matrix file. This should be a csv file.
+
+The design matrix file should be a csv file with the following columns: "sample" and "group".
+The sample column should contain the names of the samples as found in the PEAKS files.
+The group column should contain the group that the sample belongs to.
+This can be any thing, as long as it is the same per group.
+
+Here is an example of what the design matrix file should look like for the test files.
+
+```{r}
+DT::datatable(mspms::design_matrix)
+```
 

### 2024-04-16 troubleshooting.Rmd
major changes

---
title: "troubleshooting"
output: html_document
date: "2024-04-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

### Loading the files ###
lfq_filename = "test/protein-peptides-lfq_2.csv"
#file "protein-peptides.csv" exported from PEAKS identification
id_filename = "test/protein-peptides-id_2.csv"

# Prepare the data for normalyzer analysis
prepared_data = prepare_for_normalyzer(lfq_filename,id_filename)

design_matrix = readr::read_csv("test/design_matrix_2.csv")
normalyzed_data = normalyze(prepared_data,design_matrix)

outliers = handle_outliers(normalyzed_data,design_matrix)

imputed = impute(outliers)
joined_with_library = join_with_library(imputed)
final_data = add_cleavages(joined_with_library)

prepared_for_stats = prepare_for_stats(final_data,design_matrix)

anova = mspms_anova(prepared_for_stats)

ttests = mspms_t_tests(prepared_for_stats)

```

### 2024-04-16 R/check_for_cleavage_after_AA.R

@@ -1,0 +1,28 @@
+#' check_for_cleavage_after_AA
+#'
+#' This function checks if the cleavage site is after a vector of specified amino acids.
+#'
+#' Note that the cleavage sequence is intended to be represented as 4 AA to the left and right of cleavage site, as single letter codes.
+#'
+#' @param cleavage_seq = this is the cleavage sequecne that is being checked. Represented as 4 AA to the left and right of cleavage site
+#' @param AA_vector = this is a vector of AA that you want to check to see if the cleavage site is after. For example c("W","L","Y","F").
+#'
+#' @return
+#' a logical vector of the same length as the input cleavage_seq
+#' @export
+#'
+#' @examples
+#'
+#' #this is true because the cleavage site is after a W
+#' check_for_cleavage_after_AA("GLYWAFAA",c("W","L","Y","F"))
+#'
+#' #this is false because the cleavage site is not after a W, L, Y, or F
+#' check_for_cleavage_after_AA("GLYAAFAA",c("W","L","Y","F"))
+#'
+check_for_cleavage_after_AA = function(cleavage_seq,AA_vector){
+
+  # check if the cleavage site is in the AA_list
+  out = stringr::str_sub(cleavage_seq,4,4) %in% AA_list
+
+  return(out)
+}

### 2024-04-16 R/mspms_log2fc.R

@@ -1,0 +1,36 @@
+#' mspms_log2fc
+#'
+#' Calculates the log2fc for each timepoint within each condition relative to time 0.
+#'
+#' @param prepared_for_stats = this is the data that has been prepared for stats using the prepare for stats function.
+#'
+#' @return a data frame with the log2fc for each timepoint within each condition relative to time 0
+#' @export
+#'
+#' @examples
+mspms_log2fc = function(prepared_for_stats){
+
+  # Extracting just the control data
+  control_data = prepared_for_stats %>%
+    dplyr::filter(time == 0) %>%
+    dplyr::group_by(condition,Peptide) %>%
+    dplyr::summarise(control_mean = mean(value,na.rm = TRUE))
+
+
+  # Extracting just the reference data
+  reference_data = prepared_for_stats %>%
+    dplyr::group_by(Peptide,time,condition) %>%
+    dplyr::summarise(reference_mean = mean(value,na.rm = TRUE))
+
+  # Calculating the log2fc
+
+  log2fc = dplyr::inner_join(control_data,reference_data,by = c("Peptide","condition")) %>%
+    dplyr::mutate(comparison = paste0(condition,".T0","_",condition,".T",time),
+                  log2fc = log2(reference_mean/control_mean)) %>%
+    tibble::as.tibble()
+
+
+  return(log2fc)
+
+
+}

### 2024-04-16 R/normalyze.R

@@ -1,41 +1,55 @@
-#' normalyze: this function normalyzes the data using the NormalyzerDE package. 
+#' normalyze: This function performs normalization of MSP-MS data.
 #'
-#' @param prepared_data = this is the data that has been prepared
-#' @param design_matrix = this is the design matrix
-#' @param outdir = this is the directory you would like NormalyzerDE::nomralyzer to write output to. 
+#' This function uses the NormalyzerDE package to perform normalization. It returns the median normalized data as a data frame.
+#' For this normalization the Intensity of each variable in a given sample is divided by the median of intensities of all variables
+#' in the sample and then multiplied by the mean of median of sum of intensities of all variables in all samples.
+#'
+#' @param prepared_data = this is the data that has been prepared via mspms::prepare_for_normalyzer().
+#' @param design_matrix = this is the design matrix. This should be a data frame with a column named "sample" that contains the names of the samples in the data,
+#' a column named "group" that contains the group that the sample belongs to a column named "time" that contains the time the sample was incubated for,
+#' and on named "condition" that contains condition cooresponding to each sample (commonly some kind of protease inhibitor).
+#'
+#' @param outdir = this is the directory you would like NormalyzerDE::nomralyzer to write output to.Default is current working dir.
 #'
 #' @return
-#' files written to the outdir specified
-#' the normalyzed data is then read in as a data frame. 
+#' Files from NormalyzerDE are written to the outdir specified.
+#'
+#' The normalized data is read in as a data frame, and returned as a R object.
 #' @export
 #'
 #' @examples
 normalyze = function(prepared_data,design_matrix,outdir = getwd()){
-  
+
   # Extracting only the data
   dataOnly = prepared_data[, design_matrix$sample]
 
   #Extracting only the annotation columns
   annotOnly = prepared_data[, !(colnames(prepared_data) %in% design_matrix$sample)]
-  
+
   # Making a summarized Experiment object
   sumExpObj = SummarizedExperiment::SummarizedExperiment(
     as.matrix(dataOnly),
     colData = design_matrix,
     rowData = annotOnly)
-  
+
 
   # Making the jobName be the date folowed by brief tag
-  jobName = paste0(Sys.Date(),"_msp-ms_normalyze_output")
+  jobName = paste0(Sys.Date(),"_mspms_normalyze_output")
[…]

### 2024-04-16 R/nterm_cleavage.R

@@ -1,12 +1,13 @@
-#' nterm_cleavage:
+#' nterm_cleavage
 #'
-#' @param peptide_sequence = this is the peptide sequence measured. _ denotes clevage site.
+#' Finding the cleavage sequences on the N terminus of a given peptide in reference to the peptide library it was derived from.
+#'
+#' @param peptide_sequence = this is the peptide sequence in single leter AA code. _ denotes cleavage site..
 #' @param library_match_sequence  = this is the sequence of the corresponding peptide in the library to the peptide_sequence.
-#' @param library_real_sequence = this is the library match sequence with the Ls transformed to Ms (For some reason we don't understand.)
+#' @param library_real_sequence = this is the library match sequence with the Ls transformed to Ms (This is what the legacy code did so it is kept this way in case there was a good reason for it)
 #'
 #' @return
 #' a data frame with the peptide sequence, cleavage sequences 4 AA on the left and right of the n term cleavage, and the position of the n term cleavage in the library sequence.
-
 #' @export
 #'
 #' @examples
@@ -84,16 +85,3 @@ nterm_cleavage = function(peptide_sequence,library_match_sequence,library_real_s
   }
 
 
-# ###Testing
-# cleave_tab = readr::read_csv("legacy/protein-peptides_prepared.csv")
-#
-#
-# library(dplyr)
-#    test =  purrr::pmap_df(list( gsub("\\.","_",cleave_tab$Annotated_Sequence),cleave_tab$MatchSequence,cleave_tab$RealSequence),nterm_cleavage)
-#
-#
-#    test_vis = cbind(cleave_tab$RealSequence,cleave_tab$Annotated_Sequence,cleave_tab$`N-term`,test$nterm,cleave_tab$Cleave_Loc,test$nterm_cleavage_pos)
-#
-#    identical(cleave_tab$`N-term`,test$nterm)
-#    test_vis
-#

### 2024-04-16 R/plot_heatmap.R

@@ -1,0 +1,26 @@
+#' plot_heatmap
+#'
+#' This produces a heatmaply interactive heatmap of the data.
+#'
+#' @param cleavage_added_data = this is the data that has been processed by the add_cleavage() function.
+#' @param scale = how would you like the data scaled? default is none, but can be done by "row", "column", or "none"
+#'
+#' @return a heatmaply interactive heatmao
+#' @export
+#'
+#' @examples
+plot_heatmap = function(prepared_for_stats,scale = "column"){
+
+  heatmap_data = prepared_for_stats %>%
+    dplyr::select(sample,Peptide,condition,time,value) %>%
+    tidyr::pivot_wider(names_from = Peptide,values_from = value,values_fn = mean) %>%
+    tibble::column_to_rownames("sample") %>%
+    # if a peptide has an na remove it
+    dplyr::select_if(~ !any(is.na(.))) %>%
+    dplyr::mutate(time = as.factor(time))
+
+
+ plot =  heatmaply::heatmaply(heatmap_data,scale = scale,showticklabels = c(FALSE,TRUE))
+
+ return(plot)
+}

### 2024-04-16 mspms_shiny/app.R

@@ -11,7 +11,9 @@ ui <- dashboardPage(#skin = "midnight",
       sidebarMenu(
         # Making the tabs
         menuItem("About", tabName = "about", icon = icon("magnifying-glass-chart")),
-        menuItem("Output", tabName = "output", icon = icon("chart-simple"))
+        menuItem("Output", tabName = "output", icon = icon("database")),
+        menuItem("Stats", tabName = "stats", icon = icon("star-of-life")),
+        menuItem("DataViz", tabName = "viz", icon = icon("chart-simple"))
       ),
       #Putting the place to load the files
       fileInput("upload1", "Upload PEAKS LFQ file",accept = ".csv"),
@@ -31,19 +33,50 @@ ui <- dashboardPage(#skin = "midnight",
                 downloadButton(
                   "downloadData",
                   label = "Download"
-                )))
-                  )
-                  )
+                ))),
+        tabItem(tabName = "stats",
+                fluidRow(
+                  box(DT::DTOutput('ttest_results'),width = 6),
+                  box(DT::DTOutput('anova_results'),width = 6)),
+                downloadButton(
+                  "downloadttest",
+                  label = "Download t-test"
+                  ),
+                downloadButton(
+                  "downloadanova",
+                  label = "Download anova"
+                )
+                ),
+        tabItem(tabName = "viz",
+                fluidRow(
+                  box(plotOutput("plot1"),width = 6),
+                  box(plotOutput("plot2"),width = 6)
+                )
+                )
+        )
     )
+)
+
 
 
 # Define server logic
 server <- function(input, output){
 
+  # Rendering the about page
+  output$about <- renderUI({
+    includeHTML("./about.html")})
+
+
+  # Reading in the design matrix
+  design_matrix = reactive({
+    readr::read_csv(input$upload3$datapath)
+  })
+
+  # Processing the data normalization data
   processed_data = reactive({
         prepared_data =  mspms::prepare_for_normalyzer(input$upload1$datapath,input$upload2$datapath)
         # mspms workflow
-        design_matrix=readr::read_csv(input$upload3$datapath)
+        design_matrix = design_matrix()
[…]

### 2024-04-16 mspms_shiny/about.Rmd
more changes

@@
         theme: cosmo
-        code_folding: hide
         highlight: tango
@@
 
-This file should look like the following:
-
-```{r}
-DT::datatable(head(mspms::peaks_lfq,2))
-```
-
 2. The second file should be the ID file as exported from PEAKS. This file should also be a csv file.
-
-This file should look like the following:
-
-```{r}
-DT::datatable(head(mspms::peaks_id,2))
-```
 
@@
 
-The design matrix file should be a csv file with the following columns: "sample" and "group".
+The design matrix file should be a csv file with the following columns: "sample", "group","condition","time".
 The sample column should contain the names of the samples as found in the PEAKS files.
@@
 This can be any thing, as long as it is the same per group.
+The time column should contain the duration of incubation, and the condition column should include any experimental condition such as protease inhibitor.
 
-Here is an example of what the design matrix file should look like for the test files.
+You will also see four tabs at the side of the page, About, Output, Stats, and DataViz
 
-```{r}
-DT::datatable(mspms::design_matrix)
-```
-
-You will also see two tabs at the top of the page, About and Output.. 
-About links to these instructions.
+About links to the instructions shown here
 
@@
 Be patient with the output tab, it takes a while for the code to execute. When it is finished, the table will appear and the data will be ready for download.
+
+Stats contains a table of the results of the t-tests and ANOVA tests. Be patient, when it is finished the talbe will appear and the data will be ready for download.
+
+DataViz contains a PCA plot, heatmap of the data, and volcano plots of all the data.  Be patient,this takes a while to render.

### 2024-04-16 R/plot_heatmap.R

@@ -20,7 +20,7 @@ plot_heatmap = function(prepared_for_stats,scale = "column"){
     dplyr::mutate(time = as.factor(time))
 
 
- plot =  heatmaply::heatmaply(heatmap_data,scale = scale,showticklabels = c(FALSE,TRUE))
+ heatmaply::heatmaply(heatmap_data,scale = scale,showticklabels = c(FALSE,TRUE))
 
- return(plot)
+
 }

### 2024-04-16 legacy/MSP_data_processing_1.R

@@ -1,79 +1,0 @@
-#enter directory location
-directory <- "legacy/R Scripts New"
-#file "protein-peptides.csv" exported from PEAKS label-free quantification
-lfq_filename <- "legacy/protein-peptides-lfq.csv"
-#file "protein-peptides.csv" exported from PEAKS identification
-id_filename <- "legacy/protein-peptides-id.csv"
-#specify mass spec samples, replicates should be denoted with the same numbers
-sample <- c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,4)
-
-
-
-##Optioanl, whether to fix random number generation
-#set.seed(12345)
-numsample <- length(sample)
-library(data.table)
-library(qvalue)
-library(fitdistrplus)
-library(truncnorm)
-library(gtools)
-
-#set working directory, all files should be in this directory
-
-#setwd(directory)
-
-#function "FDRcorrection", find peptides < 1%FDR
-FDRcorrection <- function(lfq, id){
-        df <- read.csv(lfq, sep = ",", header=TRUE, row.names = NULL, stringsAsFactors = FALSE)
-        refdata <- read.csv(id, sep = ",", header=TRUE, row.names = NULL, stringsAsFactors = FALSE)
-        rels <- c(refdata[,4])
-        ls <- NULL
-        for (i in 1:nrow(df)){
-                ls <- c(ls, df[i,1:4], refdata[(which(rels == df[i,4])[1]),6:12], df[i,5:ncol(df)],which(rels == df[i,4])[1])
-        }
-        data <- as.matrix.data.frame(matrix(ls,ncol=(ncol(df)+8),byrow=TRUE))
-        cnames1 <- colnames(df[1:4])
-        cnames2 <- colnames(refdata[6:12])
-        cnames3 <- colnames(df[5:ncol(df)])
-        colnames(data) <- c(cnames1,cnames2,cnames3,"matchwhich")
-        data<-data[complete.cases(data[ , ncol(data)]),]
-}
-
-#function "matfornorm", filter out Quality <= 0.3 and prepare the file for Normalyzer
-matfornorm <- function (sample,df){
-        data <- NULL
-        ls <- NULL
-        cnames <- NULL
-        
-        # Literally just filtering by quality score in the most confusing way. 
[…]

### 2024-04-18 README.Rmd
updated n term and cterm cleavage functions to allow for user to specify n of AA to the side of cleavage event. Also updated docs and down stream functions.
A couple of other minor things in here too

@@
 
-Sequences are presented as the 4 amino acids on both sides of a cleavage. X indicates that there was nothing on that side in the library because it was cleaved close to the edge.
+Sequences are presented as the user specified number of amino acids on both sides of a cleavage.
+The default is 4, but there is interest in looking at motifs that are farther away from the cut site.
+Note that X indicates that there was nothing on that side in the library because it was cleaved close to the edge.
 
 ```{r}
-cleavage_added_data = add_cleavages(joined_with_library)
+cleavage_added_data = add_cleavages(joined_with_library,n_residues = 4)
 

### 2024-04-18 tests/testthat.R

@@ -1,0 +1,12 @@
+# This file is part of the standard setup for testthat.
+# It is recommended that you do not modify it.
+#
+# Where should you do additional test configuration?
+# Learn more about the roles of various files in:
+# * https://r-pkgs.org/testing-design.html#sec-tests-files-overview
+# * https://testthat.r-lib.org/articles/special-files.html
+
+library(testthat)
+library(mspms)
+
+test_check("mspms")

### 2024-04-18 README.Rmd
added ability to plot time course.

@@
 ![](www/plot_heatmap_output.png)
+
+#### Ploting time course.
+
+We also provide a function for plotting the mean intensity and standard deviation over time for each peptide in the data set by conditions. Facets show peptide, color shows the condition.
+
+This is best used in combination with the ANOVA function. Below, we will use anova to calculate the peptides where we see an effect of time (in the DMSO group), and then plot these.
+
+```{r}
+
+top10sig = anova_stats %>% 
+  arrange(p.adj) %>%
+  pull(Peptide) %>% 
+  head(10)
+
+p1 = prepared_for_stats %>%
+  dplyr::filter(Peptide %in% top10sig) %>%
+  mspms::plot_time_course()
+
+p1
+```

### 2024-04-22 README.Rmd
added function to calculate all possible cleavages.

@@
 
-## Workflow
+## Overview
 
-So how does the msps data normalization process work?
+There are 4 different types of functions in this package.  Those that are involved in:
 
-1. Takes two input files from PEAKS and combines them
-3. Normalizes values and then does a reverse log2 transformation
-4. Looks for outliers across replicates. Removes them. 
-5. Imputes data for missing values (likely to be very low intesity).
-6. Figures out the locations of the detected clevages within the library of peptide sequences used. Is it cleaved at the N or C terminus, or both?
+1.
+Making mspms generically useful.
+These functions are focused on making mspms more generally useful to a wider audience outside of the very specific workflow traditionally used in the O’Donoghue lab.
+Allows for the use of other types of upstream proteomic data processing, different peptide libraries, etc.
 
-This package also contains tools for doing some basic statistics and data visualizations.
+2. Data processing/ normalization. These functions allow the user to normalize and process the MSP-MS data.
 
-7. Calculates the fold change and the p/q value across experimental conditions using T-tests.   
-8. Conducts ANOVA tests to determine if there are any peptides that are significantly different across all time points.  
-9. Conducts a hierarchical clustering analysis to visualize overall patterns in the data.
+3. Statistics. These methods allow the user to perform basic statistics on the normalyzed/processed data.
+
+4. Data visualization. These functions allow the user to visualize the data in a number of useful way.
+
+**Making mspms generically useful**.
+
+1. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.
+
+**Data Proessing/ Normalization**.
+
+1. *prepare_for_normalyzer()*:Takes two input files from PEAKS and combines them.    
+2. *normalyze()*: Normalizes values and then does a reverse log2 transformation.    
+3. *handle_outliers()*: Looks for outliers across replicates. Removes them.     
+4. *impute()*: Imputes data for missing values (not including NAs introduced by handle_outliers()).    
+5. *join_with_library()*: Joins the normalyzed data with the peptide library sequences.  
+6.
+*add_cleavages()*: Figures out the locations of the detected clevages within the library of peptide sequences used.
+Is it cleaved at the N or C terminus, or both?
+Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.
+7. *polish()*: combines the cleavage info into one column. Discards peptides that were cleaved on both sides or not at all.  
+8. *prepare_for_stats()*: Reshapes the data into a long format and appends the data in the design matrix to make statistical testing easy to perform.
+
+**Statistics**.   
+1. *log2fc_t_test()*: Calculates the fold change and the p/q value across experimental conditions using T-tests.      
+2. *mspms_anova()*: Conducts ANOVA tests to determine if there are any peptides that are significantly different across all time points.
+
+**Data Visualization**.   
+1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an interactive heatmap to visualize overall patterns in the data.    
+2. *plot_pca()*: PCA analysis to visualize the data in a 2D space.  
+3. *plot_time_course()*: Plot peptides over time by condition.  
+4. *plot_cleavage_motif()*: Visualize the sequence specificity of the cleavage sites. Designed to be similar to what is implemented in IceLogo.
+
+## Making Generically Usefull
+
+### Calculating all cleavages
+
+We might want to calculate all possible cleavages for the peptide library sequences.
+This is useful for downstream analysis, especially when we are looking at the specificity of the cleavage sites via  as this requires a background of all possible cleavages
+
+We can do this by specifying the number of amino acids after the cleavage site that we are interested in. First lets try 4, which is the default.
+
+```{r}
+all_peptide_sequences = mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
+                        n_AA_after_cleavage=4)
+head(all_peptide_sequences)
+
+```
+
+We could also try 5 AA after each cleavage site
+
+```{r}
+all_peptide_sequences = mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
+                        n_AA_after_cleavage=5)
+head(all_peptide_sequences)
+```
+
+## Data Normalization/ Processing
 
@@
 ### Loading the files ###
-lfq_filename = "test/protein-peptides-lfq.csv"
+lfq_filename = "tests/protein-peptides-lfq.csv"
 #file "protein-peptides.csv" exported from PEAKS identification
-id_filename = "test/protein-peptides-id.csv"
+id_filename = "tests/protein-peptides-id.csv"
 
@@
 ```{r}
-design_matrix = readr::read_csv("test/design_matrix.csv")
+design_matrix = readr::read_csv("tests/design_matrix.csv")
 
@@
 ```{r}
-design_matrix = readr::read_csv("test/design_matrix.csv")
+design_matrix = readr::read_csv("tests/design_matrix.csv")
 outliers = handle_outliers(normalyzed_data,design_matrix)
@@
 
-### polishing
+### Polishing
 
@@
 
-### Conducting stats
+## Statistics
 
@@
 
-#### T tests
+### T tests
 
@@
 
-#### log2FC
+### log2FC
 
@@
 
-#### log2fc_t_tests. 
+### log2fc_t_tests.
+
 Sometimes (such as when you want to make volcano plots) it is useful to look at the log2fc and the t test statistics at the same time. This function does that.
@@
 
-#### ANOVA
+### ANOVA
 
@@
 
-### Common Data Visualizations
+## Common Data Visualizations
 
@@
 
-#### PCA
+### Plotting sequence specificity motif
+
+Here use an approach similar to what is implemented in ICELogo to visualize the sequence specificity of the cleavage sites.
+This was used as reference to build the code: https://iomics.ugent.be/icelogoserver/resources/manual.pdf.
+
+```{r}
+cleavage_seqs = mspms::prepared_for_stats %>%
+  mspms::polish() %>% 
+  filter(condition == "DMSO",time == 240) %>%
+  pull(cleavage_seq)
+
+background_universe = mspms::all_possible_8mers_from_228_library
+
+mspms::plot_cleavage_motif(cleavage_seqs,background_universe)
+```
+
+```{r}
+test_seq = readLines("tests/cleavage_seq_test.txt")
+
+mspms::plot_cleavage_motif(test_seq,background_universe)
+```
+
+### PCA
 
@@
 ```{r}
-
 mspms::plot_pca(prepared_for_stats)
@@
 
-#### Hierchical clustering
+### Hierchical clustering
 
@@
 
-#### Ploting time course.
+### Ploting time course.
 

### 2024-04-22 R/handle_outliers.R

@@ -10,6 +10,9 @@
 #'
 #' @examples
 #'
+#' # Here we handle the outliers from the test data
+#'handled_outliers = handle_outliers(mspms::normalyzed_data,mspms::design_matrix)
+
 handle_outliers = function(normalyzed_data,design_matrix){
 
   # Putting into the long format, adding group information, based on sample_time_replicate naming convention

### 2024-04-22 R/mspms_log2fc.R

@@ -8,6 +8,8 @@
 #' @export
 #'
 #' @examples
+#'
+#' log2fc = mspms_log2fc(mspms::prepared_for_stats)
 mspms_log2fc = function(prepared_for_stats){
 
   # Extracting just the control data

### 2024-04-22 R/normalyze.R

@@ -14,10 +14,13 @@
 #' @return
 #' Files from NormalyzerDE are written to the outdir specified.
 #'
-#' The normalized data is read in as a data frame, and returned as a R object.
+#' The normalized data is returned as a tibble.
 #' @export
 #'
 #' @examples
+#'
+#' normalyzed_data = normalyze(mspms::prepared_data,mspms::design_matrix,outdir = getwd())
+#'
 normalyze = function(prepared_data,design_matrix,outdir = getwd()){
 
   # Extracting only the data

### 2024-04-22 R/plot_pca.R

@@ -9,7 +9,7 @@
 #'
 #' @examples
 #'
-#'
+#'plot_pca(mspms::prepared_for_stats)
 #'
 #'
 plot_pca = function(prepared_for_stats){

### 2024-04-22 R/polish.R

@@ -1,0 +1,24 @@
+#' polish
+#'
+#' This function is used to polish the cleavage data. It combines the nterm and cterm cleavage information into a single column and removes rows that
+#' don't have any cleavage information or have both nterm and cterm cleavage information.
+#'
+#' @param cleavage_added_data
+#'
+#' @return a data frame with the cleavage information combined into a single column and rows with no cleavage information or double information removed.
+#' @export
+#'
+#' @examples
+#' polished = polish(mspms::cleavage_added_data)
+polish = function(cleavage_added_data){
+
+  out = cleavage_added_data %>%
+    dplyr::mutate(cleavage_seq = dplyr::case_when(!is.na(nterm) & is.na(cterm) ~ nterm,
+                                                  !is.na(cterm) & is.na(nterm) ~ cterm,
+                                                  TRUE ~NA),.after = "cterm_cleavage_pos") %>%
+    dplyr::filter(!is.na(cleavage_seq)) %>%
+    tibble::as_tibble()
+
+
+  return(out)
+}

### 2024-04-22 R/post_process.R

@@ -1,23 +1,0 @@
-#' polish
-#'
-#' This function is used to polish the cleavage data. It combines the nterm and cterm cleavage information into a single column and removes rows that
-#' don't have any cleavage information or have both nterm and cterm cleavage information.
-#'
-#' @param cleavage_added_data
-#'
-#' @return a data frame with the cleavage information combined into a single column and rows with no cleavage information or double information removed.
-#' @export
-#'
-#' @examples
-polish = function(cleavage_added_data){
-
-  out = cleavage_added_data %>%
-    dplyr::mutate(cleavage_seq = dplyr::case_when(!is.na(nterm) & is.na(cterm) ~ nterm,
-                                                  !is.na(cterm) & is.na(nterm) ~ cterm,
-                                                  TRUE ~NA),.after = "cterm_cleavage_pos") %>%
-    dplyr::filter(!is.na(cleavage_seq)) %>%
-    tibble::as.tibble()
-
-
-  return(out)
-}

### 2024-04-22 R/prepare_for_stats.R

@@ -5,10 +5,11 @@
 #' @param cleavage_added_data = this is the processed data with cleavage information added. Intended to be downstream of the add_cleavages() function.
 #' @param design_matrix = this is the design matrix with "sample",group","time","condition" columns.
 #'
-#' @return a data frame in long format with the design matrix combined.
+#' @return a tibble in long format with the design matrix combined.
 #' @export
 #'
 #' @examples
+#' prepare_for_stats(mspms::cleavage_added_data,mspms::design_matrix)
 prepare_for_stats = function(cleavage_added_data,design_matrix){
 
   #Figuring out where the samples start

### 2024-04-26 README.Rmd
made changes in the data that is retained to allow for the development of different parsers for outputs from different software

@@
 
-# msp-ms
+# mspms
 
@@
 **Making mspms generically useful**.   
-
-1. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.
+1. *prepare_peaks()*:Takes two input files from PEAKS and combines them. 
+2. *prepare_pd()*: prepares exported files from proteome discoverer.  
+  - *extract_design_matrix_pd()*: Extracts the design matrix from the proteome discoverer file.  
+3. *prepare_spectronaut()*: prepares exported files from spectronaut - note in development, not yet implemented.  
+4. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.
 
 **Data Proessing/ Normalization**. 
-
-1. *prepare_for_normalyzer()*:Takes two input files from PEAKS and combines them.    
-2. *normalyze()*: Normalizes values and then does a reverse log2 transformation.    
-3. *handle_outliers()*: Looks for outliers across replicates. Removes them.     
-4. *impute()*: Imputes data for missing values (not including NAs introduced by handle_outliers()).    
-5. *join_with_library()*: Joins the normalyzed data with the peptide library sequences.  
-6.
+1. *normalyze()*: Normalizes values and then does a reverse log2 transformation.    
+2. *handle_outliers()*: Looks for outliers across replicates. Removes them.     
+3. *impute()*: Imputes data for missing values (not including NAs introduced by handle_outliers()).    
+4. *join_with_library()*: Joins the normalyzed data with the peptide library sequences.  
+5.
 *add_cleavages()*: Figures out the locations of the detected clevages within the library of peptide sequences used.
@@
 Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.
-7. *polish()*: combines the cleavage info into one column. Discards peptides that were cleaved on both sides or not at all.  
-8. *prepare_for_stats()*: Reshapes the data into a long format and appends the data in the design matrix to make statistical testing easy to perform.
+6. *polish()*: combines the cleavage info into one column. Discards peptides that were cleaved on both sides or not at all.  
+7. *prepare_for_stats()*: Reshapes the data into a long format and appends the data in the design matrix to make statistical testing easy to perform.
 
@@
 
-## Making Generically Usefull
+## Making Generically Usefull.
+
+### Formating Peaks File Outputs.
+
+All analysis performed by this package is downstream of the data generated by the PEAKS software. The first step is to combine the two files generated by PEAKS.
+
+The files coming from peaks should be generated according to the instructions found [here](www/PeaksDataAnalysis_howto.pdf).
+
+```{r}
+library(dplyr)
+library(mspms)
+
+### Loading the files ###
+lfq_filename = "tests/protein-peptides-lfq.csv"
+#file "protein-peptides.csv" exported from PEAKS identification
+id_filename = "tests/protein-peptides-id.csv"
+
+# Prepare the data for normalyzer analysis
+prepared_data = prepare_peaks(lfq_filename,id_filename)
+
+# saving to data folder in package for tests. Ignore when using on your own data
+usethis::use_data(prepared_data,overwrite = TRUE)
+```
 
@@
 ## Data Normalization/ Processing
-
-### Combining Peaks file outputs
-
-All analysis performed by this package is downstream of the data generated by the PEAKS software. The first step is to combine the two files generated by PEAKS.
-
-The files coming from peaks should be generated according to the instructions found [here](www/PeaksDataAnalysis_howto.pdf).
-
-```{r}
-library(dplyr)
-library(mspms)
-
-### Loading the files ###
-lfq_filename = "tests/protein-peptides-lfq.csv"
-#file "protein-peptides.csv" exported from PEAKS identification
-id_filename = "tests/protein-peptides-id.csv"
-
-# Prepare the data for normalyzer analysis
-prepared_data = prepare_for_normalyzer(lfq_filename,id_filename)
-
-# saving to data folder in package for tests. Ignore when using on your own data
-usethis::use_data(prepared_data,overwrite = TRUE)
-```
 
@@
 
-```{r}
-test_seq = readLines("tests/cleavage_seq_test.txt")
-
-mspms::plot_cleavage_motif(test_seq,background_universe)
-```
-
 ### PCA

### 2024-04-26 R/handle_outliers.R

@@ -16,10 +16,10 @@
 handle_outliers = function(normalyzed_data,design_matrix){
 
   # Putting into the long format, adding group information, based on sample_time_replicate naming convention
-  index = which(names(normalyzed_data) == "z")+1
+  index = which(names(normalyzed_data) == "Protein Accession")+1
 
   long_data = normalyzed_data %>%
-    tidyr::pivot_longer(index:length(.),names_to = "sample") %>%
+    tidyr::pivot_longer(dplyr::all_of(index:length(.)),names_to = "sample") %>%
     dplyr::inner_join(design_matrix,by = "sample") %>%
     dplyr::filter(!is.na(value))
 
@@ -65,12 +65,11 @@ handle_outliers = function(normalyzed_data,design_matrix){
     dplyr::select(-group,-condition,-time) %>%
     tidyr::pivot_wider(names_from = sample, values_from = value) %>%
     tidyr::pivot_longer(index:length(.),names_to = "sample_id") %>%
-    tibble::as.tibble()
+    tibble::as_tibble()
 
 
   return(NAs_added_back)
 }
-
 
 
 

### 2024-04-26 R/plot_cleavage_motif.R

@@ -10,15 +10,10 @@
 #' @export
 #'
 #' @examples
-#'
-#'cleavage_seqs = mspms::cleavage_added_data %>%
-#'filter(condtion == "DMSO",time == 240) %>%
-#'pull(cleavage_seq)
-#'
-#'background_universe = mspms::mspms::all_possible_8mers_from_228_library
-#'
-#'
-#'plot_clavage_motif(cleavage_seqs,background_universe)
+#'cleavage_seqs = mspms::prepared_for_stats %>%
+#' mspms::polish() %>%
+#' filter(condition == "DMSO",time == 240) %>%
+#' pull(cleavage_seq)
 
 plot_cleavage_motif = function(cleavage_seqs,background_universe = mspms::all_possible_8mers_from_228_library){
 
@@ -37,17 +32,28 @@ plot_cleavage_motif = function(cleavage_seqs,background_universe = mspms::all_po
   # Counting the number in each position
   bg_count = bg_seq %>%
     purrr::map_df(table) %>%
-    t() %>%
-    as.data.frame() %>%
-    tibble::rownames_to_column("AA")
+    t()
 
-  #Fixing names
-  names(bg_count) = c("AA",paste0("position",1:nchar_bg))
+  #replacing NA with 0
+  bg_count[is.na(bg_count)] = 0
 
   bg_prop = bg_count %>%
+    as.data.frame() %>%
+    tibble::rownames_to_column("AA") %>%
     mutate(across(where(is.numeric), prop.table)) %>%
     tibble::column_to_rownames("AA") %>%
     as.matrix()
+
+  ### calculating the SD ###
+
+  # defining function per Icelogo manual
+  standard_deviation = function(proportion_mat,count_mat){
+    (proportion_mat/colSums(count_mat))^(1/2)
+  }
+
+  # applying to bg data
+  bg_sd = standard_deviation(bg_prop,bg_count)
+
 
   bg_bits = mspms::calc_bits(bg_prop)
 
@@ -67,27 +73,75 @@ plot_cleavage_motif = function(cleavage_seqs,background_universe = mspms::all_po
 
  clev_count = clev_seq %>%
    purrr:::map_df(table) %>%
-   t() %>%
-   as.data.frame() %>%
-   tibble::rownames_to_column("AA")
+   t()
 
+  #replacing NA with 0
   clev_count[is.na(clev_count)] = 0
[…]

### 2024-04-26 R/prepare_for_stats.R

@@ -13,7 +13,7 @@
 prepare_for_stats = function(cleavage_added_data,design_matrix){
 
   #Figuring out where the samples start
-  start_of_samples = names(cleavage_added_data)[which(names(cleavage_added_data) == "z")+1]
+  start_of_samples = which(names(cleavage_added_data) == "cterm_cleavage_pos")+1
 
   # make into long format, combine with design matrix.
   long = cleavage_added_data %>%

### 2024-04-27 README.Rmd
fixed to clear all devtools::check() warnings

@@
   comment = "#>",
-  cache = TRUE
+  cache = FALSE
 )
@@
 ### Loading the files ###
-lfq_filename = "tests/protein-peptides-lfq.csv"
+lfq_filename = "tests/testdata/protein-peptides-lfq.csv"
 #file "protein-peptides.csv" exported from PEAKS identification
-id_filename = "tests/protein-peptides-id.csv"
+id_filename = "tests/testdata/protein-peptides-id.csv"
 
 # Prepare the data for normalyzer analysis
-prepared_data = prepare_peaks(lfq_filename,id_filename)
+peaks_prepared_data = prepare_peaks(lfq_filename,id_filename)
 
 # saving to data folder in package for tests. Ignore when using on your own data
-usethis::use_data(prepared_data,overwrite = TRUE)
+usethis::use_data(peaks_prepared_data,overwrite = TRUE)
 ```
@@
 ```{r}
-design_matrix = readr::read_csv("tests/design_matrix.csv")
+design_matrix = readr::read_csv("tests/testdata/design_matrix.csv")
 
@@
 
-msp-ms uses the .... package to do normalization under the hood.
+msp-ms uses the normalyzede package to do normalization under the hood.
 
@@
 ```{r}
-normalyzed_data = normalyze(prepared_data,design_matrix)
+normalyzed_data = normalyze(peaks_prepared_data,design_matrix)
 
@@
 ```{r}
-design_matrix = readr::read_csv("tests/design_matrix.csv")
+design_matrix = readr::read_csv("tests/testdata/design_matrix.csv")
 outliers = handle_outliers(normalyzed_data,design_matrix)

### 2024-04-27 R/impute.R

@@ -67,8 +67,8 @@ impute = function(outlier_handled_data,noise=0.05,nsd=1){
     tidyr::pivot_wider(names_from = "sample_id",
                        values_from = imputed_values) %>%
     dplyr::mutate(Peptide = gsub("\\.","_",Peptide),
-                  Peptide_no_cleavage = gsub("_","",Peptide),.before = 5) %>%
-    tibble::as.tibble()
+                  Peptide_no_cleavage = gsub("_","",Peptide),.after = Peptide) %>%
+    tibble::as_tibble()
 
   return(output)
 

### 2024-04-27 R/prepare_peaks.R

@@ -1,0 +1,61 @@
+#' prepare_peaks
+#' This function prepares the data out from peaks for downstream normalization via normalyzer.
+#' Briefly, it combines the two files by peptide IDs and filters out peptides that have quality scores <= 0.3. 0s are also replaced with NAs
+#' Note that there can be multiple matches between these, currently only keeps the values for the best quality scores. This could be a bug.
+#'
+#' @param lfq_filepath = this is the filepath to the first PEAKS output table
+#' @param id_filepath = this is the filepath to the second PEAKS output table
+#'
+#' @return a data frame containing the combined columns from the lfq and id files, with quality scores > 0.3 and 0s replaced with NAs.
+#' @export
+#'
+#' @examples
+#'
+#' prepare_peaks("tests/protein-peptides-lfq.csv","tests/protein-peptides-id.csv")
+#'
+prepare_peaks = function(lfq_filepath,
+                         id_filepath){
+
+  # Reading in the label free quantification data
+  lfq = readr::read_csv(lfq_filepath,guess_max = 10)
+
+  # Reading in the ids
+  id = readr::read_csv(id_filepath) %>%
+    #only keep the Peptide ID with the highest score in case there are more than one.
+    # This is essentially what the current script does, is it intended? Seems like a bug to me
+    dplyr::group_by(Peptide) %>%
+    dplyr::filter(`-10lgP` == max(`-10lgP`)) %>%
+    dplyr::ungroup() %>%
+    dplyr::select(Peptide,RT,6:12)
+
+
+  # Combining data frame and filtering to only contain quality scores > 0.3
+  output = lfq %>%
+    # Sometimes a peptide is detected more than once at different retention times.
+    # We will only keep the one with the highest quality score.
+    dplyr::group_by(Peptide) %>%
+    dplyr::filter(Quality == max(Quality)) %>%
+    dplyr::ungroup() %>%
+    dplyr::inner_join(id,by = c("Peptide"),multiple = "first") %>%
+    dplyr::filter(Quality > 0.3) %>%
+    tibble::as.tibble()
+
+
+  # finding the columns with our samples
+
+  start = which(names(output) == "Avg. Area") + 1
+  end = which(names(output) == "Sample Profile (Ratio)") - 1
+
+  # Selecting only the columns that we care about
+  output = output %>%
+    dplyr::mutate(Peptide = gsub("\\.","_",Peptide)) %>%
[…]

### 2024-04-27 tests/testthat/test-cterm_cleavage.R

@@ -2,7 +2,7 @@
 testthat::test_that("works with no cleavage site", {
   testthat::expect_equal(
     mspms::cterm_cleavage("ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",4),
-    data.frame(peptide = "ABCDEFGHIJKLMN",
+    tibble::tibble(peptide = "ABCDEFGHIJKLMN",
                cterm = NA,
                cterm_cleavage_pos = NA)
 )
@@ -13,7 +13,7 @@ testthat::test_that("works with no cleavage site", {
 testthat::test_that("works with one clevage site", {
   testthat::expect_equal(
     mspms::cterm_cleavage("ABCDEFGHIJKLM_N", "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",4),
-    data.frame(peptide = "ABCDEFGHIJKLM_N",
+    tibble::tibble(peptide = "ABCDEFGHIJKLM_N",
                cterm = "JKLMNXXX",
                cterm_cleavage_pos = 13)
   )
@@ -24,7 +24,7 @@ testthat::test_that("works with one clevage site", {
 testthat::test_that("works with two clevage site", {
   testthat::expect_equal(
     mspms::cterm_cleavage("A_BCDEFGHIJKLM_N", "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",4),
-    data.frame(peptide = "A_BCDEFGHIJKLM_N",
+    tibble::tibble(peptide = "A_BCDEFGHIJKLM_N",
                cterm = "JKLMNXXX",
                cterm_cleavage_pos = 13)
   )
@@ -35,7 +35,7 @@ testthat::test_that("works with two clevage site", {
 testthat::test_that("works with 6 AA past", {
   testthat::expect_equal(
     mspms::cterm_cleavage("A_BCDEFGHIJKLM_N", "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",6),
-    data.frame(peptide = "A_BCDEFGHIJKLM_N",
+    tibble::tibble(peptide = "A_BCDEFGHIJKLM_N",
                cterm = "HIJKLMNXXXXX",
                cterm_cleavage_pos = 13)
   )
@@ -46,7 +46,7 @@ testthat::test_that("works with 6 AA past", {
 testthat::test_that("works with 10 AA past", {
[…]

### 2024-04-27 tests/testthat/test-nterm_cleavage.R

@@ -2,7 +2,7 @@
 testthat::test_that("works with no cleavage site", {
   testthat::expect_equal(
     mspms::nterm_cleavage("ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",4),
-    data.frame(peptide = "ABCDEFGHIJKLMN",
+    tibble::tibble(peptide = "ABCDEFGHIJKLMN",
                nterm = NA,
                nterm_cleavage_pos = NA)
   )
@@ -13,7 +13,7 @@ testthat::test_that("works with no cleavage site", {
 testthat::test_that("works with one clevage site", {
   testthat::expect_equal(
     mspms::nterm_cleavage("A_BCDEFGHIJKLM_N", "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",4),
-    data.frame(peptide = "A_BCDEFGHIJKLM_N",
+    tibble::tibble(peptide = "A_BCDEFGHIJKLM_N",
                nterm = "XXXABCDE",
                nterm_cleavage_pos = 1)
   )
@@ -24,7 +24,7 @@ testthat::test_that("works with one clevage site", {
 testthat::test_that("works with two clevage site", {
   testthat::expect_equal(
     mspms::nterm_cleavage("A_BCDEFGHIJKLM_N", "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",4),
-    data.frame(peptide = "A_BCDEFGHIJKLM_N",
+    tibble::tibble(peptide = "A_BCDEFGHIJKLM_N",
                nterm = "XXXABCDE",
                nterm_cleavage_pos = 1)
   )
@@ -35,7 +35,7 @@ testthat::test_that("works with two clevage site", {
 testthat::test_that("works with 6 AA past", {
   testthat::expect_equal(
     mspms::nterm_cleavage("A_BCDEFGHIJKLM_N", "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",6),
-    data.frame(peptide = "A_BCDEFGHIJKLM_N",
+    tibble::tibble(peptide = "A_BCDEFGHIJKLM_N",
                nterm = "XXXXXABCDEFG",
                nterm_cleavage_pos = 1)
   )
@@ -46,7 +46,7 @@ testthat::test_that("works with 6 AA past", {
 testthat::test_that("works with 10 AA past", {
[…]

### 2024-05-02 README.Rmd
cleared all notes except for global binding ones
changed github actions

@@
 <!-- badges: start -->
+[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
 <!-- badges: end -->
@@
 2. *prepare_pd()*: prepares exported files from proteome discoverer.  
-  - *extract_design_matrix_pd()*: Extracts the design matrix from the proteome discoverer file.  
-3. *prepare_spectronaut()*: prepares exported files from spectronaut - note in development, not yet implemented.  
-4. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.
+3. *extract_design_matrix_pd()*: Extracts the design matrix from the proteome discoverer file.  
+4. *prepare_spectronaut()*: prepares exported files from spectronaut - note in development, not yet implemented.  
+5. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.
 
 **Data Proessing/ Normalization**.  
-1. *normalyze()*: Normalizes values and then does a reverse log2 transformation.    
+1. *normalyze()*: Normalizes values.     
 2. *handle_outliers()*: Looks for outliers across replicates. Removes them.       
@@
 3. *plot_time_course()*: Plot peptides over time by condition.  
-4. *plot_cleavage_motif()*: Visualize the sequence specificity of the cleavage sites. Designed to be similar to what is implemented in IceLogo.
+4. *plot_cleavage_motif()*: Visualize the sequence specificity of the cleavage sites. Designed to be equivalent to what is implemented in IceLogo.
 
@@
 
+### Formating Proteome Discoverer File Outputs.
+
+We can prepare proteome discover files for analysis with mspms as follows.
+
+```{r}
+prepared_proteome_discoverer = prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
+```
+
+We can also extract a design matrix from the naming scheme used by proteome discoverer as follows.
+
+```{r}
+ design_matrix = extract_design_matrix_pd("tests/testdata/proteome_discoverer_output.xlsx")
+```
+
 ### Calculating all cleavages
@@
 We might want to calculate all possible cleavages for the peptide library sequences.
-This is useful for downstream analysis, especially when we are looking at the specificity of the cleavage sites via  as this requires a background of all possible cleavages
+This is useful for downstream analysis, especially when we are looking at the specificity of the cleavage sites via plot_cleavage_motif() as this requires a background of all possible cleavages
 
@@
 
-msp-ms uses the normalyzede package to do normalization under the hood.
+mspms uses the NormalyzerDE package to do normalization under the hood.
 
-Now we can normalyze the data.
+We can normalyze the data as follows:  
 ```{r}
@@
 
-We have a lot of missing, or 0 values. For these, we need to impute them so we can do downstream statistics
-Data is imputated by ....
+We have a lot of missing, or 0 values. For these, we need to impute them so we can do downstream statistics.
 
@@
 
-Next we need to join everything with the sequences of the peptide library
+Next we need to join everything with the sequences of the peptide library.  
 ```{r}
@@
 
-### Calcuating clevages.
+### Calcuating cleavages.
 
@@
 
-We can also calculate the log 2 fc. The comparisons here are the same as for the T tests
+We can also calculate the log 2 fold change. The comparisons here are the same as for the T tests.  
 ```{r}
@@
 
-Now that we have this data we can visualize these results easily using the ggplot2 package.
+Now that we have this data we can visualize these results easily using the ggplot2 package as follows.
 
@@
 
-![](www/plot_heatmap_output.png)
+![](plot_heatmap_output.png)
 

### 2024-05-02 README.Rmd
added code coverage

@@
 [![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
+[![Codecov test coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
 <!-- badges: end -->

### 2024-05-21 R/plot_volcano.R
update to man

@@ -24,7 +24,7 @@ plot_volcano = function(log2fc_t_test_data,log2fc_threshold = 3){
     ggplot2::geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
     ggplot2::geom_vline(xintercept = log2fc_threshold, linetype = "dashed",color = "red")+
     ggplot2::geom_vline(xintercept = -log2fc_threshold, linetype = "dashed",color = "red")+
-    ggplot2::labs(x = "Log2 (Time 0/X)",y = "-log10(p value)")+
+    ggplot2::labs(x = "Log2 (Time TX/T0)",y = "-log10(p value)")+
     ggplot2::facet_wrap(~ condition + group2,ncol = ncol)
 
   return(p1)

### 2024-05-21 tests/test_new_functions.Rmd
tests

---
title: "test_random_sample_motif"
output: html_document
date: "2024-05-17"
---

```{r}

library(dplyr)
# Reading in the probabilities of an AA at each position
  probabilities = readxl::read_excel("testdata/motif_sampling_probabilities.xlsx") %>%
    tibble::column_to_rownames("AA") %>%
    t() %>%
    as.data.frame()

  # Switching non na values with the AA supplied
  replace_non_na <- function(x, colname) {
    sapply(x, function(y) ifelse(is.na(y), NA, colname))
  }

  # Defining function to look for columns with zeros
  is.zero = function(x) {
    x == 0
  }

  positions_with_0 = probabilities %>%
    dplyr::select_if(~any(is.zero(.)))

  # switching to zeros
  positions_with_0_mod <- as.data.frame(lapply(names(positions_with_0), function(col)
    replace_non_na(positions_with_0[[col]], col)))
  names(positions_with_0_mod) <- names(positions_with_0)

    # Changing  NAs to . for pattern extraction
  positions_with_0_mod = positions_with_0_mod %>%
    dplyr::mutate(
      dplyr::across(everything(), tidyr::replace_na, ".")
    )

  ### removing the peptides with zeros ###

  zero_removed = tibble::tibble()

  col_names_with_zero = colnames(positions_with_0_mod)

   for(i in 1:ncol(positions_with_0_mod)){
    # Extracting the name of the column
    col_name_with_zero = colnames(positions_with_0_mod)[i]

    logic = paste0(positions_with_0_mod[[i]],collapse = "")
    T_F = !grepl(logic,mspms::all_possible_8mers_from_228_library)
    cleavages = list(mspms::all_possible_8mers_from_228_library[T_F])

    output = tibble::tibble(i = cleavages)

    names(output)[names(output) == 'i'] = col_name_with_zero

    ## Adding to Tibble ##
    zero_removed = tibble::add_column(output,zero_removed)

   }

  zero_removed_vector = zero_removed$P[[1]]

  ### Extracting the Peptides that we would expect to be there ###

  df_modified <- as.data.frame(lapply(names(probabilities), function(col) replace_non_na(probabilities[[col]], col)))
  names(df_modified) <- names(probabilities)

  # Changing  NAs to . for pattern extraction
  df_modified = df_modified %>%
    dplyr::mutate(
      dplyr::across(everything(), tidyr::replace_na, ".")
    ) %>% 
    dplyr::select(-col_names_with_zero)

  # Defining a tibble where the columns have only peptides with the specificed
  # cleavages at the specificed positions

  cleavage_seq_per_AA = tibble::tibble()

  for(i in 1:ncol(df_modified)){

    # Extracting the name of the column
    col_name = colnames(df_modified)[i]

    logic = paste0(df_modified[[i]],collapse = "")
    T_F = grepl(logic,zero_removed_vector)
    cleavages = list(zero_removed_vector[T_F])

    output = tibble::tibble(i = cleavages)

    names(output)[names(output) == 'i'] = col_name

    ## Adding to Tibble ##
    cleavage_seq_per_AA = tibble::add_column(output,cleavage_seq_per_AA,)

  }

trypsin_seqs = c(cleavage_seq_per_AA$R[[1]],cleavage_seq_per_AA$K[[1]])

trypsin_seqs
```

```{r}
mspms::plot_cleavage_motif(c(trypsin_seqs),mspms::all_possible_8mers_from_228_library)

```

```{r}
sum(grepl(".......K",mspms::all_possible_8mers_from_228_library))
sum(grepl(".......K",trypsin_seqs))

sum(grepl("....P...",mspms::all_possible_8mers_from_228_library))
sum(grepl("....P...",trypsin_seqs))

sum(grepl("...K....",mspms::all_possible_8mers_from_228_library))
sum(grepl("...K....",trypsin_seqs))

sum(grepl("...R....",mspms::all_possible_8mers_from_228_library))
sum(grepl("...R....",trypsin_seqs))
```

There is a problem where the motif plotter won't plot residues if they have a count of zero. These may be the most differnt though! Investigating below:

```{r,fig.height = 5, fig.width = 3}
cleavage_seqs = readLines("testdata/in_silico_trypsin_digest_out.csv")
background_universe = mspms::all_possible_8mers_from_228_library
type = "fold_change"

plot_icelogo = function(cleavage_seqs,
                        background_universe = mspms::all_possible_8mers_from_228_library,
                        type = "percent_difference") {
  ## First need to format this such that we have a matrix with the counts at each location of the sequence
  nchar_bg = nchar(background_universe[1])

  # do this for the background first.
  background_universe = tibble::tibble(sequences = background_universe)

  # separate the sequences into individual amino acids
  bg_seq = tidyr::separate(
    background_universe,
    col = 1,
    into = paste0("position", 1:(nchar_bg + 1)),
    sep = "",
    remove = TRUE
  ) %>%
    dplyr::select(-.data$position1)

  # Counting the number in each position
  bg_count = bg_seq %>%
    purrr::map_df(table) %>%
    t()

  #replacing NA with 0
  bg_count[is.na(bg_count)] = 0

  bg_prop = bg_count %>%
    as.data.frame() %>%
    tibble::rownames_to_column("AA") %>%
    dplyr::mutate(dplyr::across(dplyr::where(is.numeric), prop.table)) %>%
    tibble::column_to_rownames("AA") %>%
    as.matrix()

  # Now doing this for the experimental clevages
  nchar_cleav = nchar(cleavage_seqs[[1]])

  # done for the background first
  cleavage_seqs = tibble::tibble(sequences = cleavage_seqs)

  clev_seq = tidyr::separate(
    cleavage_seqs,
    col = 1,
    into = paste0("position", 1:(nchar_cleav + 1)),
    sep = "",
    remove = TRUE
  ) %>%
    dplyr::select(-.data$position1)

  # counting the number of time each AA appears at each position
  clev_count = clev_seq %>%
    purrr::map_df(table) %>%
    t()

  # replacing NA with 0 
  clev_count[is.na(clev_count)] = 0

  clev_prop = clev_count %>%
    as.data.frame() %>%
    tibble::rownames_to_column("AA") %>%
    dplyr::mutate(dplyr::across(dplyr::where(is.numeric), prop.table)) %>%
    tibble::column_to_rownames("AA") %>%
    as.matrix()

  ### calculating the SD ###
  # defining function per Icelogo manual
  standard_deviation = function(proportion_mat, count_mat) {
    (proportion_mat / colSums(count_mat)) ^ (1 / 2)
  }

  # applying to bg data
  bg_sd = standard_deviation(bg_prop, clev_count)

  ### Calculating Z score ###

  # Defining function per icelogo manual
  zscore = function(cleav_prop, bg_prop, bg_sd) {
    (cleav_prop - bg_prop) / bg_sd
  }

  # applying to our data
  zscores = zscore(clev_prop, bg_prop, bg_sd) %>%
    as.data.frame()

  names(zscores) = paste0("P", c((nchar_cleav / 2):1,
                                 paste0(1:(nchar_cleav / 2), "'")))

  #finding the significant zscores
  sig_zscores = zscores %>%
    tibble::rownames_to_column("AA") %>%
    tidyr::pivot_longer(2:length(.), names_to = "position", values_to = "zscore") %>%
    dplyr::filter(abs(zscore) > 1.96) %>%
    dplyr::mutate(aa_position = paste0(.data$AA, ".", .data$position))

  # Calculating percent difference
  pd = (clev_prop * 100) - (bg_prop * 100) %>%
    as.data.frame()

  names(pd) = paste0("P", c((nchar_cleav / 2):1,
                            paste0(1:(nchar_cleav / 2), "'")))

  ### Determining the type of plot to create ###

  if (type == "percent_difference") {
    sig_final = pd %>%
      tibble::rownames_to_column("AA") %>%
      tidyr::pivot_longer(2:length(.), names_to = "position", values_to = "percent_difference") %>%
      dplyr::mutate(aa_position = paste0(.data$AA, ".", .data$position)) %>%
      dplyr::filter(.data$aa_position %in% sig_zscores$aa_position) %>%
      dplyr::select(-.data$aa_position) %>%
      tidyr::pivot_wider(
        names_from = .data$position,
        values_from = .data$percent_difference,
        names_sort = FALSE
      )

    # Dealing with the problem introduced by the pivot wider function
    missing_cols = 9 - ncol(sig_final)
    missing_data = as.data.frame(matrix(nrow = nrow(sig_final), ncol = missing_cols, NA))
    `%!in%` = Negate(`%in%`)
    missing_names = names(pd)[names(pd) %!in% names(sig_final)]
    names(missing_data) = missing_names

    final = sig_final %>%
      dplyr::bind_cols(missing_data) %>%
      dplyr::relocate(paste0("P", c((nchar_cleav / 2):1,
                                    paste0(
                                      1:(nchar_cleav / 2), "'"
                                    )))) %>%
      tibble::column_to_rownames("AA") %>%
      as.matrix()

    # Plotting the icelogo
    p1 = ggseqlogo::ggseqlogo(final,
                              font = "helvetica_light" ,
                              method = 'custom',
                              seq_type = "AA") +
      ggplot2::scale_x_continuous(labels = paste0("P", c((nchar_cleav / 2):1,
                                                         paste0(
                                                           1:(nchar_cleav / 2), "'"
                                                         ))),
                                  breaks = 1:nchar_cleav) +
      ggplot2::ylab("p < 0.05 Percent Difference")

    return(p1)

  } else if (type == "fold_change") {
    fold_change = (clev_prop * 100) / (bg_prop * 100)

    converted_fc = fold_change

    # Here we run into a problem with infinite values. If a value is 0 in the count, but greater than that in the background - we have a problem.
    # To deal with this, ice logo just shows an infinite value as taking up the entire scale.
    # We could mimic that here by setting the columns with infinite values to the max value of the fold change/ the number of things in the columnns

    max_sum = max(purrr::map_df(as.data.frame(converted_fc), sum), na.rm = T)
    final_converted_fc  = data.frame(row.names = rownames(converted_fc))

    for (i in 1:ncol(converted_fc)) {
      if (0 %in% converted_fc[, i]) {
        num_zero = sum(converted_fc[, i] == 0, na.rm = TRUE)

        col = replace(converted_fc[, i],
                      converted_fc[, i] == 0,
                      (max_sum / num_zero * -1))

        final_converted_fc =  cbind(final_converted_fc, col)
      } else{
        final_converted_fc = cbind(final_converted_fc, converted_fc[, i])
      }
    }
    names(final_converted_fc) = paste0("P", c((nchar_cleav / 2):1,
                                              paste0(1:(nchar_cleav / 2), "'")))
    sig_final = final_converted_fc %>%
      tibble::rownames_to_column("AA") %>%
      tidyr::pivot_longer(2:length(.), names_to = "position", values_to = "fold_change") %>%
      dplyr::mutate(aa_position = paste0(.data$AA, ".", .data$position)) %>%
      dplyr::filter(.data$aa_position %in% sig_zscores$aa_position) %>%
      dplyr::select(-.data$aa_position) %>%
      tidyr::pivot_wider(
        names_from = .data$position,
        values_from = .data$fold_change,
        names_sort = FALSE
      )

    # Dealing with the problem introduced by the pivot wider function
    missing_cols = 9 - ncol(sig_final)
    missing_data = as.data.frame(matrix(nrow = nrow(sig_final), ncol = missing_cols, NA))
    `%!in%` = Negate(`%in%`)
    missing_names = names(final_converted_fc)[names(final_converted_fc) %!in% names(sig_final)]

    names(missing_data) = missing_names
    final = sig_final %>%
      dplyr::bind_cols(missing_data) %>%
      dplyr::relocate(paste0("P", c((nchar_cleav / 2):1,
                                    paste0(
                                      1:(nchar_cleav / 2), "'"
                                    ))), ) %>%
      tibble::column_to_rownames("AA") %>%
      as.matrix()

    # Plotting the motif
    p1 = ggseqlogo::ggseqlogo(final,
                              font = "helvetica_light" ,
                              method = 'custom',
                              seq_type = "AA") +
      ggplot2::scale_x_continuous(labels = paste0("P", c((nchar_cleav / 2):1,
                                                         paste0(
                                                           1:(nchar_cleav / 2), "'"
                                                         ))),
                                  breaks = 1:nchar_cleav) +
      ggplot2::ylab("p < 0.05 log2 fold change")

    return(p1)

  } else {
    stop("type must be either 'percent_difference' or 'fold_change'")
  }
}

```

```{r}

readr::write_lines(trypsin_seqs,"testdata/in_silico_trypsin_digest_out.csv")

readr::write_lines(mspms::all_possible_8mers_from_228_library,"testdata/background_out.csv")
```

### 2024-05-21 RMD_template/RMD_template.Rmd
updates to package structure.

---
title: "MSPMS Report"
date: "`r Sys.Date()`"
output:
  html_document:
        theme: cosmo
        toc: true
        toc_depth: 6
        toc_float:
            collapsed: true
            smooth_scroll: true
        code_folding: hide
        highlight: tango
        df_print: paged
params:
  prepared_data: NULL
  design_matrix: NULL
  peptide_library: mspms::peptide_library
  n_residues: 4
---

```{r setup, include=FALSE}
library(dplyr)
ggplot2::theme_set(ggplot2::theme_minimal())
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# Info

This report has been automatically generated by the mspms package. It contains standard results of the mspms analysis pipeline.

R session info is as follows:

```{r}
sessionInfo()
```
# Data Processing

```{r}
# The standard mspms data processing pipeline is used here. 
mspms = mspms::mspms(prepared_data,design_matrix,peptide_library,n_residues)
```

# Overview 
```{r}
mspms::plot_pca(mspms)
```

**Figure 1.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.

```{r,fig.height = 8, fig.width = 10}
mspms::plot_heatmap(mspms)
```

**Figure 2.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.

# Statistics

```{r}
t1 = mspms::mspms_anova(mspms) %>% 
  dplyr::arrange(p.adj)

DT::datatable(t1)
```

**Table 1. ** Here anova results are displayed for each condition (testing to see if there is an effect on time).

```{r}
t2 = log2fc_t_test(mspms) %>% 
  dplyr::arrange(p.adj)

DT::datatable(t2)
```

**Table 2.** Here T-test results are displayed as compared to time 0 for each condition

```{r}
plot_volcano(t2)
```

**Figure 3.** Here the Log2fc and T tests are displayed as volcano plots relative to T0 for each condition.

# Cleavage Motifs

```{r,fig.height = 10, fig.width = 6}
print(mspms::plot_all_icelogos(mspms))
```

**Figure 4.** Here the cleavage motifs enriched in each condition are shown.

### 2024-05-21 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
updates to package structure.

--- 2024-05-21 RMD_template/RMD_template.Rmd
@@
 # The standard mspms data processing pipeline is used here. 
-mspms = mspms::mspms(prepared_data,design_matrix,peptide_library,n_residues)
+mspms = mspms::mspms(prepared_data,design_matrix)
 ```
@@
 
-```{r,fig.height = 8, fig.width = 10}
+```{r}
 mspms::plot_heatmap(mspms)
@@
 
-```{r,fig.height = 10, fig.width = 6}
+```{r}
 print(mspms::plot_all_icelogos(mspms))

### 2024-05-31 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
major changes

@@
 library(dplyr)
-ggplot2::theme_set(ggplot2::theme_minimal())
+ggplot2::theme_set(ggplot2::theme_bw())
 knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
@@
 # The standard mspms data processing pipeline is used here. 
-mspms = mspms::mspms(prepared_data,design_matrix)
+mspms_data = mspms::mspms(prepared_data,design_matrix)
 ```
@@
 ```{r}
-mspms::plot_pca(mspms)
+mspms::plot_pca(mspms_data)
 ```
@@
 
-```{r}
-mspms::plot_heatmap(mspms)
+```{r,fig.height = 8, fig.width = 10}
+mspms::plot_heatmap(mspms_data)
 ```
@@
 ```{r}
-t1 = mspms::mspms_anova(mspms) %>% 
+t1 = mspms::mspms_anova(mspms_data) %>% 
   dplyr::arrange(p.adj)
@@
 ```{r}
-t2 = log2fc_t_test(mspms) %>% 
+t2 = log2fc_t_test(mspms_data) %>% 
   dplyr::arrange(p.adj)
@@
 
-```{r}
-print(mspms::plot_all_icelogos(mspms))
+```{r,fig.height = 10, fig.width = 6}
+print(mspms::plot_all_icelogos(mspms_data))
 ```

### 2024-05-31 R/calc_AA_percent_difference.R

@@ -1,0 +1,23 @@
+#' calc_AA_percent_difference
+#'
+#' calculate the percent difference between a matrix of background proportions and a matrix of experimentally observed proportions.
+#'
+#' @param background_prop_matrix = proportion matrix of aa per position from background cleavage sequences
+#' @param experimental_prop_matrix = a proportion matrix of aa per position from experimental cleavage sequences
+#'
+#' @return a data frame of percent differences
+#' @export
+#' @examples
+#'experimental_prop_matrix = matrix(c(0.2,0.2,0.2,0.2),nrow = 1, dimnames = list("A"))
+#'background_prop_matrix = matrix(c(0.4,0.4,0.4,0.4),nrow = 1, dimnames = list("A"))
+#'calc_AA_percent_difference(background_prop_matrix, experimental_prop_matrix)
+calc_AA_percent_difference = function(background_prop_matrix,
+                                      experimental_prop_matrix){
+
+  # Calculating percent difference
+  pd = (experimental_prop_matrix * 100) - (background_prop_matrix * 100) %>%
+    as.data.frame()
+
+  return(pd)
+
+}

### 2024-05-31 R/calc_bits.R

@@ -1,15 +1,0 @@
-#' calc_bits
-#'
-#'function to calculate the bits of a given proportion as described: https://iomics.ugent.be/icelogoserver/resources/manual.pdf
-#'
-#' @param proportion  a numeric value representing the proportion of a given amino acid
-#'
-#' @return a numeric value representing the bits of the given proportion
-#' @export
-#'
-#' @examples
-#' calc_bits(0.25)
-calc_bits = function(proportion){
-  bits = (proportion * log2(proportion)) * -1
-  return(bits)
-}

### 2024-05-31 R/count_cleavages_per_pos.R

@@ -1,0 +1,52 @@
+#' count_cleavages_per_pos
+#'
+#' count the number of cleavages per position
+#'
+#' @param mspms_data = this is the data that has gone through the mspms pipeline.
+#' @param sig_peptides = this is a list of peptides of interest.
+#'
+#' @return a tibble with the number of cleavages per position
+#' @export
+#'
+#' @examples
+#'# Extracting the significant peptides for DMSO
+#'sig_peptides = mspms::mspms_anova(mspms::mspms_data) %>%
+#'  dplyr::filter(p.adj <= 0.05,
+#'                condition == "MZB") %>%
+#'  dplyr::pull(Peptide)
+#'
+#'# Counting the number of cleavages per position
+#'counts = count_cleavages_per_pos(mspms::mspms_data,sig_peptides)
+#'
+#'#Plotting the results
+#'p1 = counts %>%
+#'  ggplot2::ggplot(ggplot2::aes(x = cleavage_pos,y = n)) +
+#'  ggplot2::geom_point()+
+#'  ggplot2::geom_line()
+#'p1
+#'
+count_cleavages_per_pos = function(mspms_data,sig_peptides){
+
+  positions = 1:13
+
+  count = mspms_data %>%
+    dplyr::select(Peptide,cleavage_pos) %>%
+    dplyr::distinct() %>%
+    dplyr::filter(Peptide %in% sig_peptides) %>%
+    dplyr::group_by(cleavage_pos) %>%
+    dplyr::summarise(n = dplyr::n()) %>%
+    dplyr::ungroup()
+
+
+  missing = positions[!positions %in% count$cleavage_pos]
+
+  missing_t = tibble::tibble(cleavage_pos = missing, n = rep(0,length(missing)))
+
+  out = dplyr::bind_rows(count,missing_t) %>%
+    dplyr::arrange(n)
+
+  return(out)
+
+}
+
+

### 2024-05-31 R/cterm_cleavage.R

@@ -9,7 +9,6 @@
 
 #' @return a tibble with the peptide sequence, cleavage sequences n number of AA on the left and right of the c term cleavage, and the position of the c term cleavage in the library sequence.
 #' @export
-#'
 #' @examples
 #' peptide_sequence = "YWLSTHLAGKR_R"
 #' library_match_sequence = "YWLSTHLAGKRRDW"

### 2024-05-31 R/generate_report.R

@@ -11,7 +11,7 @@
 #' @return a knited .html report of the mspms analysis.
 #' @export
 #'
-#' @examplesIf is.TRUE(FALSE)
+#' @examplesIf isTRUE(FALSE)
 #'
 #' generate_report(prepared_data = mspms::peaks_prepared_data,
 #' design_matrix = mspms::design_matrix,
@@ -23,7 +23,7 @@ generate_report = function(prepared_data,
                            n_residues = 4,
                            outdir = getwd()){
 
-  rmarkdown::render("inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd",
+  rmarkdown::render(system.file("rmarkdown/templates/mspms_report/skeleton/skeleton.RMD",package = "mspms"),
                     params = list(prepared_data = prepared_data,
                                   design_matrix = design_matrix,
                                   peptide_library = peptide_library,

### 2024-05-31 R/log2fc_t_test.R

@@ -2,23 +2,23 @@
 #'
 #' Gives log2fc and statistics from t tests across conditions. Gives differences relative to time 0 within each condition.
 #'
-#' @param prepared_for_stats = this is data that has been prepared using prepare_for_stats().
+#' @param mspms_data = this is data that has been run through the mspms pipeline.
 #'
-#' @return a tibblecontaining log2fc and t test statistics across conditions
+#' @return a tibble containing log2fc and t test statistics across conditions
 #' @export
 #'
 #' @examples
 #'
-#' log2fc_and_t_test = log2fc_t_test(mspms::prepared_for_stats)
+#' log2fc_and_t_test = log2fc_t_test(mspms::mspms_data)
 #'
 #'
-log2fc_t_test = function(prepared_for_stats){
+log2fc_t_test = function(mspms_data){
 
   # Calculating the log2fc
-  log2fc = mspms::mspms_log2fc(prepared_for_stats)
+  log2fc = mspms::mspms_log2fc(mspms_data)
 
   # Calculating the stats
-  stat = mspms::mspms_t_tests(prepared_for_stats)
+  stat = mspms::mspms_t_tests(mspms_data)
 
   #Combining them together
   log2fc_stats = dplyr::inner_join(log2fc,stat,by = c("Peptide","comparison")) %>%

### 2024-05-31 R/log2fct_time.R

@@ -1,0 +1,64 @@
+#' log2fct_time
+#'
+#' preforms log2fc and t test analysis for a user specified condition comparing two specified times.
+#'
+#' @param mspms_data processed mspms data
+#' @param within_condition the condition you would like to compare within
+#' @param ref_time the reference (denominator) time
+#' @param comparison_time the comparison (numerator) time
+#'
+#' @return a tibble
+#' @export
+#'
+#' @examples
+#'
+#'log2t = log2fct_time(mspms::mspms_data,
+#'                     within_condition = "DMSO",
+#'                     ref_time = "0",
+#'                     comparison_time = "60")
+#'
+log2fct_time = function(mspms_data,
+                             within_condition,
+                             ref_time,
+                             comparison_time){
+
+  # Filtering the data set to only include the relevant conditions and time point.
+  f = mspms_data %>%
+    dplyr::filter(time %in% c(ref_time,comparison_time),
+           condition == within_condition)
+
+  # Performing the statistics
+  stat = f %>%
+    dplyr::group_by(Peptide,cleavage_seq) %>%
+    rstatix::t_test(value ~ time,ref.group = ref_time) %>%
+    rstatix::adjust_pvalue(method = "fdr") %>%
+    tibble::as_tibble()
+
+  # Extracting the reference means
+  reference_data = f %>%
+    dplyr::filter(condition == ref_time) %>%
+    dplyr::group_by(.data$Peptide) %>%
+    dplyr::summarise(reference_mean = mean(.data$value,na.rm = TRUE))
+
+  # Extracting the comparison means
+  comparison_data = f %>%
+    dplyr::filter(condition == comparison_time) %>%
+    dplyr::group_by(.data$Peptide) %>%
+    dplyr::summarise(comparison_mean = mean(.data$value,na.rm = TRUE))
+
+  # Calculating the log2fc
+  log2fc = dplyr::inner_join(comparison_data,reference_data,by = c("Peptide")) %>%
+    dplyr::mutate(log2fc = log2(.data$comparison_mean /.data$reference_mean)) %>%
+    dplyr::mutate(comparison = paste0(comparison_time,
+                                      "/",ref_time, " within ", within_condition)) %>%
+    tibble::as_tibble()
+
+
+  # Combining log2fc with statistics
+  out = dplyr::inner_join(log2fc,stat,by = "Peptide")
+
+  return(out)
+
+}
+
+

### 2024-05-31 R/nterm_cleavage.R

@@ -9,7 +9,6 @@
 
 #' @return a tibble with the peptide sequence, cleavage sequences n specified number of AA on the left and right of the n term cleavage, and the position of the n term cleavage in the library sequence.
 #' @export
-#'
 #' @examples
 #'
 #' peptide_sequence = "F_LAHWVGI"

### 2024-05-31 R/polish.R

@@ -7,7 +7,6 @@
 #'
 #' @return a tibble with the cleavage information combined into a single column and rows with no cleavage information or double information removed.
 #' @export
-#'
 #' @examples
 #' polished = polish(mspms::cleavage_added_data)
 polish = function(cleavage_added_data){
@@ -17,6 +16,8 @@ polish = function(cleavage_added_data){
                                                   !is.na(cterm) & is.na(nterm) ~ cterm,
                                                   TRUE ~NA),.after = "cterm_cleavage_pos") %>%
     dplyr::filter(!is.na(.data$cleavage_seq)) %>%
+    dplyr::mutate(cleavage_pos = dplyr::case_when(is.na(cterm_cleavage_pos) ~ nterm_cleavage_pos,
+                                    TRUE ~ cterm_cleavage_pos),.after = "cleavage_seq") %>%
     tibble::as_tibble()
 
 

### 2024-05-31 R/prepare_for_stats.R

@@ -7,7 +7,6 @@
 #'
 #' @return a tibble in long format with the design matrix combined.
 #' @export
-#'
 #' @examples
 #' prepare_for_stats(mspms::cleavage_added_data,mspms::design_matrix)
 prepare_for_stats = function(cleavage_added_data,design_matrix){

### 2024-05-31 tests/testthat/test-calc_AA_fc.R

@@ -1,0 +1,4 @@
+test_that("throws an error without row names", {
+  expect_error(mspms::calc_AA_fc(matrix(c(0.4,0.2),nrow = 1),
+                                 matrix(c(0.2,0.4),nrow = 1)))
+})

### 2024-06-04 R/add_cleavages.R
updates to pass Rcmd check and bioccheck

@@ -53,17 +53,17 @@ add_cleavages <- function(joined_with_library, n_residues = 4) {
   # Building final data frame.
   output <- dplyr::bind_cols(joined_with_library, cleavages) %>%
     dplyr::select(
-      .data$Peptide,
-      .data$library_reference_id,
-      .data$library_match_sequence,
-      .data$library_real_sequence,
-      .data$nterm,
-      .data$nterm_cleavage_pos,
-      .data$cterm,
-      .data$cterm_cleavage_pos,
+      "Peptide",
+      "library_reference_id",
+      "library_match_sequence",
+      "library_real_sequence",
+      "nterm",
+      "nterm_cleavage_pos",
+      "cterm",
+      "cterm_cleavage_pos",
       dplyr::everything(),
-      -.data$peptide,
-      -.data$Peptide_no_cleavage,
+      -"peptide",
+      -"Peptide_no_cleavage",
       -dplyr::any_of("RT")
     ) %>%
     tibble::as_tibble()

### 2024-06-04 R/calc_sig_zscores.R

@@ -14,6 +14,8 @@
 #' rownames(zscores) <- c("A", "B", "C", "D", "E", "F")
 #' calc_sig_zscores(zscores)
 calc_sig_zscores <- function(zscores, pval = 0.05) {
+  # no visible binding for global variable .
+  . <- NULL
   # converting p value to zscore threshold. Divide by two since it is two tailed.
   threshold <- qnorm(p = pval / 2, lower.tail = FALSE)
 
@@ -24,7 +26,7 @@ calc_sig_zscores <- function(zscores, pval = 0.05) {
       names_to = "position",
       values_to = "zscore"
     ) %>%
-    dplyr::filter(abs(zscore) > threshold) %>%
+    dplyr::filter(abs(.data$zscore) > threshold) %>%
     dplyr::mutate(aa_position = paste0(.data$AA, ".", .data$position))
 
   return(sig_zscores)

### 2024-06-04 R/data.R

@@ -13,7 +13,8 @@
 #'   ...
 #' }
 #' @source <mspms processed data originally from PEAKS files found in
-#'  "tests/testdata/protein-peptides-id.csv" and "tests/testdata/protein-peptides-lfq.csv">
+#'  "tests/testdata/protein-peptides-id.csv" and
+#'   "tests/testdata/protein-peptides-lfq.csv">
 "prepared_for_stats"
 
 #' all possible 8mers from the 228 peptide library used by the O’Donoghue

### 2024-06-04 R/log2fct_condition.R

@@ -24,25 +24,25 @@ log2fct_condition <- function(mspms_data,
   # Filtering the data set to include the relevant conditions and time point.
   f <- mspms_data %>%
     dplyr::ungroup() %>%
-    dplyr::filter(condition %in% c(ref_condition, comparison_condition)) %>%
-    dplyr::filter(time == at_time)
+    dplyr::filter(.data$condition %in% c(ref_condition, comparison_condition)) %>%
+    dplyr::filter(.data$time == at_time)
 
   # Performing the statistics
   stat <- f %>%
-    dplyr::group_by(Peptide, cleavage_seq) %>%
+    dplyr::group_by(.data$Peptide, .data$cleavage_seq, .data$cleavage_pos) %>%
     rstatix::t_test(value ~ condition, ref.group = ref_condition) %>%
     rstatix::adjust_pvalue(method = "fdr") %>%
     tibble::as_tibble()
 
   # Extracting the control data
   reference_data <- f %>%
-    dplyr::filter(condition == ref_condition) %>%
+    dplyr::filter(.data$condition == ref_condition) %>%
     dplyr::group_by(.data$Peptide) %>%
     dplyr::summarise(reference_mean = mean(.data$value, na.rm = TRUE))
 
   # Extracting the experimental data
   comparison_data <- f %>%
-    dplyr::filter(condition == comparison_condition) %>%
+    dplyr::filter(.data$condition == comparison_condition) %>%
     dplyr::group_by(.data$Peptide) %>%
     dplyr::summarise(comparison_mean = mean(.data$value, na.rm = TRUE))
 

### 2024-06-04 R/normalyze.R

@@ -33,6 +33,8 @@
 #' )
 #'
 normalyze <- function(prepared_data, design_matrix, outdir = getwd()) {
+  # No visible binding for global variable.
+  . <- NULL
   # Extracting only the data
   dataOnly <- prepared_data[, design_matrix$sample]
 

### 2024-06-04 R/plot_pca.R

@@ -17,8 +17,8 @@ plot_pca <- function(mspms_data, color = "time", shape = "condition") {
   # dealing with no visible binding for global variable ‘.’ NOTE
   . <- NULL
   PCA_data <- mspms_data %>%
-    dplyr::select(.data$sample, .data$Peptide, .data$group, .data$condition, .data$time, .data$value) %>%
-    tidyr::pivot_wider(names_from = .data$Peptide, values_from = .data$value, values_fn = mean) %>%
+    dplyr::select("sample", "Peptide", "group", "condition", "time", "value") %>%
+    tidyr::pivot_wider(names_from = "Peptide", values_from = "value", values_fn = NULL) %>%
     # if a peptide has an na remove it
     dplyr::select_if(~ !any(is.na(.)))
 
@@ -51,8 +51,7 @@ plot_pca <- function(mspms_data, color = "time", shape = "condition") {
     ggplot2::geom_point() +
     ggplot2::stat_ellipse(level = 0.95) +
     ggplot2::xlab(paste0("PCA-1 ", round(Prop_of_var[1] * 100, 2), "%")) +
-    ggplot2::ylab(paste0("PCA-2 ", round(Prop_of_var[2] * 100, 2), "%")) +
-    ggplot2::theme_bw()
+    ggplot2::ylab(paste0("PCA-2 ", round(Prop_of_var[2] * 100, 2), "%"))
 
   return(plot)
 }

### 2024-06-04 README.Rmd
update to man

--- 2024-05-02 README.Rmd
@@
 ---
-output: github_document
+output:
+  github_document: default
+  self_contained: default
 ---
@@
 )
+
+ggplot2::theme_set(
+  ggplot2::theme_minimal()
+)
+
 ```
@@
 [![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
+[![Codecov test coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
+[![BioC status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)
 <!-- badges: end -->
 
-The goal of mspms is provide a concise code-base for the normalization and data processing required to analyze data from the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).
+The goal of mspms is provide a concise code-base for the normalization and 
+data processing required to analyze data from the 
+[Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).
 
-Additionally, we provide a [graphical user interface powered by shiny apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to utilize the method without requiring any R coding knowledge.
+Additionally, we provide a [graphical user interface powered by shiny apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to
+utilize the method without requiring any R coding knowledge.
 
@@
 
-You can install the released version of mspms from github
+You can install the development version of mspms from github.
 
 ```{r,eval = FALSE}
-devtools::install_github("baynec2/mspms")
+devtools::install_github("baynec2/mspms",
+                         ref = "develop")
 ```
+
+## Quickstart
+
+To generate a general report using your own data, run the following code. 
+It requires data that has been prepared for mspms data analysis by a 
+converter function, and a design matrix. For more information see subsequent 
+sections.
+
+```{r,eval = FALSE}
+mspms::generate_report(
+  prepared_data = mspms::peaks_prepared_data,
+  design_matrix = mspms::design_matrix,
+  outdir = "../Desktop/mspms_report"
+)
+```
+
+The above command will generate a .html file containing a general mspms analysis.
+
+There is much more that can be done using the mspms package- see the following 
+sections for more information.
 
@@
 
-There are 4 different types of functions in this package.  Those that are involved in:
+There are 4 different types of functions in this package.
 
-1.
-Making mspms generically useful.
-These functions are focused on making mspms more generally useful to a wider audience outside of the very specific workflow traditionally used in the O’Donoghue lab.
-Allows for the use of other types of upstream proteomic data processing, different peptide libraries, etc.
+1. Making mspms generically useful. These functions are focused on making mspms 
+more generally useful to a wider audience outside of the very specific workflow
+traditionally used in the O’Donoghue lab. Allows for the use of other types of
+upstream proteomic data processing, different peptide libraries, etc.
 
-2. Data processing/ normalization. These functions allow the user to normalize and process the MSP-MS data.
+2. Data processing/ normalization. These functions allow the user to normalize
+and process the MSP-MS data.
 
-3. Statistics. These methods allow the user to perform basic statistics on the normalyzed/processed data.
+3. Statistics. These methods allow the user to perform basic statistics on the 
+normalized/processed data.
 
-4. Data visualization. These functions allow the user to visualize the data in a number of useful way.
+4. Data visualization. These functions allow the user to visualize the data.
 
@@
 2. *prepare_pd()*: prepares exported files from proteome discoverer.  
-3. *extract_design_matrix_pd()*: Extracts the design matrix from the proteome discoverer file.  
-4. *prepare_spectronaut()*: prepares exported files from spectronaut - note in development, not yet implemented.  
-5. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.
+3. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide
+library sequences.
 
-**Data Proessing/ Normalization**.  
+**Data Processing/ Normalization**.  
 1. *normalyze()*: Normalizes values.     
 2. *handle_outliers()*: Looks for outliers across replicates. Removes them.       
-3. *impute()*: Imputes data for missing values (not including NAs introduced by handle_outliers()).    
-4. *join_with_library()*: Joins the normalyzed data with the peptide library sequences.  
-5.
-*add_cleavages()*: Figures out the locations of the detected clevages within the library of peptide sequences used.
-Is it cleaved at the N or C terminus, or both?
-Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.
-6. *polish()*: combines the cleavage info into one column. Discards peptides that were cleaved on both sides or not at all.  
-7. *prepare_for_stats()*: Reshapes the data into a long format and appends the data in the design matrix to make statistical testing easy to perform.
+3. *impute()*: Imputes data for missing values (not including NAs introduced 
+by handle_outliers()).  
+4. *join_with_library()*: Joins the normalized data with the peptide library 
+sequences.  
+5. *add_cleavages()*: Figures out the locations of the detected cleavages within
+the library of peptide sequences used. Is it cleaved at the N or C terminus, 
+or both? Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.      
+6. *polish()*: combines the cleavage info into one column. Discards peptides
+that were cleaved on both sides or not at all.  
+7. *prepare_for_stats()*: Reshapes the data into a long format and appends the 
+data in the design matrix to make statistical testing easy to perform.  
+8 *mspms():* combines all of the above functions into one convenient function.
 
-**Statistics**.   
-1. *log2fc_t_test()*: Calculates the fold change and the p/q value across experimental conditions using T-tests.      
-2. *mspms_anova()*: Conducts ANOVA tests to determine if there are any peptides that are significantly different across all time points.
+**Calculations/ Statistics**.   
+1. *mspms_log2fc()*: Calculates the log2 fold change across experimental 
+conditions relative to T0.  
+2.*mspms_t_tests()*: Conducts T-tests to determine if there are any peptides 
+that are significantly different across experimental conditions relative to time 
+0.  
+3. *log2fc_t_test()*: Combines results from mspms_log2fc() and mspms_t_tests() 
+into one convenient tibble.   
+4. *log2fct_condition()*: Calculates the log2 fold change and t tests for the 
+specified condition at a user specified time.    
+5.*log2fct_time()*: Calculates the log2 fold change and t tests between the 
+specified times within a user specified time.   
+6. *mspms_anova()*: Conducts ANOVA tests to determine if there are any peptides 
+that are significantly different across all time points.   
+7. *count_cleavages_per_pos():* Counts the number of cleavages at each position 
+of the library. Useful for determining endoprotease vs exoprotease activity.
 
 **Data Visualization**.   
-1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an interactive heatmap to visualize overall patterns in the data.    
+1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an
+interactive heatmap to visualize overall patterns in the data.    
 2. *plot_pca()*: PCA analysis to visualize the data in a 2D space.  
 3. *plot_time_course()*: Plot peptides over time by condition.  
-4. *plot_cleavage_motif()*: Visualize the sequence specificity of the cleavage sites. Designed to be equivalent to what is implemented in IceLogo.
+4. *plot_cleavages_per_pos()*: Plot the number of cleavages at each position 
+in the library.Useful for determining endoprotease vs exoprotease activity.
+
+**Icelogo**.
+
+1. *calc_AA_count_of_motif()*: calculates the count of each amino acid at each
+position in a vector of motifs.  
+2. *calc_AA_prop_of_motif()*: calculates the proportion of each amino acid at 
+each position from the output of calc_AA_count_of_motif().  
+3. *calc_AA_motif_zscore()*: calculates z scores for each amino acid at each 
+position in a vector of motifs.  
+4. *calc_AA_percent_difference()*: calculates the percentage difference of 
+5. *prepare_sig_p_dif()*: prepares the data for plotting percent difference 
+icelogos. Removes AAs at positions that are not significantly different.  
+6. *prepare_fc()*: prepares the data for plotting fold change icelogos.
+Removes AAs at positions that are not significantly different.  
+7. *plot_pd_icelogo()*: plots a percent difference icelogo.  
+8. *plot_fc_icelogo()*: plots a fold change ice logo.  
+9. *plot_icelogo()*:  wraps plot_pd_icelogo() and plot_fc_icelogo() into 
+one function.  
+10. *plot_all_icelogos()*: convenience function to plot all icelogos relative to 
+time 0 from an experiment.    
+**Reports**. 
+1. *generate_report()*: Generates a simple analysis of the mspms data that 
+should be generically applicable to all mspms experiments.
 
@@
 
-All analysis performed by this package is downstream of the data generated by the PEAKS software. The first step is to combine the two files generated by PEAKS.
+All analysis performed by this package is downstream of the data generated by 
+the PEAKS software. The first step is to combine the two files generated by
+PEAKS.
 
-The files coming from peaks should be generated according to the instructions found [here](www/PeaksDataAnalysis_howto.pdf).
+The files coming from peaks should be generated according to the instructions 
+found [here](www/PeaksDataAnalysis_howto.pdf).
 
@@
 ### Loading the files ###
-lfq_filename = "tests/testdata/protein-peptides-lfq.csv"
-#file "protein-peptides.csv" exported from PEAKS identification
-id_filename = "tests/testdata/protein-peptides-id.csv"
+lfq_filename <- "tests/testdata/protein-peptides-lfq.csv"
+# file "protein-peptides.csv" exported from PEAKS identification
+id_filename <- "tests/testdata/protein-peptides-id.csv"
 
 # Prepare the data for normalyzer analysis
-peaks_prepared_data = prepare_peaks(lfq_filename,id_filename)
-
-# saving to data folder in package for tests. Ignore when using on your own data
-usethis::use_data(peaks_prepared_data,overwrite = TRUE)
+peaks_prepared_data <- prepare_peaks(lfq_filename, id_filename)
 ```
@@
 ```{r}
-prepared_proteome_discoverer = prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
-```
-
-We can also extract a design matrix from the naming scheme used by proteome discoverer as follows.
-
-```{r}
- design_matrix = extract_design_matrix_pd("tests/testdata/proteome_discoverer_output.xlsx")
+prepared_proteome_discoverer <- prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
 ```
@@
 
-We might want to calculate all possible cleavages for the peptide library sequences.
-This is useful for downstream analysis, especially when we are looking at the specificity of the cleavage sites via plot_cleavage_motif() as this requires a background of all possible cleavages
+We might want to calculate all possible cleavages for the peptide library 
+sequences. This is useful for downstream analysis, especially when we are 
+looking at the specificity of the cleavage sites via plot_cleavage_motif() as
+this requires a background of all possible cleavages
 
-We can do this by specifying the number of amino acids after the cleavage site that we are interested in. First lets try 4, which is the default.
+We can do this by specifying the number of amino acids after the cleavage site 
+that we are interested in. First lets try 4, which is the default.
 
 ```{r}
-all_peptide_sequences = mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
-                        n_AA_after_cleavage=4)
+all_peptide_sequences <- mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
+  n_AA_after_cleavage = 4
+)
 head(all_peptide_sequences)
-
 ```
@@
 ```{r}
-all_peptide_sequences = mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
-                        n_AA_after_cleavage=5)
+all_peptide_sequences <- mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
+  n_AA_after_cleavage = 5
+)
 head(all_peptide_sequences)
@@
 ```{r}
-design_matrix = readr::read_csv("tests/testdata/design_matrix.csv")
+design_matrix <- readr::read_csv("tests/testdata/design_matrix.csv")
 
 head(design_matrix)
-
 ```
@@
 ```{r}
-normalyzed_data = normalyze(peaks_prepared_data,design_matrix)
+normalyzed_data <- normalyze(peaks_prepared_data, design_matrix)
 
-# saving to data folder in package for tests. Ignore when using on your own data
-usethis::use_data(normalyzed_data,overwrite = TRUE)
 ```
@@
 
-Here we use a dixon test from the outliers package to detect outliers from each of our replicates.
-We need to know what samples are part of which groups, so we need to specify the design matrix here too.
-Make sure that the column header names are "sample" and "group" just like before.
+Here we use a Dixon test from the outliers package to detect outliers from each 
+of our replicates. We need to know what samples are part of which groups, so we 
+need to specify the design matrix here too. Make sure that the column header 
+names are "sample" and "group" just like before.
 
 ```{r}
-design_matrix = readr::read_csv("tests/testdata/design_matrix.csv")
-outliers = handle_outliers(normalyzed_data,design_matrix)
+design_matrix <- readr::read_csv("tests/testdata/design_matrix.csv")
+outliers <- handle_outliers(normalyzed_data, design_matrix)
 
-# saving to data folder in package for tests. Ignore when using on your own data
-usethis::use_data(outliers,overwrite = TRUE)
 ```
@@
 
-We have a lot of missing, or 0 values. For these, we need to impute them so we can do downstream statistics.
+We have a lot of missing, or 0 values. For these, we need to impute them so we 
+can do downstream statistics.
 
 ```{r}
-imputed = impute(outliers)
+imputed <- mspms::impute(outliers)
 
-# saving to data folder in package for tests. Ignore when using on your own data
-usethis::use_data(imputed,overwrite = TRUE)
 ```
@@
 ```{r}
-joined_with_library = join_with_library(imputed)
-
-# saving to data folder in package for tests. Ignore when using on your own data
-usethis::use_data(joined_with_library,overwrite = TRUE)
+joined_with_library <- mspms::join_with_library(imputed)
 ```
@@
 
-Sequences are presented as the user specified number of amino acids on both sides of a cleavage.
-The default is 4, but there is interest in looking at motifs that are farther away from the cut site.
-Note that X indicates that there was nothing on that side in the library because it was cleaved close to the edge.
+Sequences are presented as the user specified number of amino acids on 
+both sides of a cleavage. The default is 4, but there is interest in looking at
+motifs that are farther away from the cut site. Note that X indicates that there
+was nothing at that position because it is over the end.
 
 ```{r}
-cleavage_added_data = add_cleavages(joined_with_library,n_residues = 4)
+cleavage_added_data <- mspms::add_cleavages(joined_with_library, n_residues = 4)
 
 head(cleavage_added_data)
+```
 
-# saving to data folder in package for tests. Ignore when using on your own data
-usethis::use_data(cleavage_added_data,overwrite = TRUE)
+### Adding metadata back
+
+We need to add information about our samples back to perform downstream 
+statistics on the data. We accomplish this with the prepare for stats file.
+
+```{r}
+prepared_for_stats <- mspms::prepare_for_stats(cleavage_added_data,
+                                               design_matrix)
 ```
@@
 
-Someimes there is a need to polish the data a bit for downstream analysis.
-This function does that by removing combining the cterm and nterm cleavage information into one column while removing any rows that don't have any cleavage information or have cleavage information on the cterm and nterm.
+This function polishes the data by combining the cterm and nterm cleavage 
+information into one column while removing any rows that don't have any 
+cleavage information or have cleavage information on the cterm and nterm.
 
 ```{r}
-polished_data = polish(cleavage_added_data)
+polished_data <- mspms::polish(cleavage_added_data)
+```
+
+### Complete workflow
+
+There are a lot of steps to get to the final normalized data we listed above. 
+We have combined all of these steps into one function, mspms() for convenience.
+This offers a nice balance between allowing the user the option to use 
+individual steps and providing a quick way to get to the final data.
+
+```{r}
+mspms_data <- mspms::mspms(peaks_prepared_data, design_matrix)
 ```
@@
 
-mspms provides a number of convenience functions to conduct statistics on the data.
+mspms provides a number of convenience functions to conduct statistics
+on the data.
 
-First, we need to prepare the data for stats.
-This involves reshaping the data into the long format so it is easy to conduct statistics on and then appending the data in the design matrix that allows us to determine what sample contains what conditions/time and conduct the appropriate statistics.
+First, we need to prepare the data for stats. This involves reshaping the data 
+into the long format so it is easy to conduct statistics on and then appending 
+the data in the design matrix that allows us to determine what sample contains 
+what conditions/time and conduct the appropriate statistics.
 
-```{r}
-prepared_for_stats = prepare_for_stats(cleavage_added_data,design_matrix)
-
-usethis::use_data(prepared_for_stats,overwrite = TRUE)
-```
-
-Noe we can conduct the statistics.
+Now we can conduct the statistics:
 
@@
 
-Ttests are performed within each condition and compared to time 0.
+T tests are performed within each condition and compared to time 0.
 
-For example, for an experiment where there are two conditions, DMSO and MZB as well as 3 time points, 0, 1, and 2, the t tests would be as follows:
+For example, for an experiment where there are two conditions, 
+DMSO and MZB as well as 3 time points, 0, 1, and 2, the t tests would be
+as follows:
 
@@
 # Perform T test
-t_test_stats = mspms::mspms_t_tests(prepared_for_stats)
+t_test_stats <- mspms::mspms_t_tests(mspms_data)
 ```
@@
 
-We can also calculate the log 2 fold change. The comparisons here are the same as for the T tests.  
+We can also calculate the log 2 fold change. 
+The comparisons here are the same as for the T tests.  
 ```{r}
-# calclulate log2fc
-log2fc = mspms::mspms_log2fc(prepared_for_stats)
+# calculate log2fc
+log2fc <- mspms::mspms_log2fc(mspms_data)
 
@@
 
-Sometimes (such as when you want to make volcano plots) it is useful to look at the log2fc and the t test statistics at the same time. This function does that.
+Oftentimes (such as when you want to make volcano plots) it is most useful to
+look at the log2fc and the t test statistics at the same time.
 
 ```{r}
-log2fc_t_test = mspms::log2fc_t_test(prepared_for_stats)
+log2fc_t_test <- mspms::log2fc_t_test(mspms_data)
 
@@
 
-Now that we have this data we can visualize these results easily using the ggplot2 package as follows.
+### log2fct_condition
+
+You may wish to compare two conditions at a specific time point. 
+This function allows this
 
 ```{r}
-library(ggplot2)
+condition_t = mspms::log2fct_condition(mspms_data,
+                                       ref_condition = "DMSO",
+                                       comparison_condition = "MZB",
+                                       at_time = 60)
 
-p1 = log2fc_t_test %>% 
-  ggplot(aes(x = log2fc,y = -log10(p.adj)))+
-  geom_point(size =0.5)+
-  geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
-  geom_vline(xintercept = 3, linetype = "dashed",color = "red")+
-  geom_vline(xintercept = -3, linetype = "dashed",color = "red")+
-  theme_minimal()+
-  labs(x = "Log2 Fold Change",y = "-log10(p value)")+
-  facet_wrap(~comparison,scales = "free")
+head(condition_t)
 
-p1
+```
+
+### log2fct_time
+
+You may wish to compare two time points within a specific condition. 
+This function allows
+
+```{r}
+time_t = mspms::log2fct_time(mspms_data,
+                             within_condition = "DMSO",
+                             ref_time = 0,
+                             comparison_time = 60)
+
+head(time_t)
 ```
@@
 
-We also might want to perfom an anova. Here, we have it set up to show the effect of time within each condition.
+We also might want to perform an anova. Here, we have it set up to show the 
+effect of time within each condition.
 
-For example, for an experiment where there are two conditions, DMSO and MZB as well as 3 time points, 0, 1, and 2, the anova would statistics for the effect of time for each peptide within DMSO or MZB.
+For example, for an experiment where there are two conditions, DMSO and MZB as
+well as 3 time points, 0, 1, and 2, the anova would statistics for the effect of
+time for each peptide within DMSO or MZB.
 
@@
 # Doing ANOVA
-anova_stats = mspms::mspms_anova(prepared_for_stats)
+anova_stats <- mspms::mspms_anova(mspms_data)
 
@@
 
+### volcano plots
+
+```{r}
+volcano_plot <- mspms::plot_volcano(log2fc_t_test)
+
+volcano_plot
+```
+
 ### Plotting sequence specificity motif
 
-Here use an approach similar to what is implemented in ICELogo to visualize the sequence specificity of the cleavage sites.
-This was used as reference to build the code: https://iomics.ugent.be/icelogoserver/resources/manual.pdf.
+Here use an approach similar to what is implemented in ICELOGO to visualize the
+sequence specificity of the cleavage sites. This was used as reference to build 
+the code: https://iomics.ugent.be/icelogoserver/resources/manual.pdf.
+
+Below we can see this applied to see what cleavage sequences are enriched among
+the peptides significantly different between time 0 and time 60 
+within DMSO.
 
 ```{r}
-cleavage_seqs = mspms::prepared_for_stats %>%
-  mspms::polish() %>% 
-  filter(condition == "DMSO",time == 240) %>%
-  pull(cleavage_seq)
+cleavage_seqs <- time_t %>%
+  dplyr::filter(log2fc > 3, p.adj <= 0.05) %>%
+  dplyr::pull(cleavage_seq)
 
-background_universe = mspms::all_possible_8mers_from_228_library
+background_universe <- mspms::all_possible_8mers_from_228_library
 
-mspms::plot_cleavage_motif(cleavage_seqs,background_universe)
+mspms::plot_icelogo(cleavage_seqs, background_universe)
 ```
 
-### PCA
-
-We can generate a PCA plot to visualize the data. Here, the colors show the different time points while the shape shows the different conditions.
+We could also look at the fold change instead of the percent difference
 
 ```{r}
-mspms::plot_pca(prepared_for_stats)
+mspms::plot_icelogo(cleavage_seqs, background_universe, type = "fold change")
+```
+
+We also provide a convenience function for looking at all icelogos relative to 
+time 0 within the experiment.
+
+This function includes cleavages that were significant at any time point once in 
+the experimental set of the ICELOGO.
+
+```{r}
+mspms::plot_all_icelogos(mspms_data)
+```
+
+### PCA.
+
+We can generate a PCA plot to visualize the data. 
+Here, the colors show the different time points while the shape shows 
+the different conditions.
+
+```{r}
+mspms::plot_pca(mspms_data)
 ```
@@
 
-We can also generate an interactive heatmap with the hierchical clustering results.
+We can also generate an interactive heatmap with the hierarchical clustering 
+results.
 
-The format of this readme does not allow for interactivity, so a static picture of the output is shown instead.
+The format of this readme does not allow for interactivity, so a static picture 
+of the output is shown instead.
 
 ```{r,eval = FALSE}
-
-mspms::plot_heatmap(prepared_for_stats)
-
+mspms::plot_heatmap(mspms_data)
 ```
@@
 
-We also provide a function for plotting the mean intensity and standard deviation over time for each peptide in the data set by conditions. Facets show peptide, color shows the condition.
+We also provide a function for plotting the mean intensity and standard 
+deviation over time for each peptide in the data set by conditions.
+Facets show peptide, color shows the condition.
 
-This is best used in combination with the ANOVA function. Below, we will use anova to calculate the peptides where we see an effect of time (in the DMSO group), and then plot these.
+This can be used in combination with the ANOVA function.
+Below, we will use anova to calculate the peptides where we see an 
+effect of time (in the DMSO group), and then plot these.
 
@@
 
-top10sig = anova_stats %>% 
-  arrange(p.adj) %>%
-  pull(Peptide) %>% 
+top10sig <- anova_stats %>%
+  dplyr::arrange(p.adj) %>%
+  dplyr::pull(Peptide) %>%
   head(10)
 
-p1 = prepared_for_stats %>%
+p1 <- mspms_data %>%
   dplyr::filter(Peptide %in% top10sig) %>%
@@
 ```
+
+## Calculating the number of cleavages at each position in the library
+
+We can also calculate the number of cleavages at each position of the library.
+Here we will do this for peptides significantly different relative to time 0 for
+DMSO and MZB.
+
+```{r}
+# Performing T tests
+t_tests = log2fc_t_test(mspms_data) %>% 
+  dplyr::filter(p.adj < 0.05,
+                log2fc > 3)
+
+count = mspms::count_cleavages_per_pos(t_tests)
+
+p1 = mspms:::plot_cleavages_per_pos(count)
+
+p1
+```

### 2024-06-07 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
Changed github actions to work in develop branch and polished some of the plotting functions.

@@
 **Figure 4.** Here the cleavage motifs enriched in each condition are shown. 
+Enriched cleavage motifs are defined as having a p.adj <= 0.05 and a log2fc >3
+relative to time 0.
+
+# Position Specificity
+
+```{r}
+sig = t2 %>% 
+  filter(p.adj <= 0.05, log2fc >3)
+
+d = count_cleavages_per_pos(sig)
+
+plot_cleavages_per_pos(d)
+```
+
+**Figure 5.** Here the count of significant cleavages at each position of the 
+peptide library are shown for each condition (facet) per time point (color). 
+Significant cleavages are defined as having a p.adj <= 0.05 and a log2fc > 3 
+relative to time 0.

### 2024-06-07 R/plot_cleavages_per_pos.R

@@ -4,6 +4,11 @@
 #'
 #' @param count_of_cleavages = tibble with number of cleavages per position, for
 #' each condition and time point
+#' @param facets = condition, grid, or wrap If grid facets for each condition and 
+#' time are shown even if data does not exist. 
+#' Wrap on the other hand is condensed. None shows plots overlayed by time with
+#' different colors.
+#' 
 #'
 #' @return a ggplot2 object
 #' @export
@@ -14,17 +19,36 @@
 #'                   cleavage_pos = 1:13,
 #'                   n = 1:13)
 #' plot_cleavages_per_pos(c)
-plot_cleavages_per_pos = function(count_of_cleavages){
-
+plot_cleavages_per_pos = function(count_of_cleavages,
+                                  facets = "condition"){
   p1 = ggplot2::ggplot(count_of_cleavages,
                        ggplot2::aes_string(x = "cleavage_pos",
                                            y = "n")) +
     ggplot2::geom_point() +
     ggplot2::geom_line()+
-    ggplot2::facet_wrap(~condition + time, scales = "free_y") +
     ggplot2::scale_x_continuous(breaks = seq(0, 13))
 
-  return(p1)
+  if(facets == "grid"){
+    p2 = p1 + ggplot2::facet_grid(rows = vars(condition),
+                                  cols = vars(time)) 
+  } else if (facets == "wrap"){
+    p2 = p1 + ggplot2::facet_wrap(~condition + time,
+                                  ncol = ncol,
+                                  scales = "free_y") 
+  } else if(facets == "condition"){
+    p2 = ggplot2::ggplot(count_of_cleavages,
+                         ggplot2::aes_string(x = "cleavage_pos",
+                                             y = "n",
+                                             color = "time"))+
+      ggplot2::geom_point()+
+      ggplot2::geom_line()+
+      ggplot2::scale_x_continuous(breaks = seq(0, 13))+
+      ggplot2::facet_wrap(~condition)
+  } else {
+    stop("facets must be either condition,grid or wrap")
+  }
+  return(p2)
 }
+
 
 

### 2024-06-07 R/plot_icelogo.R

@@ -20,6 +20,8 @@ plot_icelogo <- function(cleavage_seqs,
                          background_universe = mspms::all_possible_8mers_from_228_library,
                          pval = 0.05,
                          type = "percent_difference") {
+  # calculate n 
+  n = length(cleavage_seqs)
   # calculation proportions of background
   background_counts <- mspms::calc_AA_count_of_motif(background_universe)
   background_proportions <- mspms::calc_AA_prop_of_motif(background_counts)
@@ -45,7 +47,8 @@ plot_icelogo <- function(cleavage_seqs,
       background_prop_matrix = background_proportions
     )
     final_pd <- mspms::prepare_sig_p_dif(pd, sig_zscores)
-    plot <- mspms::plot_pd_icelogo(final_pd)
+    plot <- mspms::plot_pd_icelogo(final_pd)+
+      ggplot2::labs(subtitle = paste0("n = ",n))
     return(plot)
   } else if (type == "fold change") {
     fc <- mspms::calc_AA_fc(
@@ -53,7 +56,8 @@ plot_icelogo <- function(cleavage_seqs,
       background_prop_matrix = background_proportions
     )
     final_fc <- mspms::prepare_fc(fc, sig_zscores)
-    plot <- mspms::plot_fc_icelogo(final_fc)
+    plot <- mspms::plot_fc_icelogo(final_fc) +
+      ggplot2::labs(subtitle = paste0("n = ",n))
     return(plot)
   } else {
     stop("type must be either 'percent_difference' or 'fold change'")

### 2024-06-17 README.Rmd
change install to main
Changed github actions to work in develop branch and polished some of the plotting functions.
README files
updating README to include new functions.

@@
 ---
-output:
-  github_document: default
-  self_contained: default
+output: github_document
+
 ---
@@
   collapse = TRUE,
-  comment = "#>",
-  cache = FALSE
+  comment = "#>"
 )
@@
 
-Additionally, we provide a [graphical user interface powered by shiny apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to
-utilize the method without requiring any R coding knowledge.
+Additionally, we provide a 
+[graphical user interface powered by shiny apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) 
+that allows for a user to utilize the method without requiring any R coding
+knowledge.
 
@@
 
-1. Making mspms generically useful. These functions are focused on making mspms 
+1. Pre-processing data. These functions are focused on making mspms 
 more generally useful to a wider audience outside of the very specific workflow
@@
 
-**Making mspms generically useful**.   
-1. *prepare_peaks()*:Takes two input files from PEAKS and combines them. 
+**Preprocessing data**.   
+1. *prepare_peaks()*:Takes two exported files from PEAKS software, 
+combines them, and prepares for processing by the mspms R package. 
 2. *prepare_pd()*: prepares exported files from proteome discoverer.  
@@
 of the library. Useful for determining endoprotease vs exoprotease activity.
+8. *count_cleavages_per_pos()*: Counts the number of cleavages at the position
+of the initial peptide detected peptides were cleaved from.
+
+**Icelogo calculations**
+
+1. *calc_AA_count_of_motif()*: calculates the count of each amino acid at each
+position in a vector of motifs.  
+2. *calc_AA_prop_of_motif()*: calculates the proportion of each amino acid at 
+each position from the output of calc_AA_count_of_motif().  
+3. *calc_AA_motif_zscore()*: calculates z scores for each amino acid at each 
+position in a vector of motifs.  
+4. *calc_sig_zscore()*: given a matrix of zscores, determines which are significant given a user defined p value. 
+5. *calc_AA_percent_difference()*: calculates the percentage difference of 
+6. *calc_AA_fold_change()*: calculates the fold change of a matrix containing 
+the proportions of amino acids at each position relative to a background matrix. 
+7. *prepare_sig_p_dif()*: prepares the data for plotting percent difference 
+icelogos. Removes AAs at positions that are not significantly different.  
+8. *prepare_pd()*:prepares a matrix with significant amino acid percent 
+differences by position for plotting.
+9. *prepare_fc()*: prepares the data for plotting fold change icelogos.
+10. *prepare_icelogo_data()*: uses all the functions described above to 
+prepare the data for plotting icelogos. 
+11. *extract_re()*: extracts regular expressions specifying amino acids above a
+defined threshold cutoff for each position of a cleavage motif from an icelogo 
+matrix.
 
@@
 in the library.Useful for determining endoprotease vs exoprotease activity.
+5.*plot_icelogo()*: Plot an icelogo given an icelogo matrix. 
+6.*plot_all_icelogos()*: generates icelogos representing enriched sequences 
+relative to time 0 for each conditions. This function compares cleavage 
+sequences corresponding to peptides that are significant 
+(p.adj < 0.05 and log2 fold change > 3) to a background of all possible
+sequences present in the initial peptide library.
 
-**Icelogo**.
-
-1. *calc_AA_count_of_motif()*: calculates the count of each amino acid at each
-position in a vector of motifs.  
-2. *calc_AA_prop_of_motif()*: calculates the proportion of each amino acid at 
-each position from the output of calc_AA_count_of_motif().  
-3. *calc_AA_motif_zscore()*: calculates z scores for each amino acid at each 
-position in a vector of motifs.  
-4. *calc_AA_percent_difference()*: calculates the percentage difference of 
-5. *prepare_sig_p_dif()*: prepares the data for plotting percent difference 
-icelogos. Removes AAs at positions that are not significantly different.  
-6. *prepare_fc()*: prepares the data for plotting fold change icelogos.
-Removes AAs at positions that are not significantly different.  
-7. *plot_pd_icelogo()*: plots a percent difference icelogo.  
-8. *plot_fc_icelogo()*: plots a fold change ice logo.  
-9. *plot_icelogo()*:  wraps plot_pd_icelogo() and plot_fc_icelogo() into 
-one function.  
-10. *plot_all_icelogos()*: convenience function to plot all icelogos relative to 
-time 0 from an experiment.    
 **Reports**. 
@@
 
-## Making Generically Usefull.
+## Data Pre-processing
 
@@
 
-All analysis performed by this package is downstream of the data generated by 
-the PEAKS software. The first step is to combine the two files generated by
-PEAKS.
+Analysis downstream of data exported from PEAKS software is supported.
 
-The files coming from peaks should be generated according to the instructions 
-found [here](www/PeaksDataAnalysis_howto.pdf).
+exported files are generated in the O'donoghue lab as follows:
+
+1. Create a New Project in PEAKS.  
+2. Select data to add to PEAKS project.  
+3. Add data, rename samples, set correct parameters. 
+  • Highlight all the added data.  
+  • Select Create new sample for each file. 
+  • Select appropriate instrument and fragmentation method. 
+  •  Set Enzyme to “None”. 
+  • Rename samples according to enzymes, time points, and replicates. 
+      • Can also ensure appropriate data is selected for these   
+  • Select “Data Refinement” to continue. 
+4. Data Refinement. 
+  • On the top right, select MSP-MS as the predefined parameters. 
+  •  Parameters shown are what should be used for MSP-MS data analysis. 
+  •  Select “Identification” to continue. 
+5. Peptide Identification. 
+  • On the top right, select MSP-MS as the predefined parameters. 
+  • Parameters shown are what should be used for MSP-MS data analysis. 
+  • Identification -> Select Database -> TPD_237. 
+  • Identification -> no PTMs. 
+  • Can remove the PTMs by highlighting them and selecting Remove  
+6. Label Free Quantification. 
+  • Make sure Label Free is selected. 
+  • Group samples -> add quadruplicates of samples to new group.
 
@@
 ```{r}
-prepared_proteome_discoverer <- prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
+prepared_proteome_discoverer <-
+  prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
 ```
@@
 ```{r}
-all_peptide_sequences <- mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
+all_peptide_sequences <- mspms::calculate_all_cleavages(
+  mspms::peptide_library$library_real_sequence,
   n_AA_after_cleavage = 4
@@
 ```{r}
-all_peptide_sequences <- mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
+all_peptide_sequences <- mspms::calculate_all_cleavages(
+  mspms::peptide_library$library_real_sequence,
   n_AA_after_cleavage = 5
@@
 
-We could also look at the fold change instead of the percent difference
-
+We could also look at the fold change instead of the percent difference.
 ```{r}
-mspms::plot_icelogo(cleavage_seqs, background_universe, type = "fold change")
+mspms::plot_icelogo(cleavage_seqs, background_universe, type = "fold_change")
 ```
@@
 This function includes cleavages that were significant at any time point once in 
-the experimental set of the ICELOGO.
+the experimental set of the ICELOGO. Below we show this applied to create
+percent difference and fold change icelogos.
 
 ```{r}
-mspms::plot_all_icelogos(mspms_data)
+plot_all_icelogos(mspms_data)
+
+plot_all_icelogos(mspms_data, type = "fold_change")
 ```
+
+### Extracting regular expressions from icelogos
+
+It might be useful for the user to extract the information contained in the 
+Icelogo and use it to filter their data for peptides that are predicted 
+to be well cleaved, or not cleaved according to the icelogo plot.
+
+For example, we can get the amino acids that are enriched at each position of 
+the icelogo using the code below.
+
+```{r}
+
+# Getting data for Icelog 
+cleavage_seqs <- time_t %>%
+  dplyr::filter(log2fc > 3, p.adj <= 0.05) %>%
+  dplyr::pull(cleavage_seq)
+
+background_universe <- mspms::all_possible_8mers_from_228_library
+
+# Plotting icelogo
+mspms::plot_icelogo(cleavage_seqs, background_universe)
+
+# corresponding icelogo data to the plot
+icelogo_matrix = mspms::prepare_icelogo_data(cleavage_seqs, background_universe)
+
+# Extracting regular expressions
+RE_pos = mspms::extract_re(icelogo_matrix)
+
+RE_pos
+```
+
+We can also do this for amino acids that are negatively enriched.
+
+```{r}
+RE_neg = mspms::extract_re(icelogo_matrix,type = "negative")
+
+RE_neg
+```
+
+Once we have these regular expressions, we can use it to filter our data as
+follows:
+
+```{r}
+pos_enriched = mspms_data %>% 
+  dplyr::filter(grepl(RE_pos, cleavage_seq))
+```
+
+In this case, none of the peptides in the original data set match all the rules
+in our library (peptide library does not contain all possible combinations, so
+none of the peptides was a match to the combination of rules at each position)
+
+We could then filter to only include peptides that are greater than a certain
+threshold percent difference. We can then visualize these results easily with 
+the plot_time_course function described further below. Here we will only 
+consider AAs with a percent difference greater than 10 at each position.
+
+```{r}
+# Extracting regular expressions 
+RE_pos = mspms::extract_re(icelogo_matrix,threshold = 10)
+
+RE_pos
+
+# Filtering data for peptides that have cleavage sequences matching regular 
+# expressions
+f = mspms_data %>% 
+  dplyr::filter(grepl(RE_pos, cleavage_seq))
+
+#visualizing these peptides that match over time.
+plot_time_course(f)
+```
+
+Here we see that these peptides were cleaved over time in DMSO, as would be 
+expected based on the icelogo plot.
 
@@
 
-![](plot_heatmap_output.png)
+![](README_files/plot_heatmap_output.png)
 

### 2024-07-03 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
added download buttons

@@
 mspms_data = mspms::mspms(prepared_data,design_matrix)
+
+mspms_data %>%
+  downloadthis::download_this(
+    output_name = "mspms_data",
+    output_extension = ".csv",
+    button_label = "Download data",
+    button_type = "warning",
+    has_icon = TRUE,
+    icon = "fa fa-save",
+    csv2 = FALSE
+  )
 ```
@@
 DT::datatable(t1)
+
+t1 %>%
+  downloadthis::download_this(
+    output_name = "mspms_anova",
+    output_extension = ".csv",
+    button_label = "Download anova",
+    button_type = "warning",
+    has_icon = TRUE,
+    icon = "fa fa-save",
+    csv2 = FALSE
+  )
 ```
@@
 DT::datatable(t2)
+
+t2 %>%
+  downloadthis::download_this(
+    output_name = "mspms_t-tests",
+    output_extension = ".csv",
+    button_label = "Download ttests",
+    button_type = "warning",
+    has_icon = TRUE,
+    icon = "fa fa-save",
+    csv2 = FALSE
+  )
 ```

### 2024-07-25 README.Rmd
Incorporated new qc check functionality

@@
 
-There are 4 different types of functions in this package.
+There are 5 different types of functions in this package.
 
@@
 
-2. Data processing/ normalization. These functions allow the user to normalize
+2. QC checks. These functions allow the user to conduct quality control checks
+
+3. Data processing/ normalization. These functions allow the user to normalize
 and process the MSP-MS data.
 
-3. Statistics. These methods allow the user to perform basic statistics on the 
+4. Statistics. These methods allow the user to perform basic statistics on the 
 normalized/processed data.
 
-4. Data visualization. These functions allow the user to visualize the data.
+5. Data visualization. These functions allow the user to visualize the data.
 
@@
 library sequences.
+
+**QC checks**
+1. *qc_check()*: Conducts quality control checks on the data.
+2. *plot_qc_check()*: Plots the results of the qc_checks() function.
+3. *find_nd_peptides()*: Finds library peptides that were not detected.
+4. *plot_nd_peptides()*: Plots the results of the find_nd_peptides() function.
 
@@
 position in a vector of motifs.  
-4. *calc_sig_zscore()*: given a matrix of zscores, determines which are significant given a user defined p value. 
+4. *calc_sig_zscore()*: given a matrix of zscores, determines which are
+significant given a user defined p value. 
 5. *calc_AA_percent_difference()*: calculates the percentage difference of 
@@
 ```
+
+## QC Checks
+
+In theory, all of the full length peptides present in the library 
+should be detected in the MSP-MS experiment. In practice, this isn't the case
+because some peptides fly worse than others.
+
+The check_qc() function let's us know what percentage of these full length
+peptides are missing. This function also checks to see what
+
+We can use it as follows:
+
+```{r}
+qc_check <- mspms::qc_check(prepared_data = mspms::peaks_prepared_data,
+                            peptide_library = mspms::peptide_library,
+                            design_matrix = mspms::design_matrix)
+
+head(qc_check)
+```
+
+These results can then be visualized as follows using plot_qc_check()
+
+First lets look at the percent of full length peptides in the library detected
+
+```{r}
+plot_qc_check(qc_check)[1]
+```
+
+We can see here that most samples have more than 90% of the full length peptides
+detected. We would expect to see ~90% of the peptides detected at T0. Notably 
+less than this would suggest that something funky happened to the sample and it
+
+all of these samples look reasonably good.
+
+Now let's look at all the percentage of peptides in the library detected, 
+including both full length and detected fragments.
+
+```{r}
+plot_qc_check(qc_check)[2]
+```
+
+Here we see that most samples have more than 90% of the peptides detected. 
+This is as would be expected with quality data, thus Ww can consider these 
+samples to have passed QC. Note that we see more peptides here compared to the 
+previous qc plot, since some of the digested peptide fragments fly better than
+the parent peptides on the mass spec.
+
+It may be helpful to know what peptides from the library were not detected in a 
+given experiment We can examine this as follows:
+
+```{r}
+nd_peptides = mspms::find_nd_peptides(mspms::peaks_prepared_data,
+                                      mspms::peptide_library,
+                                      mspms::design_matrix)
+
+head(nd_peptides)
+```
+
+we can then visualize as follows using the plot_nd_peptides() function.
+
+```{r}
+plot_nd_peptides(nd_peptides)
+```
+Here we can see that some of the full length peptides are not detected in 100% 
+of samples. Additionally, some of these library peptides are not detected, even
+when considering possible fragment peptides.
 

### 2024-07-25 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
Incorporated new qc check functionality

@@
 
+# QC Checks
+
+```{r}
+qc_checks = mspms::qc_check(prepared_data,peptide_library,design_matrix)
+
+DT::datatable(qc_checks)
+
+qc_checks %>%
+  downloadthis::download_this(
+    output_name = "qc_checks",
+    output_extension = ".csv",
+    button_label = "Download qc_check",
+    button_type = "warning",
+    has_icon = TRUE,
+    icon = "fa fa-save",
+    csv2 = FALSE
+  )
+```
+
+**Table 1.** Here the percentage of the peptide library that is detected in 
+each sample is displayed.
+
+```{r}
+mspms::plot_qc_check(qc_checks)
+
+```
+
+**Figure 1.** Here the results of the qc checks are displayed.
+
+```{r}
+nd_peptides = mspms::find_nd_peptides(prepared_data,
+                                      peptide_library,
+                                      design_matrix)
+
+DT::datatable(nd_peptides)
+
+nd_peptides %>%
+  downloadthis::download_this(
+    output_name = "nd_peptides",
+    output_extension = ".csv",
+    button_label = "Download nd_peptides",
+    button_type = "warning",
+    has_icon = TRUE,
+    icon = "fa fa-save",
+    csv2 = FALSE
+  )
+```
+
+**Table 2.** Here the peptides missing from each sample are shown
+
+```{r}
+plot_nd_peptides(nd_peptides)
+```
+
+**Figure 2.** Here the library peptides not detected in each samples are shown.
+
 # Overview 
@@
 
-**Figure 1.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.
+**Figure 3.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.
 
@@
 
-**Figure 2.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.
+**Figure 4.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.
 
@@
 
-**Table 1. ** Here anova results are displayed for each condition (testing to see if there is an effect on time).
+**Table 3. ** Here anova results are displayed for each condition (testing to see if there is an effect on time).
 
 ```{r}
-t2 = log2fc_t_test(mspms_data) %>% 
+t2 = mspms::log2fc_t_test(mspms_data) %>% 
   dplyr::arrange(p.adj)
@@
 
-**Table 2.** Here T-test results are displayed as compared to time 0 for each condition
+**Table 4.** Here T-test results are displayed as compared to time 0 for each condition
 
 ```{r}
-plot_volcano(t2)
+mspms::plot_volcano(t2)
 ```
 
-**Figure 3.** Here the Log2fc and T tests are displayed as volcano plots relative to T0 for each condition.
+**Figure 5.** Here the Log2fc and T tests are displayed as volcano plots relative to T0 for each condition.
 
@@
 ```{r,fig.height = 10, fig.width = 6}
+
+tryCatch({
   print(mspms::plot_all_icelogos(mspms_data))
+}, error = function(e) {
+  print(e)
+})
 ```
 
-**Figure 4.** Here the cleavage motifs enriched in each condition are shown. 
+**Figure 6.** Here the cleavage motifs enriched in each condition are shown. 
 Enriched cleavage motifs are defined as having a p.adj <= 0.05 and a log2fc >3
@@
 
-d = count_cleavages_per_pos(sig)
+d = mspms::count_cleavages_per_pos(sig)
 
-plot_cleavages_per_pos(d)
+tryCatch({
+  print(mspms::plot_cleavages_per_pos(d))
+}, error = function(e) {
+  print(e)
+})
 ```
 
-**Figure 5.** Here the count of significant cleavages at each position of the 
+**Figure 7.** Here the count of significant cleavages at each position of the 
 peptide library are shown for each condition (facet) per time point (color). 

### 2024-09-04 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
fixes

@@
 
+```{r}
+plot_rt_qc(prepared_data,design_matrix)
+```
+
+**Figure 3.** Here the distribution of peptides detected per retention time are 
+shown for each group.
+
 # Overview 
@@
 
-**Figure 3.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.
+**Figure 4.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.
 
@@
 
-**Figure 4.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.
+**Figure 5.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.
 
@@
 sig = t2 %>% 
-  filter(p.adj <= 0.05, log2fc >3)
+  dplyr::filter(p.adj <= 0.05, log2fc >3)
 

### 2024-09-04 R/generate_report.R

@@ -9,6 +9,7 @@
 #'  cleavage sequences for left and right of the cut site.
 #' @param outdir = this is the output directory you would like to render the
 #' report to.
+#' @param output_file = this is the file name to export. 
 #'
 #' @return a knited .html report of the mspms analysis.
 #' @export
@@ -24,7 +25,11 @@ generate_report <- function(prepared_data,
                             design_matrix,
                             peptide_library = mspms::peptide_library,
                             n_residues = 4,
-                            outdir = getwd()) {
+                            outdir = getwd(),
+                            output_file = paste0(Sys.Date(),
+                                                 "_mspms_report.html")
+                            ) {
+
   rmarkdown::render(
     system.file("rmarkdown/templates/mspms_report/skeleton/skeleton.RMD",
       package = "mspms"
@@ -37,6 +42,7 @@ generate_report <- function(prepared_data,
     ),
     clean = TRUE,
     output_dir = outdir,
-    output_file = paste0(Sys.Date(), "_mspms_report.html")
+    output_file = output_file
   )
+
 }

### 2024-09-19 README.Rmd
major structural changes after Bioconductor feedback

--- 2024-06-04 README.Rmd
@@
 ---
-output:
-  github_document: default
-  self_contained: default
+output: github_document
 ---
@@
 <!-- badges: start -->
+
 [![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
-[![Codecov test coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
-[![BioC status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)
+[![Codecov test
+coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
+[![BioC
+status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)
+
 <!-- badges: end -->
 
-The goal of mspms is provide a concise code-base for the normalization and 
-data processing required to analyze data from the 
-[Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).
+The goal of mspms is provide a R Package that can be used to analyze data from
+the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS)
+method](https://pubmed.ncbi.nlm.nih.gov/36948708/).
 
-Additionally, we provide a [graphical user interface powered by shiny apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to
+Additionally, we provide a [graphical user interface powered by shiny
+apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to
 utilize the method without requiring any R coding knowledge.
@@
 
-To generate a general report using your own data, run the following code. 
-It requires data that has been prepared for mspms data analysis by a 
-converter function, and a design matrix. For more information see subsequent 
-sections.
+To generate a general report using your own data, run the following code. It
+requires data that has been prepared for mspms data analysis by a converter
+function, and a design matrix. For more information see subsequent sections.
 
@@
   prepared_data = mspms::peaks_prepared_data,
-  design_matrix = mspms::design_matrix,
+  colData = mspms::design_matrix,
   outdir = "../Desktop/mspms_report"
@@
 
-The above command will generate a .html file containing a general mspms analysis.
+The above command will generate a .html file containing a generic mspms
+analysis.
 
@@
 
-There are 4 different types of functions in this package.
+Functions in this package are logically divided into 4 broad types of functions.
 
-1. Making mspms generically useful. These functions are focused on making mspms 
-more generally useful to a wider audience outside of the very specific workflow
-traditionally used in the O’Donoghue lab. Allows for the use of other types of
-upstream proteomic data processing, different peptide libraries, etc.
+1.  Pre-processing data. These functions are focused on getting all of the data
+    necessary for a mspms analysis together and in a consistent format.
 
-2. Data processing/ normalization. These functions allow the user to normalize
-and process the MSP-MS data.
+2.  Data processing. These functions allow the user to process the proteomics
+    data (imputation, normalization, etc).
 
@@
 
-**Making mspms generically useful**.   
-1. *prepare_peaks()*:Takes two input files from PEAKS and combines them. 
-2. *prepare_pd()*: prepares exported files from proteome discoverer.  
-3. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide
-library sequences.
+**Note that only a subset of functions are exported to the user.** Helper
+functions can be found in the \_helper_functions.R files.
 
-**Data Processing/ Normalization**.  
-1. *normalyze()*: Normalizes values.     
-2. *handle_outliers()*: Looks for outliers across replicates. Removes them.       
-3. *impute()*: Imputes data for missing values (not including NAs introduced 
-by handle_outliers()).  
-4. *join_with_library()*: Joins the normalized data with the peptide library 
-sequences.  
-5. *add_cleavages()*: Figures out the locations of the detected cleavages within
-the library of peptide sequences used. Is it cleaved at the N or C terminus, 
-or both? Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.      
-6. *polish()*: combines the cleavage info into one column. Discards peptides
-that were cleaved on both sides or not at all.  
-7. *prepare_for_stats()*: Reshapes the data into a long format and appends the 
-data in the design matrix to make statistical testing easy to perform.  
-8 *mspms():* combines all of the above functions into one convenient function.
+## Pre-processing data
 
-**Calculations/ Statistics**.   
-1. *mspms_log2fc()*: Calculates the log2 fold change across experimental 
-conditions relative to T0.  
-2.*mspms_t_tests()*: Conducts T-tests to determine if there are any peptides 
-that are significantly different across experimental conditions relative to time 
-0.  
-3. *log2fc_t_test()*: Combines results from mspms_log2fc() and mspms_t_tests() 
-into one convenient tibble.   
-4. *log2fct_condition()*: Calculates the log2 fold change and t tests for the 
-specified condition at a user specified time.    
-5.*log2fct_time()*: Calculates the log2 fold change and t tests between the 
-specified times within a user specified time.   
-6. *mspms_anova()*: Conducts ANOVA tests to determine if there are any peptides 
-that are significantly different across all time points.   
-7. *count_cleavages_per_pos():* Counts the number of cleavages at each position 
-of the library. Useful for determining endoprotease vs exoprotease activity.
+Pre-processing data here refers to standardizing the data exported from
+different software solutions for analyzing mass spectrometry-based proteomics
+and adding useful information specific to the MSP-MS assay.
 
-**Data Visualization**.   
-1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an
-interactive heatmap to visualize overall patterns in the data.    
-2. *plot_pca()*: PCA analysis to visualize the data in a 2D space.  
-3. *plot_time_course()*: Plot peptides over time by condition.  
-4. *plot_cleavages_per_pos()*: Plot the number of cleavages at each position 
-in the library.Useful for determining endoprotease vs exoprotease activity.
+Specifically, the data from the file exported from the upstream software is
+parsed to contain the **peptide** (detected peptide sequence, cleavages are
+represented as "\_"), **library_id** (name of the peptide in the peptide library
+each detected peptide maps to), and intensities for each sample.
 
-**Icelogo**.
+Information about each peptide relative to the peptide it was derived from is
+then added. This includes the **cleavage_seq** (the motif the peptide a user
+specified n residues up and down from the cleavage event), **cleavage_pos**
+(position of the peptide library the cleavage product was cleaved at ),
+**peptide_type** (whether the detected peptide is a cleavage product of the
+peptide library, or a full length member of the peptide library).
 
-1. *calc_AA_count_of_motif()*: calculates the count of each amino acid at each
-position in a vector of motifs.  
-2. *calc_AA_prop_of_motif()*: calculates the proportion of each amino acid at 
-each position from the output of calc_AA_count_of_motif().  
-3. *calc_AA_motif_zscore()*: calculates z scores for each amino acid at each 
-position in a vector of motifs.  
-4. *calc_AA_percent_difference()*: calculates the percentage difference of 
-5. *prepare_sig_p_dif()*: prepares the data for plotting percent difference 
-icelogos. Removes AAs at positions that are not significantly different.  
-6. *prepare_fc()*: prepares the data for plotting fold change icelogos.
-Removes AAs at positions that are not significantly different.  
-7. *plot_pd_icelogo()*: plots a percent difference icelogo.  
-8. *plot_fc_icelogo()*: plots a fold change ice logo.  
-9. *plot_icelogo()*:  wraps plot_pd_icelogo() and plot_fc_icelogo() into 
-one function.  
-10. *plot_all_icelogos()*: convenience function to plot all icelogos relative to 
-time 0 from an experiment.    
-**Reports**. 
-1. *generate_report()*: Generates a simple analysis of the mspms data that 
-should be generically applicable to all mspms experiments.
+All of the information is then loaded as a QFeatures object along with colData
+(metadata about each sample) where it is stored as a SummarizedExperiment object
+named "peptides".
 
-## Making Generically Usefull.
+#### Uploading your colData
 
-### Formating Peaks File Outputs.
+As briefly described above, the colData contains metadata describing the samples
+you are trying to analyze.
 
-All analysis performed by this package is downstream of the data generated by 
-the PEAKS software. The first step is to combine the two files generated by
-PEAKS.
+It is very important that this contains at least the following columns
+-"quantCols","group","condition",and "time".
 
-The files coming from peaks should be generated according to the instructions 
-found [here](www/PeaksDataAnalysis_howto.pdf).
+The quantCols field contains the name of each sample. These names must match
+whatever the samples are named in the files exported from the upstream
+proteomics software.
+
+The group field contains text specifying what group each sample belongs to. All
+replicate samples should have the same group name
+
+The condition field contains the experimental conditions that vary in your
+MSP-MS experiment (this depends on the experimental question, could be type of
+protease, the type of inhibitor, etc)
+
+The time field contains the duration of incubation of your protease of interest
+with the peptide library for each sample.
+
+An example of a valid colData file is shown below.
+
+```{r}
+colData = readr::read_csv("inst/extdata/colData.csv")
+usethis::use_data(colData,overwrite = TRUE)
+
+head(colData)
+```
+
+#### PEAKS
+
+Analysis using a .csv exported from [PEAKS
+software](https://www.bioinfor.com/peaks-studio/) is supported.
+
+an exported .csv file can be generated from PEAKS as follows:
+
+1.  Create a New Project in PEAKS.\
+2.  Select data to add to PEAKS project.\
+3.  Add data, rename samples, set correct parameters. • Highlight all the added
+    data.\
+    • Select Create new sample for each file.\
+    • Select appropriate instrument and fragmentation method.\
+    • Set Enzyme to “None”.\
+    • Rename samples according to enzymes, time points, and replicates.\
+    • Can also ensure appropriate data is selected for these.\
+    • Select “Data Refinement” to continue.\
+4.  Data Refinement.\
+    • On the top right, select MSP-MS as the predefined parameters.\
+    • Parameters shown are what should be used for MSP-MS data analysis.\
+    • Select “Identification” to continue.\
+5.  Peptide Identification.\
+    • On the top right, select MSP-MS as the predefined parameters.\
+    • Parameters shown are what should be used for MSP-MS data analysis.\
+    • Identification -\> Select Database -\> TPD_237.\
+    • Identification -\> no PTMs.\
+    • Can remove the PTMs by highlighting them and selecting Remove\
+6.  Label Free Quantification.\
+    • Make sure Label Free is selected.\
+    • Group samples -\> add quadruplicates of samples to new group.
 
@@
 
-### Loading the files ###
-lfq_filename <- "tests/testdata/protein-peptides-lfq.csv"
-# file "protein-peptides.csv" exported from PEAKS identification
-id_filename <- "tests/testdata/protein-peptides-id.csv"
+# File path of peaks lfq file
+lfq_filename <- "inst/extdata/peaks_protein-peptides-lfq.csv"
 
-# Prepare the data for normalyzer analysis
-peaks_prepared_data <- prepare_peaks(lfq_filename, id_filename)
+# Prepare the data 
+peaks_prepared_data <- mspms::prepare_peaks(lfq_filename,
+                                            colData,
+                                            0.3,
+                                            n_residues = 4)
+
+usethis::use_data(peaks_prepared_data,overwrite = TRUE)
+
+peaks_prepared_data
 ```
 
-### Formating Proteome Discoverer File Outputs.
+#### Proteome Discoverer
 
-We can prepare proteome discover files for analysis with mspms as follows.
+Analysis using a .csv file exported from [proteome
+discoverer](https://www.thermofisher.com/us/en/home/industrial/mass-spectrometry/liquid-chromatography-mass-spectrometry-lc-ms/lc-ms-software/multi-omics-data-analysis/proteome-discoverer-software.html)
+is supported.
+
+To generate this .csv file:
 
 ```{r}
-prepared_proteome_discoverer <- prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
-```
-
-### Calculating all cleavages
-
-We might want to calculate all possible cleavages for the peptide library 
-sequences. This is useful for downstream analysis, especially when we are 
-looking at the specificity of the cleavage sites via plot_cleavage_motif() as
-this requires a background of all possible cleavages
-
-We can do this by specifying the number of amino acids after the cleavage site 
-that we are interested in. First lets try 4, which is the default.
-
-```{r}
-all_peptide_sequences <- mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
-  n_AA_after_cleavage = 4
-)
-head(all_peptide_sequences)
-```
-
-We could also try 5 AA after each cleavage site
-
-```{r}
-all_peptide_sequences <- mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
-  n_AA_after_cleavage = 5
-)
-head(all_peptide_sequences)
-```
-
-## Data Normalization/ Processing
-
-### Loading design matrix
-
-Before we normalyze data, we need to know what samples are in what groups. We can do that by defining a design matrix. 
-```{r}
-design_matrix <- readr::read_csv("tests/testdata/design_matrix.csv")
-
-head(design_matrix)
-```
-### Normalyzing data
-
-mspms uses the NormalyzerDE package to do normalization under the hood.
-
-We can normalyze the data as follows:  
-```{r}
-normalyzed_data <- normalyze(peaks_prepared_data, design_matrix)
 
@@
 
-### Handling Outliers
+#### Fragpipe
 
-Here we use a Dixon test from the outliers package to detect outliers from each 
-of our replicates. We need to know what samples are part of which groups, so we 
-need to specify the design matrix here too. Make sure that the column header 
-names are "sample" and "group" just like before.
+Data exported from [Fragpipe](https://github.com/Nesvilab/FragPipe) as a .tsv
+file is supported.
+
+To generate this file, run a [LFQ Fragpipe
+workflow](https://fragpipe.nesvilab.org/docs/tutorial_lfq.html) using your
+desired parameters (we recommend using match between runs).
+
+You will find a file named "combined_peptide.tsv" in the output folder from
+Fragpipe.
+
+This can be loaded into mspms as follows:
 
 ```{r}
-design_matrix <- readr::read_csv("tests/testdata/design_matrix.csv")
-outliers <- handle_outliers(normalyzed_data, design_matrix)
+lfq_file_path = "inst/extdata/fragpipe_combined_peptide.tsv"
 
+fragpipe_prepared_data = prepare_fragpipe(lfq_file_path,
+                                          colData)
+
+fragpipe_prepared_data
 ```
 
-### Imputation of data
+## Data Processing
 
-We have a lot of missing, or 0 values. For these, we need to impute them so we 
-can do downstream statistics.
+Data processing includes imputation and normalization of the proteomics data.
+
+**Add more info**
 
 ```{r}
-imputed <- mspms::impute(outliers)
+# does set seed matter?
+set.seed(2)
+processed_qf = process_qf(peaks_prepared_data)
 
-```
+usethis::use_data(processed_qf,overwrite = TRUE)
 
-### Joining with Library
-
-Next we need to join everything with the sequences of the peptide library.  
-```{r}
-joined_with_library <- mspms::join_with_library(imputed)
-```
-
-### Calcuating cleavages.
-
-Next, we need to determine where the peptide sequences are cleaved.
-
-We check both the N and C terminus.
-
-Sequences are presented as the user specified number of amino acids on 
-both sides of a cleavage. The default is 4, but there is interest in looking at
-motifs that are farther away from the cut site. Note that X indicates that there
-was nothing at that position because it is over the end.
-
-```{r}
-cleavage_added_data <- mspms::add_cleavages(joined_with_library, n_residues = 4)
-
-head(cleavage_added_data)
-```
-
-### Adding metadata back
-
-We need to add information about our samples back to perform downstream 
-statistics on the data. We accomplish this with the prepare for stats file.
-
-```{r}
-prepared_for_stats <- mspms::prepare_for_stats(cleavage_added_data,
-                                               design_matrix)
-```
-
-### Polishing
-
-This function polishes the data by combining the cterm and nterm cleavage 
-information into one column while removing any rows that don't have any 
-cleavage information or have cleavage information on the cterm and nterm.
-
-```{r}
-polished_data <- mspms::polish(cleavage_added_data)
-```
-
-### Complete workflow
-
-There are a lot of steps to get to the final normalized data we listed above. 
-We have combined all of these steps into one function, mspms() for convenience.
-This offers a nice balance between allowing the user the option to use 
-individual steps and providing a quick way to get to the final data.
-
-```{r}
-mspms_data <- mspms::mspms(peaks_prepared_data, design_matrix)
+processed_qf
 ```
@@
 
-mspms provides a number of convenience functions to conduct statistics
-on the data.
+MSP-MS analysis has traditionally used T-tests to what peptides are
+significantly different relative to T0. The authors of the assay suggest using a
+T-test derived FDR adjusted p value threshold \<= 0.05 and log2 fold change
+threshold \>= 3.
 
-First, we need to prepare the data for stats. This involves reshaping the data 
-into the long format so it is easy to conduct statistics on and then appending 
-the data in the design matrix that allows us to determine what sample contains 
-what conditions/time and conduct the appropriate statistics.
-
-Now we can conduct the statistics:
-
-### T tests
-
-T tests are performed within each condition and compared to time 0.
-
-For example, for an experiment where there are two conditions, 
-DMSO and MZB as well as 3 time points, 0, 1, and 2, the t tests would be
-as follows:
-
-* DMSO.TO vs DMSO.T0. 
-* DMSO.T0 vs DMSO.T1. 
-* DMSO.T0 vs DMSO.T2.
-
-* MZB.T0 vs MZB.T0. 
-* MZB.T0 vs MZB.T1. 
-* MZB.T0 vs MZB.T2.
+We provide a user-friendly implementation of this statistical test as follows:
 
 ```{r}
-# Perform T test
-t_test_stats <- mspms::mspms_t_tests(mspms_data)
+log2fc_t_test_data = mspms::log2fc_t_test(processed_qf = processed_qf)
+
+usethis::use_data(log2fc_t_test_data,overwrite = TRUE)
 ```
 
-### log2FC
+Please note that statistical analyses are one of the strengths of the
+Bioconductor project and several packages provide alternative statistical
+approaches that are more flexible than t-tests and could be used with the
+QFeatures objects produced by mspms.
 
-We can also calculate the log 2 fold change. 
-The comparisons here are the same as for the T tests.  
+This includes packages using empirical Bayesian approaches (see
+[limma](https://bioconductor.org/packages/release/bioc/html/limma.html)) and
+mixed modelling (see
+[msqrob2](https://bioconductor.org/packages/release/bioc/html/msqrob2.html)),
+that are suited for proteomics data.
+
+Users are encouraged to use these excellent packages.
+
+## Data Visualizations
+
+We provide a number of data visualizations as part of the mspms_package.
+
+### QC Checks
+
+A critical initial step in MSP-MS data analysis is a rigorous quality control
+assessment. Given the reliance on a specific peptide library, calculating the
+percentage of library peptides detected in each sample provides a
+straightforward yet powerful indicator of data integrity. Anomalously low
+detection rates signal potential instrument or experimental problems.
+
+We can see the percentage of the peptide library that were undetected among each
+group of replicates as follows (for full length peptides a threshold of \~\< 10
+is considered good while for cleavage products a threshold \~\< 5 is good):
+
 ```{r}
-# calculate log2fc
-log2fc <- mspms::mspms_log2fc(mspms_data)
-
-head(log2fc)
+plot_qc_check(processed_qf,
+              full_length_threshold = 10,
+              cleavage_product_threshold = 5)
 ```
 
-### log2fc_t_tests.
-
-Oftentimes (such as when you want to make volcano plots) it is most useful to
-look at the log2fc and the t test statistics at the same time.
+We can also see in what percentage of samples each peptide in the peptide
+library were undetected in.
 
 ```{r}
-log2fc_t_test <- mspms::log2fc_t_test(mspms_data)
-
-head(log2fc_t_test)
+plot_nd_peptides(processed_qf)
 ```
 
-### log2fct_condition
+### Heatmap
 
-You may wish to compare two conditions at a specific time point. 
-This function allows this
+We can inspect our data using an interactive heatmap as follows:
 
 ```{r}
-condition_t = mspms::log2fct_condition(mspms_data,
-                                       ref_condition = "DMSO",
-                                       comparison_condition = "MZB",
-                                       at_time = 60)
-
-head(condition_t)
-
+plot_heatmap(processed_qf)
 ```
 
-### log2fct_time
+### PCA
 
-You may wish to compare two time points within a specific condition. 
-This function allows
+We can also inspect our data using a PCA as follows.
 
 ```{r}
-time_t = mspms::log2fct_time(mspms_data,
-                             within_condition = "DMSO",
-                             ref_time = 0,
-                             comparison_time = 60)
-
-head(time_t)
+plot_pca(processed_qf)
 ```
 
-### ANOVA
+### Volcano plots
 
-We also might want to perform an anova. Here, we have it set up to show the 
-effect of time within each condition.
-
-For example, for an experiment where there are two conditions, DMSO and MZB as
-well as 3 time points, 0, 1, and 2, the anova would statistics for the effect of
-time for each peptide within DMSO or MZB.
+We can generate volcano plots showing the significant peptides from each
+condition as follows:
 
 ```{r}
-# Doing ANOVA
-anova_stats <- mspms::mspms_anova(mspms_data)
-
-head(anova_stats)
+plot_volcano(log2fc_t_test_data)
 ```
 
-## Common Data Visualizations
+### Cleavage Position Plots
 
-We also provide some functions that make common data visualizations easier.
+We can also create plots showing the number of times we observe a cleavage at
+each position of the peptide_library among a set of peptides of interest.
 
-### volcano plots
+We can do this as follows:
 
 ```{r}
-volcano_plot <- mspms::plot_volcano(log2fc_t_test)
+sig_cleavage_data = log2fc_t_test_data %>% 
+  dplyr::filter(p.adj <= 0.05,log2fc > 3)
 
-volcano_plot
-```
-
-### Plotting sequence specificity motif
-
-Here use an approach similar to what is implemented in ICELOGO to visualize the
-sequence specificity of the cleavage sites. This was used as reference to build 
-the code: https://iomics.ugent.be/icelogoserver/resources/manual.pdf.
-
-Below we can see this applied to see what cleavage sequences are enriched among
-the peptides significantly different between time 0 and time 60 
-within DMSO.
-
-```{r}
-cleavage_seqs <- time_t %>%
-  dplyr::filter(log2fc > 3, p.adj <= 0.05) %>%
-  dplyr::pull(cleavage_seq)
-
-background_universe <- mspms::all_possible_8mers_from_228_library
-
-mspms::plot_icelogo(cleavage_seqs, background_universe)
-```
-
-We could also look at the fold change instead of the percent difference
-
-```{r}
-mspms::plot_icelogo(cleavage_seqs, background_universe, type = "fold change")
-```
-
-We also provide a convenience function for looking at all icelogos relative to 
-time 0 within the experiment.
-
-This function includes cleavages that were significant at any time point once in 
-the experimental set of the ICELOGO.
-
-```{r}
-mspms::plot_all_icelogos(mspms_data)
-```
-
-### PCA.
-
-We can generate a PCA plot to visualize the data. 
-Here, the colors show the different time points while the shape shows 
-the different conditions.
-
-```{r}
-mspms::plot_pca(mspms_data)
-```
-
-### Hierchical clustering
-
-We can also generate an interactive heatmap with the hierarchical clustering 
-results.
-
-The format of this readme does not allow for interactivity, so a static picture 
-of the output is shown instead.
-
-```{r,eval = FALSE}
-mspms::plot_heatmap(mspms_data)
-```
-
-![](plot_heatmap_output.png)
-
-### Ploting time course.
-
-We also provide a function for plotting the mean intensity and standard 
-deviation over time for each peptide in the data set by conditions.
-Facets show peptide, color shows the condition.
-
-This can be used in combination with the ANOVA function.
-Below, we will use anova to calculate the peptides where we see an 
-effect of time (in the DMSO group), and then plot these.
-
-```{r}
-
-top10sig <- anova_stats %>%
-  dplyr::arrange(p.adj) %>%
-  dplyr::pull(Peptide) %>%
-  head(10)
-
-p1 <- mspms_data %>%
-  dplyr::filter(Peptide %in% top10sig) %>%
-  mspms::plot_time_course()
+p1 = mspms::plot_cleavages_per_pos(sig_cleavage_data )
 
@@
 
-## Calculating the number of cleavages at each position in the library
+### iceLogo
 
-We can also calculate the number of cleavages at each position of the library.
-Here we will do this for peptides significantly different relative to time 0 for
-DMSO and MZB.
+[Icelogos](https://www.nature.com/articles/nmeth1109-786) are a method for
+visualizing conserved patterns in protein sequences through probability theory.
+
+As part of the mspms R package, we have implemented the logic to generate
+[iceLogos](https://genesis.ugent.be/uvpublicdata/icelogo/iceLogo.pdf) in R.
+
+To use this, we first have to define a vector of cleavage sequences that we are
+interested in. We will do this for the significant CatA cleavages below.
 
 ```{r}
-# Performing T tests
-t_tests = log2fc_t_test(mspms_data) %>% 
-  dplyr::filter(p.adj < 0.05,
-                log2fc > 3)
+catA_sig_cleavages = log2fc_t_test_data %>% 
+  dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
+  dplyr::filter(condition == "CatA") %>% 
+  dplyr::pull(cleavage_seq) %>% 
+  unique()
+```
 
-count = mspms::count_cleavages_per_pos(t_tests)
+We also need to know all of the possible cleavages we could see among our
+peptide library. We can generate that as below:
 
-p1 = mspms:::plot_cleavages_per_pos(count)
+```{r}
+all_possible_8mers_from_228_library = calculate_all_cleavages(mspms::peptide_library$library_real_sequence,n_AA_after_cleavage = 4)
 
-p1
+usethis::use_data(all_possible_8mers_from_228_library,overwrite = TRUE)
 ```
+
+We can then generate an iceLogo relative to the background of all possible
+cleavages in our peptide library like so:
+
+```{r}
+plot_icelogo(catA_sig_cleavages,
+             background_universe = all_possible_8mers_from_228_library)
+```
+
+We also provide a function that conveniently plots all icelogos from all
+conditions from the experiment all together.
+
+```{r,fig.height = 10, fig.width = 4}
+sig_cleavage_data = log2fc_t_test_data %>% 
+  dplyr::filter(p.adj <= 0.05,log2fc > 3)
+
+plot_all_icelogos(sig_cleavage_data)
+```
+
+### Plotting a Time Course
+
+We can also plot the intensities of peptides of interest over time.
+
+First we need to define which peptides we are interested in. Here we will use
+the t-test results, and look at the peptide with the max log2fc.
+
+```{r}
+max_log2fc_pep = log2fc_t_test_data %>% 
+  dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
+  dplyr::filter(log2fc ==max(log2fc)) %>% 
+  pull(peptide)
+
+```
+
+We can then filter our QFeatures object to only include this peptide
+
+```{r}
+filter = QFeatures::VariableFilter(field = "peptide",
+               value = max_log2fc_pep,
+               condition = "==")
+
+sel = QFeatures::filterFeatures(processed_qf,filter)
+```
+
+And then plot as follows:
+
+```{r}
+plot_time_course(sel,summarize_by_mean = FALSE)
+```

### 2024-09-19 R/calc_AA_fc.R

@@ -1,65 +1,0 @@
-#' calc_AA_fc
-#'
-#' calculate the fold change of each AA by position.
-#'
-#' @param experimental_prop_matrix this is a matrix of the experimental
-#' proportions (from your vector of cleavage sequences) at each position.
-#' @param background_prop_matrix this is a matrix of the background proportions
-#'  of AAs at each position
-#' @param sig_zscores this is a tibble of the significant zscores.
-#' @return a matrix
-#' @export
-#' @examplesIf isTRUE(FALSE)
-#' 
-calc_AA_fc <- function(experimental_prop_matrix,
-                       background_prop_matrix,
-                       sig_zscores) {
-  #calculating the fold change of each amino acid by position
-  fold_change <- (experimental_prop_matrix * 100) /
-    (background_prop_matrix * 100)
-  
-  # Filtering to only include significant z-scores
-  prepared_fc = mspms::prepare_fc(as.data.frame(fold_change),
-                                  sig_zscores)
-
-  converted_fc <- prepared_fc
-  # Here we run into a problem with infinite values.
-  # If a value is 0 in the count, but greater than that in the background
-  # Icelogo shows infinite value as taking up the entire scale.
-  # We will mimic that here
-
-  max_sum <- max(purrr::map_df(as.data.frame(converted_fc), sum,na.rm = TRUE),
-                 na.rm = TRUE)
-  final_converted_fc <- data.frame(row.names = rownames(converted_fc))
-
-  for (i in seq_len(ncol(converted_fc))) {
-    if (0 %in% converted_fc[, i]) {
-      num_zero <- sum(converted_fc[, i] == 0, na.rm = TRUE)
-
-      col <- replace(
-        converted_fc[, i],
-        converted_fc[, i] == 0,
-        (max_sum / num_zero * -1)
-      )
-
-      final_converted_fc <- cbind(final_converted_fc, col)
-    } else {
-      final_converted_fc <- cbind(final_converted_fc, converted_fc[, i])
-    }
-  }
-  
-  colnames(final_converted_fc) <- colnames(converted_fc)
-
-  # now converting values less than 1 to converted fold change. 
- final_converted_fc2 = final_converted_fc %>% 
-   dplyr::mutate(
-     dplyr::across(dplyr::all_of(names(final_converted_fc)),
-          ~case_when(.x < 1 & .x > 0  ~ (1/.x) * -1,
-                      .default = .x)
-          )
-   )
[…]

### 2024-09-19 R/handle_outliers.R

@@ -1,76 +1,0 @@
-#' handle_outliers
-#'
-#' This function replaces outliers for MSP-MS data with NA based on the results
-#' of dixon tests as implemented in the outliers package
-#'
-#' @param normalyzed_data = this is the normalized data.
-#' Intended to be prepared using mspms::normalyze()
-#' @param design_matrix = this is the design_matrix containing "sample","group",
-#' "condition","time" columns. Intended to be read in using readr::read_csv()
-#'
-#' @return a data frame in the long format where outliers according to a dixon
-#' test have been handled.
-#' @export
-#' @examples
-#'
-#' # Here we handle the outliers from the test data
-#' handled_outliers <- handle_outliers(
-#'   mspms::normalyzed_data,
-#'   mspms::design_matrix
-#' )
-handle_outliers <- function(normalyzed_data, design_matrix) {
-  # dealing with no visible binding for global variable ‘.’ NOTE
-  . <- NULL
-  # Putting into the long format, adding group information
-  index <- which(names(normalyzed_data) == "Protein Accession") + 1
-
-  long_data <- normalyzed_data %>%
-    tidyr::pivot_longer(dplyr::all_of(index:length(.)), names_to = "sample") %>%
-    dplyr::inner_join(design_matrix, by = "sample") %>%
-    dplyr::filter(!is.na(.data$value))
-
-
-  # initalizing the data frame
-  out <- data.frame()
-
-  # Looping through each group to check for outliers.
-  for (i in unique(long_data$group)) {
-    # filtering data to only include  samples with the same condition and time
-    f <- dplyr::filter(long_data, .data$group == i)
-
-    # Looping though each peptide within each group
-    for (j in unique(long_data$Peptide)) {
-      f2 <- dplyr::filter(f, .data$Peptide == j)
-      # # only doing tests if there are between 3 and 30 values
-      if (length(f2$value) > 3 & length(f2$value) < 30) {
-        #
-        #   #is there an outlier?
-        stats <- outliers::dixon.test(f2$value)$p.value
-
-        # what would the outlier be if there was one?
-        outlier <- outliers::outlier(f2$value)
-
-        # If there is a outlier, change the value to 0 for now, NA after impute.
-        if (stats < 0.05) {
-          f2[f2$value == outlier, "value"] <- 0
-        }
-        # #adding the cv
-        # f2 = f2 %>%
[…]

### 2024-09-19 R/plot_nd_peptides.R

@@ -1,44 +1,38 @@
 #' plot_nd_peptides
 #'
-#' plot peptides from libray that were not detected in the data
-#'
-#' @param nd_peptides a tibble containing info on peptides not detected by the 
-#' find_nd_peptides function.
+#' plot the percentage of samples each peptide from library was undetected in
+#' (if the percentage is > 0). 
+#' 
+#' @param processed_qf a QFeatures object containing a SummarizedExperiment
+#' named "peptides"
+#' @param peptide_library_ids a vector of all peptide library ids in the 
+#' experiment.
 #'
 #' @return a ggplot2 object
 #' @export
 #'
 #' @examples
-#' nd_peptides <- find_nd_peptides(
-#'   mspms::peaks_prepared_data,
-#'   mspms::peptide_library,
-#'   mspms::design_matrix
-#' )
-#' plot_nd_peptides(nd_peptides)
-plot_nd_peptides <- function(nd_peptides) {
-  # Calculating number of total samples in study
-  n_samples <- length(unique(nd_peptides$sample))
-
-  # Summarizing per peptide
-  sum <- nd_peptides %>%
-    dplyr::group_by(.data$missing_peptide_id,.data$type_missing) %>%
-    dplyr::summarise(
-      n_missing = n(),
-      percent_samples_undetected =
-        round((.data$n_missing / n_samples * 100), 1)
-    )
+#' plot_nd_peptides(mspms::processed_qf)
+plot_nd_peptides <- function(processed_qf,
+                             peptide_library_ids = 
+                               mspms::peptide_library$library_id) {
+  
+  nd_peptides = calc_per_samples_library_nd(processed_qf,
+                                            peptide_library_ids) %>% 
+    dplyr::filter(.data$per_samples_undetected > 0 )
+    
   # Plotting
-  p1 <- sum %>%
+  p1 <- nd_peptides %>%
     ggplot2::ggplot(ggplot2::aes(
       x = stats::reorder(
-        .data$missing_peptide_id,
-        .data$percent_samples_undetected
+        .data$library_id,
+        .data$per_samples_undetected
       ),
-      y = .data$percent_samples_undetected
+      y = .data$per_samples_undetected
     )) +
     ggplot2::geom_point() +
     ggplot2::coord_flip() +
-    ggplot2::facet_wrap(~.data$type_missing, scales = "free_y") +
+    ggplot2::facet_wrap(~.data$peptide_type, scales = "free_y") +
     ggplot2::xlab("Library ID")
   return(p1)
 }

### 2024-09-19 R/plot_rt_qc.R

@@ -1,47 +1,0 @@
-#' plot_rt_qc
-#' plot the retention times of detected peptides as a histogram or density plot
-#' @param prepared_data  = data prepared by mspms package
-#' @param design_matrix  = design matrix
-#' @param type = type of plot - histogram or density
-#' @param binwidth = binwidth for histogram
-#' @param facet_by_group = whether to facet by group, TRUE or FALSE
-#'
-#' @return a ggplot2 object
-#' @export
-#'
-#' @examples
-#'plot_rt_qc(mspms::peaks_prepared_data,
-#'           mspms::design_matrix,
-#'           type = "density",
-#'           facet_by_group = TRUE)
-plot_rt_qc <- function(prepared_data,
-                       design_matrix,
-                       type = "histogram",
-                       binwidth = 2,
-                       facet_by_group = TRUE) {
-  # Plot the QC of the RTs
-  long <- prepared_data %>%
-    tidyr::pivot_longer(4:length(.), names_to = "sample") %>%
-    dplyr::filter(!is.na(value)) %>% 
-    dplyr::inner_join(design_matrix, by = "sample")
-  
-  # Formatting data in the long format
-  p1 <- ggplot2::ggplot(long, ggplot2::aes(x = RT)) +
-    if (type == "histogram") {
-      ggplot2::geom_histogram(binwidth = binwidth) 
-    } else if (type == "density") {
-      ggplot2::geom_density()
-    } else {
-      stop("type must be histogram or density")
-    }
-  
-  if(isTRUE(facet_by_group)) {
-    p1 <- p1 + ggplot2::facet_wrap(~.data$group, scales = "free_y")
-  }else if(isFALSE(facet_by_group)) {
-    p1 <- p1
-  }else {
-    stop("facet_by_group must be TRUE or FALSE")
-  }
-
-  return(p1)
-}

### 2024-09-19 R/prepare_fc.R

@@ -1,50 +1,0 @@
-#' prepare_fc
-#'
-#' prepare fold changes of AA by position for Icelogo visualization.
-#'
-#' @param fold_change = this is a matrix of the fold changes of the AA by
-#' position.
-#' @param sig_zscores = this is a tibble of the significant zscores.
-#'
-#' @return a matrix of the fold changes of the significant AAs at each position.
-#' @export
-#' @examplesIf isTRUE(FALSE)
-prepare_fc <- function(fold_change, sig_zscores) {
-  # No visible binding for global variable .
-  . <- NULL
-  sig_final <- fold_change %>%
-    tibble::rownames_to_column("AA") %>%
-    tidyr::pivot_longer(2:length(.),
-      names_to = "position",
-      values_to = "fold_change"
-    ) %>%
-    dplyr::mutate(aa_position = paste0(.data$AA, ".", .data$position)) %>%
-    dplyr::filter(.data$aa_position %in% sig_zscores$aa_position) %>%
-    dplyr::select(-.data$aa_position) %>%
-    tidyr::pivot_wider(
-      names_from = .data$position,
-      values_from = .data$fold_change,
-      names_sort = FALSE
-    )
-  # Dealing with the problem introduced by the pivot wider function
-  missing_cols <- ncol(fold_change) - (ncol(sig_final) - 1)
-  missing_data <- as.data.frame(matrix(
-    nrow = nrow(sig_final),
-    ncol = missing_cols, NA
-  ))
-  `%!in%` <- Negate(`%in%`)
-  missing_names <- names(fold_change)[names(fold_change) %!in%
-    names(sig_final)]
-  names(missing_data) <- missing_names
-  final <- sig_final %>%
-    dplyr::bind_cols(missing_data) %>%
-    dplyr::relocate(paste0("P", c(
-      (ncol(fold_change) / 2):1,
-      paste0(
-        seq_len(ncol(fold_change) / 2), "'"
-      )
-    ))) %>%
-    tibble::column_to_rownames("AA") %>%
-    as.matrix()
-  return(final)
-}

### 2024-09-19 R/prepare_fragpipe.R

@@ -1,0 +1,56 @@
+#' prepare_fragpipe
+#' 
+#' Prepare a label free quantification file exported from Fragpipe
+#' for subsequent mspms analysis.
+#'
+#' @param combined_peptide_filepath file path the combined_peptide.tsv file
+#' generated by FragPipe.
+#' @param colData tibble containing colData (sample metadata). Must have
+#' columns named "quantCols","group","condition",and "time". 
+#' @param peptide_library peptide library contained in the experiment 
+#' @param n_residues the number of amino acid residues before and after the
+#' cleavage site to generate a cleavage seq for.
+#'
+#' @return a QFeatures object containing a summarizedExperiment named "peptides" 
+#' @export
+#'
+#' @examples
+#' colData = readr::read_csv("inst/extdata/colData.csv")
+#' combined_peptide_filepath = "inst/extdata/fragpipe_combined_peptide.tsv"
+#' prepared_fragpipe_data = prepare_fragpipe(combined_peptide_filepath)
+prepare_fragpipe = function(combined_peptide_filepath,
+                            colData,
+                            peptide_library = mspms::peptide_library,
+                            n_residues = 4){
+  # Reading in the label free quantification data
+  lfq <- readr::read_tsv(combined_peptide_filepath,na = c("-",0,"","NA")) 
+  # Check that this file appears to be a proper fragpipe file
+  check_file_is_valid_fragpipe(lfq)
+  
+  # Reformat the columns to be consistent with mspsms framework
+  lfq_mod = lfq %>% 
+    dplyr::mutate(`Prev AA` = dplyr::case_when(is.na(`Prev AA`) ~ "",
+                                               TRUE ~ `Prev AA`)) %>% 
+    dplyr::mutate(`Next AA` = dplyr::case_when(is.na(`Next AA`) ~ "",
+                                               TRUE ~ `Next AA`)) %>% 
+    dplyr::mutate(peptide = paste0(`Prev AA`,"_",
+                                   `Peptide Sequence`,"_",`Next AA`),
+                  .before =1) %>% 
+    dplyr::mutate(peptide = gsub("^_","",peptide),
+                  peptide = gsub("_$","",peptide)) %>% 
+    dplyr::select(peptide,library_id = "Protein",
+                  dplyr::contains("MaxLFQ Intensity")) %>% 
+    dplyr::rename_with(
[…]

### 2024-09-19 R/qc_check.R

@@ -1,52 +1,0 @@
-#' qc_check
-#' Run simple quality control checks on the data. This checks to see how many
-#' full length peptides belonging to the library were identified in the data,
-#' as well as how many peptides were detected among all cleavage products
-#'
-#' @param prepared_data data prepared for analysis by the mspms package.
-#' @param peptide_library the peptide library used in the experiment
-#' @param design_matrix the design matrix containing experimental metadata
-#' @return a list with two slots, 1st slot contains a tibble with QC metrics
-#' second list contains and interpretation. 2nd slot contains a tibble with the
-#' missing peptides in each sample.
-#' @export
-#'
-#' @examples
-#' qc_check_data <- qc_check(
-#'   mspms::peaks_prepared_data,
-#'   mspms::peptide_library,
-#'   mspms::design_matrix
-#' )
-qc_check <- function(prepared_data, peptide_library, design_matrix) {
-  # Preparing the data for QC analysis
-  long <- prepared_data %>%
-    # column 4 is the start of the data
-    tidyr::pivot_longer(4:length(.), names_to = "sample") %>%
-    dplyr::filter(!is.na(value))
-  # Counting # of peptides in library
-  library_num <- length(peptide_library$library_reference_id)
-  # checking to see how many full length peptides were found in each sample
-  check_1 <- long %>%
-    dplyr::filter(!grepl(".*_.*", Peptide)) %>%
-    dplyr::group_by(sample) %>%
-    dplyr::summarise(n_undigested_library_detected = dplyr::n_distinct(Peptide))
-  # Checking to see how many components of the library are detected total
-  check_2 <- long %>%
-    dplyr::group_by(sample) %>%
-    dplyr::summarise(
-      n_library_detected =
-        dplyr::n_distinct(`Protein Accession`)
-    )
-  # Combining the checks
-  out <- dplyr::right_join(check_1, check_2, by = "sample") %>%
-    dplyr::mutate(
-      perc_undigested_library_detected =
-        round((n_undigested_library_detected / library_num * 100), 1),
-      perc_library_detected = round(
-        n_library_detected / library_num * 100, 1
-      )
-    ) %>%
-    dplyr::inner_join(design_matrix, by = "sample")
-
-  return(out)
-}

### 2024-09-19 R/qc_helper_functions.R

@@ -1,0 +1,121 @@
+#' Find library peptides that were not detected in the data.
+#'
+#' @param prepared_data a tibble with data prepared by the mspms package.
+#' @param peptide_library a tibble containing the peptide library used
+#' @param design_matrix a tibble containing the design matrix
+#'
+#' @return a tibble with the missing peptides for each sample
+#' @keywords internal
+#'
+#' @examples
+#' nd_peptides <- find_nd_peptides(
+#'   mspms::peaks_prepared_data,
+#'   mspms::peptide_library,
+#'   mspms::design_matrix
+#' )
+#'
+calc_per_samples_library_nd <- function(processed_qf,
+                                        peptide_library_ids = mspms::peptide_library$library_id) {
+  # tidying the data so we can work with it
+  mspms_data = mspms_tidy(processed_qf,"peptides")
+  #calculating n of samples
+  n_samples = length(unique(mspms_data$quantCols))
+  # converting to wide format, easier to reason with this way
+  wide = mspms_data %>% 
+    dplyr::select("peptide","library_id","peptides",
+                  "peptide_type","quantCols") 
+  # considering full length peptides only   
+  full_length = wide %>% 
+    dplyr::filter(peptide_type == "full_length") %>% 
+    tidyr::pivot_wider(names_from = "quantCols", 
+                       values_from = "peptides")
+  # What peptides are not detected at all
+  nd_full = peptide_library_ids[peptide_library_ids %!in% unique(
+    full_length$library_id)]
+  # Building a tibble of the peptides not detected
+  nd_full = tibble::tibble(library_id = nd_full,
+                           n_samples = n_samples,
+                           n_missing = n_samples)
+# Figuring the number of full length library ids missing per sample
+  n_missing_full_length = tibble::tibble(library_id = full_length$library_id,
+                                         n_samples = n_samples,
+                                         n_missing = rowSums(is.na(
+                                           full_length)))
+  # Combining the completely missing with partially missing data
+  full = dplyr::bind_rows(n_missing_full_length,nd_full) %>% 
+    dplyr::mutate(peptide_type ="full_length",.after = "library_id")
+  # Now considering cleavage products, conting the number of non nas, per sample
+  cleavage_product = wide %>% 
+    dplyr::filter(peptide_type == "cleavage_product") %>% 
+    dplyr::select(-"peptide") %>% 
+    tidyr::pivot_wider(names_from = quantCols, 
+                       values_from = peptides,
[…]

### 2024-09-26 README.Rmd
Passed Rcmd Check

@@
 )
+```
 
-```
+<img src="man/figures/logo.png" align="right" width="170"/>
 
@@
 [![Codecov test
-coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
-[![BioC
-status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)
+coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=bioconductor)
 
@@
 
-The goal of mspms is provide a R Package that can be used to analyze data from
-the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS)
-method](https://pubmed.ncbi.nlm.nih.gov/36948708/).
+## Introduction
+
+The goal of mspms is provide a R Package that can be used to easily and
+reproducibly analyze data resulting from the [Multiplex Substrate Profiling by
+Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).
 
@@
 
+Multiplex Substrate Profiling by Mass Spectrometry
+[(MSP-MS)](https://pubmed.ncbi.nlm.nih.gov/36948708/) is a powerful method used
+to determine the substrate specificity of proteases. This method is of interest
+for groups interested in the study of proteases and their role as regulators of
+biological pathways, whether applied to the study of disease states, the
+development of diagnostic and prognostic tests, generation of tool compounds, or
+rational design of protease targeting therapeutics. Analysis of the MS based
+data produced by MSP-MS is a multi-step process involving detection and
+quantification of peptides, imputation, normalization, cleavage sequence
+identification, statistics, and data visualizations. This process can be
+challenging, especially for scientists with limited programming experience. To
+overcome these issues, we developed the mspms R package to facilitate the
+analysis of MSP-MS data utilizing good software/data analysis practices.
+
+mspms differs from existing proteomics packages that are available in the
+Bioconductor project in that it is specifically designed to analyze MSP-MS data.
+This involves unique preprocessing steps and data visualizations tailored to
+this specific assay
+
+In order to do so, mspms uses several excellent packages from the Bioconductor
+project to provide a framework to easily and reproducibly analyze MSP-MS data.
+These include:
+
+-   QFeatures: management and processing of peptide quantitative features and
+    sample metadata.
+
+-   MsCoreUtils: log2 transformation, imputation (QRILC), and normalization
+    (center.median).
+
+Some excellent non-Bioconductor packages are also utilized.
+
+-   ggplot2: data visualizations
+
+-   heatmaply: interactive heatmaps
+
+-   rstatix: conduct statistics
+
+-   ggseqlogo: iceLogo motif visualization.
+
+-   downloadthis: gives users the ability to download data from standard .html
+    report generated by mspms::generate_report().
+
+Lastly, mspms takes advantage of iceLogos ([Colaert, N. et al. Nature Methods 6,
+786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to viusalize
+overrepresented amino acid motifs relative to a background set by implementing
+components of the Java software in R.
+
 ## Installation
 
-You can install the development version of mspms from github.
+You can install the stable release from Bioconductor:
+
+Alternatively, you can install the development version from github.
 
@@
 devtools::install_github("baynec2/mspms",
-                         ref = "develop")
+  ref = "develop"
+)
 ```
@@
   prepared_data = mspms::peaks_prepared_data,
-  colData = mspms::design_matrix,
   outdir = "../Desktop/mspms_report"
@@
 
-#### Uploading your colData
+#### A note on colData
 
@@
 ```{r}
-colData = readr::read_csv("inst/extdata/colData.csv")
-usethis::use_data(colData,overwrite = TRUE)
+colData <- readr::read_csv(system.file("extdata/colData.csv", package = "mspms"))
 
@@
 # File path of peaks lfq file
-lfq_filename <- "inst/extdata/peaks_protein-peptides-lfq.csv"
+lfq_filepath <- system.file("extdata/peaks_protein-peptides-lfq.csv",
+  package = "mspms"
+)
+
+colData_filepath <- system.file("extdata/colData.csv", package = "mspms")
 
 # Prepare the data
-peaks_prepared_data <- mspms::prepare_peaks(lfq_filename,
-                                            colData,
+peaks_prepared_data <- mspms::prepare_peaks(lfq_filepath,
+  colData_filepath,
   0.3,
-                                            n_residues = 4)
-
-usethis::use_data(peaks_prepared_data,overwrite = TRUE)
+  n_residues = 4
+)
 
@@
 ```{r}
-
 ```
@@
 ```{r}
-lfq_file_path = "inst/extdata/fragpipe_combined_peptide.tsv"
+combined_peptide_filepath <- system.file("extdata/fragpipe_combined_peptide.tsv",
+  package = "mspms"
+)
 
-fragpipe_prepared_data = prepare_fragpipe(lfq_file_path,
-                                          colData)
+fragpipe_prepared_data <- prepare_fragpipe(
+  combined_peptide_filepath,
+  colData_filepath
+)
 
@@
 Data processing includes imputation and normalization of the proteomics data.
+This is all done in a QFeatures object using MSCoreUtils methods. Data from 
+each step is stored as a SummarizedExperiment within the QFeatures object with
+the indicated name.
 
-**Add more info**
+1. Data is first log2 transformed ("peptides_log").   
+2. Data is normalized using the center.median method ("peptides_log_norm").  
+3. Missing values are then imputed using the QRILC method 
+("peptides_log_impute_norm"). 
+4. Data is then reverse log2 transformed ("peptides_norm").
 
 ```{r}
-# does set seed matter?
-set.seed(2)
-processed_qf = process_qf(peaks_prepared_data)
-
-usethis::use_data(processed_qf,overwrite = TRUE)
-
+processed_qf <- process_qf(peaks_prepared_data)
 processed_qf
@@
 ```{r}
-log2fc_t_test_data = mspms::log2fc_t_test(processed_qf = processed_qf)
-
-usethis::use_data(log2fc_t_test_data,overwrite = TRUE)
+log2fc_t_test_data <- mspms::log2fc_t_test(processed_qf = processed_qf)
 ```
@@
 
-We provide a number of data visualizations as part of the mspms_package.
+We provide a number of data visualizations as part of the mspms package.
 
@@
   full_length_threshold = 10,
-              cleavage_product_threshold = 5)
+  cleavage_product_threshold = 5
+)
 ```
@@
 
+### Tidying our data
+
+We can then convert our QFeatures object into a "long" tibble using the 
+mspms_tidy function.
+
+This makes the data easy to manipulate prior to ggplot2 or plotly based plotting 
+functions.
+
+```{r}
+mspms_tidy_data <- mspms_tidy(processed_qf)
+```
+
 ### Heatmap
@@
 ```{r}
-plot_heatmap(processed_qf)
+plot_heatmap(mspms_tidy_data)
 ```
@@
 ```{r}
-plot_pca(processed_qf)
+plot_pca(mspms_tidy_data, value_colname = "peptides_norm")
 ```
@@
 ```{r}
-sig_cleavage_data = log2fc_t_test_data %>% 
-  dplyr::filter(p.adj <= 0.05,log2fc > 3)
+sig_cleavage_data <- log2fc_t_test_data %>%
+  dplyr::filter(p.adj <= 0.05, log2fc > 3)
 
-p1 = mspms::plot_cleavages_per_pos(sig_cleavage_data )
+p1 <- mspms::plot_cleavages_per_pos(sig_cleavage_data)
 
@@
 ```{r}
-catA_sig_cleavages = log2fc_t_test_data %>% 
-  dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
+catA_sig_cleavages <- log2fc_t_test_data %>%
+  dplyr::filter(p.adj <= 0.05, log2fc > 3) %>%
   dplyr::filter(condition == "CatA") %>%
@@
 ```{r}
-all_possible_8mers_from_228_library = calculate_all_cleavages(mspms::peptide_library$library_real_sequence,n_AA_after_cleavage = 4)
-
-usethis::use_data(all_possible_8mers_from_228_library,overwrite = TRUE)
+all_possible_8mers_from_228_library <- calculate_all_cleavages(
+  mspms::peptide_library$library_real_sequence,
+  n_AA_after_cleavage = 4
+)
 ```
@@
 plot_icelogo(catA_sig_cleavages,
-             background_universe = all_possible_8mers_from_228_library)
+  background_universe = all_possible_8mers_from_228_library
+)
 ```
@@
 ```{r,fig.height = 10, fig.width = 4}
-sig_cleavage_data = log2fc_t_test_data %>% 
-  dplyr::filter(p.adj <= 0.05,log2fc > 3)
+sig_cleavage_data <- log2fc_t_test_data %>%
+  dplyr::filter(p.adj <= 0.05, log2fc > 3)
 
@@
 ```{r}
-max_log2fc_pep = log2fc_t_test_data %>% 
-  dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
-  dplyr::filter(log2fc ==max(log2fc)) %>% 
+max_log2fc_pep <- log2fc_t_test_data %>%
+  dplyr::filter(p.adj <= 0.05, log2fc > 3) %>%
+  dplyr::filter(log2fc == max(log2fc)) %>%
   pull(peptide)
-
 ```
 
-We can then filter our QFeatures object to only include this peptide
+We can then filter our tidy mspms data to only include this peptide and plot
 
 ```{r}
-filter = QFeatures::VariableFilter(field = "peptide",
-               value = max_log2fc_pep,
-               condition = "==")
+p1 <- mspms_tidy_data %>%
+  dplyr::filter(peptide == max_log2fc_pep) %>%
+  plot_time_course()
 
-sel = QFeatures::filterFeatures(processed_qf,filter)
+p1
 ```
 
-And then plot as follows:
-
 ```{r}
-plot_time_course(sel,summarize_by_mean = FALSE)
+sessionInfo()
 ```

### 2024-09-26 R/calculate_all_cleavages.R

@@ -1,19 +1,13 @@
 #' calculate_all_cleavages
-#'
 #' calculate all possible cleavages for a defined peptide library containing
 #' peptides of the same length.
-#'
-#' @param peptide_library_seqs The sequences of each peptide in the peptide 
+#' @param peptide_library_seqs The sequences of each peptide in the peptide
 #' library. They should all be the same length.
-#'  
-#' @param n_AA_after_cleavage  The number of AA after (and before) the cleavage 
-#' site to consider. 
-#'
-#' @return a vector of all the possible cleavages for the peptide library 
+#' @param n_AA_after_cleavage  The number of AA after (and before) the cleavage
+#' site to consider.
+#' @return a vector of all the possible cleavages for the peptide library
 #' sequences
-#'  
 #' @export
-#'
 #' @examples
 #' calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
 #'   n_AA_after_cleavage = 4
@@ -22,54 +16,46 @@ calculate_all_cleavages <- function(peptide_library_seqs,
                                     n_AA_after_cleavage = 4) {
   # calculate the length of the peptides in library.
   length <- unique(nchar(peptide_library_seqs))
-
   # The current code assumes all the sequences are the same length
   if (length(length) > 1) {
     stop("Peptide library sequences are not of the same length")
   }
-
   # first cut at every single position
-
   peptide_sequences <- c()
-
   # for each position go through and cut into two pieces
   for (i in seq_len(length - 1)) {
-    ## Doing this for the first fragment, left side of clevage
+    ## Doing this for the first fragment, left side of cleavage
     first_fragment <- substr(peptide_library_seqs, 1, i)
-
     n_char <- unique(nchar(first_fragment))
-
     if (n_AA_after_cleavage - n_char > 0) {
       n_x <- paste0(rep("X", n_AA_after_cleavage - n_char), collapse = "")
     } else {
       n_x <- ""
     }
     first_fragment_x <- paste0(n_x, first_fragment)
-
+    first_fragment_x <- substr(first_fragment_x,
+      stop = unique(nchar(first_fragment_x)),
+      start = unique(nchar(first_fragment_x)) -
+        n_AA_after_cleavage + 1
+    )
     ## Doing this for the second fragment, right side of cleavage
     second_fragment <- substr(
       peptide_library_seqs, i + 1,
       i + n_AA_after_cleavage
     )
-
[…]

### 2024-09-26 R/data.R

@@ -1,10 +1,13 @@
-#' all possible 8mers from the standard 228 peptide library that can
-#'
-#' equivalent to the result of callling mspms::calculate_all_cleavages() on a
-#'  vector of the 14 AA peptides used in the library.
+#' all_possible_8mers_from_228_library
+#' All possible 8mers from the standard (as of 26April2024)
+#' 228 MSP-MS peptide library
+#' (This is equivalent to the result of
+#' mspms::calculate_all_cleavages(mspms::peptide_library$real_cleavage_seq,n=4))
+#' vector of the 14 AA peptides used in the library.
 #' @format ## `all_possible_8mers_from_228_library`
 #' A vector with 2964 entries
-#' @source <Peptide library used in the O’Donoghue lab as of 26April2024>
+#' @source <standard peptide library used with MSP-MS method
+#' in the O’Donoghue lab as of 26April2024>
 "all_possible_8mers_from_228_library"
 
 #' peptide_library
@@ -24,14 +27,13 @@
 #' }
 #' @source <O’Donoghue lab as of 26April2024 >
 "peptide_library"
-
-
 #' peaks_prepared_data
-#' this is an example dataset from peaks that has been processed by mspms.
+#' A QFeatures object prepared from PEAKS data of cathepsin data/.
 #' @format ## `peaks_prepared_data`
-#' A tibble with 820 rows and 27 columns:
+#' An instance of class QFeatures containing 1 assays:
+#' [1] peptides: SummarizedExperiment with 2071 rows and 42 columns
 #' \describe{
-#'   \item{Peptide}{Peptide Sequence Detected}
+#'   \item{peptides}{Peptide Sequence Detected}
 #'   ...
 #' }
 #' @source <mspms processed data originally from PEAKS files found in
@@ -39,5 +41,47 @@
 #'  "tests/testdata/protein-peptides-lfq.csv">
 "peaks_prepared_data"
 
+#' processed_qf
+#' A QFeatures object prepared from PEAKS data of Cathepsin data that has been
+#' processed (imputation/normalization)
+#' @format ## `peaks_prepared_data`
+#' An instance of class QFeatures containing 5 assays:
+#' [1] peptides: SummarizedExperiment with 2071 rows and 42 columns
+#' [2] peptides_log: SummarizedExperiment with 2071 rows and 42 columns
[…]

### 2024-09-26 R/extract_re.R

@@ -1,67 +1,0 @@
-#' extract_re
-#' 
-#' extract regular expressions from an icelogo matrix. 
-#'
-#' @param icelogo_matrix = this is an icelogo matrix, as produced by 
-#' prepare_icelogo_data()
-#' @param type = "positive" or "negative" depending on whether you want to look
-#' for positively or negatively enriched amino acids
-#' @param threshold = the threshold to consider amino acids for inclusion. If 
-#' type is positive, will return AAs with values greater than the threshold.
-#' If type is negative, will return AAs with values less than the threshold.
-#' 
-#' If no amino acids are found greater than the threshold, a . will be returned 
-#' in regular expression. 
-#'
-#' @return a character vector containing a regular expression extracted from 
-#' the icelogo matrix based on user defined parameters. 
-#' @export
-#'
-#' @examples
-#'icelogo_matrix = data.frame(P4 = c(10,NA,NA),
-#'                            P3 = c(NA,10,NA),
-#'                           P2 = c(NA,NA,10),
-#'                           P1 = c(NA,NA,NA),
-#'                            `P1'` = c(NA,NA,NA),
-#'                            `P2'` = c(10,10,10),
-#'                            `P3'` = c(10,10,NA),
-#'                            `P4'` = c(NA,10,10))
-#'rownames(icelogo_matrix) = c("A","B","C")
-#'icelogo_matrix = as.matrix(icelogo_matrix)
-#'extract_re(icelogo_matrix)
-#' 
-extract_re <- function(icelogo_matrix,
-                       type = "positive",
-                       threshold = 0) {
-  ## Defining function to finding row names from icelogo matrix for
-  # enriched AAs depending on type and percent threshold cutoff.
-  find_rownames <- function(x) {
-    if (type == "positive") {
-      row_name <- names(which(x > threshold))
-    } else if (type == "negative") {
-      row_name <- names(which(x < threshold))
-    } else {
-      stop("type must be either positive or negative")
-    }
-    if (length(row_name) == 0) {
-      row_name <- "."
-    }
-    return(row_name)
-  }
-  ## Defining function to extract regular expressions from row names
-  re <- function(row_names) {
-    ## Preparing regular expressions from Row names)
-    step1 <- purrr::map(row_names, paste, collapse = "|")
-    step2 <- purrr::map(step1, append, "(", after = 0)
[…]

### 2024-09-26 R/plot_all_icelogos.R

@@ -1,26 +1,25 @@
 #' plot_all_icelogos
 #'
-#' Easily plot a iceLogo corresponding to peptides of interest across each 
+#' Easily plot a iceLogo corresponding to peptides of interest across each
 #' condition of an experiment.
 #' @param sig_cleavage_data a tibble of data of interest containing a column
 #' labeled peptide, cleavage_seq, and condition
-#' @param pval this is the pvalue threshold (<=) to consider significant when 
-#' determining the significance of the sig_cleavages relative to the background 
+#' @param pval this is the pvalue threshold (<=) to consider significant when
+#' determining the significance of the sig_cleavages relative to the background
 #' at each position of the iceLogo.
-#' @param typethis is the type of iceLogo you would like to generate,
+#' @param type this is the type of iceLogo you would like to generate,
 #'  can be either "percent_difference" or "fold_change".
 #' @param background_universe this is a list cleavages you would like to compare
 #' to as background of the iceLogo
 #'
 #' @return a ggplot object that shows the motif of the cleavage sequences
 #' @export
-#'
 #' @examples
 #' # Determining cleavages of interest
-#' sig_cleavage_data <- mspms::log2fc_t_test_data %>% 
-#' dplyr::filter(p.adj <= 0.05,log2fc > 3) 
+#' sig_cleavage_data <- mspms::log2fc_t_test_data %>%
+#'   dplyr::filter(p.adj <= 0.05, log2fc > 3)
 #' # Plotting a iceLogo for each condition.
-#'plot_all_icelogos(sig_cleavage_data)
+#' plot_all_icelogos(sig_cleavage_data)
 plot_all_icelogos <- function(sig_cleavage_data,
                               type = "percent_difference",
                               pval = 0.05,
@@ -28,23 +27,20 @@ plot_all_icelogos <- function(sig_cleavage_data,
                                 mspms::all_possible_8mers_from_228_library) {
   # initiating plot list
   plot_list <- list()
-
   # Plotting Icelogos
   for (i in unique(sig_cleavage_data$condition)) {
     # filtering data
     f <- dplyr::filter(sig_cleavage_data, .data$condition == i) %>%
       dplyr::select("peptide", "cleavage_seq") %>%
       # appending peptide to cleavage_seqs
-      dplyr::pull(.data$cleavage_seq) %>% 
+      dplyr::pull(.data$cleavage_seq) %>%
       unique()
-
     # generating ice logo
[…]

### 2024-09-26 R/plot_icelogo.R

@@ -4,66 +4,69 @@
 #' background as implemented in the iceLogo method.
 #' https://iomics.ugent.be/icelogoserver/resources/manual.pdf
 #'
-#' @param cleavage_seqs  these are the cleavage sequences of interest 
-#' @param background_universe  this is a list of cleavage sequences 
-#' to use as the background in building the iceLogo. 
-#' @param pval  this is the pvalue threshold (<=) to consider significant when 
-#' determining the significance of the sig_cleavages relative to the background 
+#' @param cleavage_seqs  these are the cleavage sequences of interest
+#' @param background_universe  this is a list of cleavage sequences
+#' to use as the background in building the iceLogo.
+#' @param pval  this is the pvalue threshold (<=) to consider significant when
+#' determining the significance of the sig_cleavages relative to the background
 #' at each position of the iceLogo.
 #' @param type this is the type of visualization you would like to perform,
 #' accepted values are  either "percent_difference" or "fold_change".
-#' @return a ggplot2 object 
+#' @return a ggplot2 object
 #' @export
 #'
 #' @examples
 #' # Determining significant cleavages for catA
-#' catA_sig_cleavages = mspms::log2fc_t_test_data %>% 
-#' dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
-#' dplyr::filter(condition == "CatA") %>% 
-#' dplyr::pull(cleavage_seq) %>% 
-#' unique()
-#' 
+#' catA_sig_cleavages <- mspms::log2fc_t_test_data %>%
+#'   dplyr::filter(p.adj <= 0.05, log2fc > 3) %>%
+#'   dplyr::filter(condition == "CatA") %>%
+#'   dplyr::pull(cleavage_seq) %>%
+#'   unique()
+#'
 #' # Plotting icelogo
 #' plot_icelogo(catA_sig_cleavages,
-#'background_universe = all_possible_8mers_from_228_library)
+#'   background_universe = all_possible_8mers_from_228_library
+#' )
 plot_icelogo <- function(cleavage_seqs,
-                         background_universe = mspms::all_possible_8mers_from_228_library,
+                         background_universe =
+                           mspms::all_possible_8mers_from_228_library,
                          pval = 0.05,
                          type = "percent_difference") {
-  # calculate n 
-  n = length(cleavage_seqs)
-# processing data to plot
[…]

### 2024-09-26 R/plotting_helper_functions.R

@@ -1,11 +1,142 @@
-#' icelogo_col_scheme
+#' calc_per_samples_library_nd
+#' Calculate the percentage of samples each library_id peptide was not detected
+#' in.
+#' @param processed_qf a QFeatures object with a SummarizedExperiment named
+#' "peptides". Intended to be prepared by one of the pre-processing
+#' prepare_x_data functions of the mspms R package.
+#' @param peptide_library_ids a character vector containing the names of the
+#' library_ids
+#' @return a tibble containing percentage of samples each library id was
+#' detected in,  both as full length, and as cleavage products.
+#' @keywords internal
+
+calc_per_samples_library_nd <- function(processed_qf,
+                                        peptide_library_ids =
+                                          mspms::peptide_library$library_id) {
+  # tidying the data so we can work with it
+  mspms_data <- mspms_tidy(processed_qf, "peptides")
+  # calculating n of samples
+  n_samples <- length(unique(mspms_data$quantCols))
+  # converting to wide format, easier to reason with this way
+  wide <- mspms_data %>%
+    dplyr::select(
+      "peptide", "library_id", "peptides",
+      "peptide_type", "quantCols"
+    )
+  # considering full length peptides only
+  full_length <- wide %>%
+    dplyr::filter(.data$peptide_type == "full_length") %>%
+    tidyr::pivot_wider(
+      names_from = "quantCols",
+      values_from = "peptides"
+    )
+  # What peptides are not detected at all
+  nd_full <- peptide_library_ids[peptide_library_ids %!in% unique(
+    full_length$library_id
+  )]
+  # Building a tibble of the peptides not detected
+  nd_full <- tibble::tibble(
+    library_id = nd_full,
+    n_samples = n_samples,
+    n_missing = n_samples
+  )
+  # Figuring the number of full length library ids missing per sample
+  n_missing_full_length <- tibble::tibble(
+    library_id = full_length$library_id,
+    n_samples = n_samples,
+    n_missing = rowSums(is.na(
+      full_length
+    ))
+  )
+  # Combining the completely missing with partially missing data
+  full <- dplyr::bind_rows(n_missing_full_length, nd_full) %>%
+    dplyr::mutate(peptide_type = "full_length", .after = "library_id")
[…]

### 2024-09-26 R/prepare_peaks.R

@@ -3,54 +3,54 @@
 #' for subsequent mspms analysis.
 #' @param lfq_filepath  this is the file path to a .csv file exported from
 #' PEAKS
-#' @param colData tibble containing colData (sample metadata). Must have
-#' columns named "quantCols","group","condition",and "time". 
-#' @param quality_threshold only consider peptides with quality scores > than 
+#' @param colData_filepath file path to .csv file containing colData.
+#' Must have columns named "quantCols","group","condition",and "time".
+#' @param quality_threshold only consider peptides with quality scores > than
 #' this threshold.
-#' @param peptide_library peptide library used in the experiment. 
+#' @param peptide_library peptide library used in the experiment.
 #' @param n_residues the number of amino acid residues before and after the
 #' cleavage site to generate a cleavage seq for.
 #' @return a QFeatures object containing a summarizedExperiment named "peptides"
 #' @export
 #' @examples
-#' colData = readr::read_csv("inst/extdata/colData.csv")
-#' lfq_filepath = "inst/extdata/peaks_protein-peptides-lfq.csv"
-#' prepared_peaks_data = prepare_peaks(lfq_filepath)
-prepare_peaks = function(lfq_filepath,
-                         colData,
-                         quality_threshold = 0.3,
-                         peptide_library = mspms::peptide_library,
-                         n_residues = 4){
-  
+#' lfq_filepath <- system.file("extdata/peaks_protein-peptides-lfq.csv", package = "mspms")
+#' colData_filepath <- system.file("extdata/colData.csv", package = "mspms")
+#' # Prepare the data
+#' peaks_prepared_data <- mspms::prepare_peaks(lfq_filepath, colData_filepath)
+prepare_peaks <- function(lfq_filepath,
+                          colData_filepath,
+                          quality_threshold = 0.3,
+                          peptide_library = mspms::peptide_library,
+                          n_residues = 4) {
   # Reading in the label free quantification data
-  lfq <- readr::read_csv(lfq_filepath,na = c("-",0,"","NA")) 
+  lfq <- readr::read_csv(lfq_filepath, na = c("-", 0, "", "NA"))
   # Check that this file appears to be a proper peaks file
   check_file_is_valid_peaks(lfq)
-   # Filtering out low quality peptides 
-  quality = lfq %>% 
-    dplyr::filter(.data$Quality > quality_threshold) 
[…]

### 2024-09-26 R/process_qf.R

@@ -1,41 +1,39 @@
 #' process_qf
 #'
-#' @param prepared_qf this is a QFeatures object containing a 
+#' @param prepared_qf this is a QFeatures object containing a
 #' SummarizedExperiment named "peptides"
 #'
-#' @return a QFeatures object containing a 
+#' @return a QFeatures object containing a
 #' SummarizedExperiments named "peptides","peptides_log","peptides_log_norm",
 #' "peptides_log_impute_norm",and "peptides_norm"
-#' 
 #' @export
-#'
 #' @examples
-#' processed_qf = process_qf(mspms::peaks_prepared_data)
-process_qf = function(prepared_qf){
-
+#' processed_qf <- process_qf(mspms::peaks_prepared_data)
+process_qf <- function(prepared_qf) {
   # Log2transform the data, add as assay
   prepared_qf_norm <- QFeatures::addAssay(prepared_qf,
-                                       QFeatures::logTransform(
-                                         prepared_qf[["peptides"]]
-                                         ),
-                                       name = "peptides_log")
-  
+    QFeatures::logTransform(
+      prepared_qf[["peptides"]]
+    ),
+    name = "peptides_log"
+  )
   # Normalize the data, add as assay
   prepared_qf_norm <- QFeatures::normalize(prepared_qf_norm,
-                                       i = "peptides_log",
-                                       name = "peptides_log_norm",
-                                       method = "center.median")
+    i = "peptides_log",
+    name = "peptides_log_norm",
+    method = "center.median"
+  )
   # Impute the data, add as assay
-  QF_mod <- QFeatures::impute(prepared_qf_norm,method = "QRILC",
-                              i = "peptides_log_norm",
-                              name = "peptides_log_impute_norm",MARGIN = 2)
-
-  # Remove Outliers if I can figure it out. 
-
-  # Here we log2 transform the data
-  QF_mod <- QFeatures::aggregateFeatures(QF_mod, i = "peptides_log_impute_norm",
-                                         "peptide",name = "peptides_norm",
-                                         fun = rlog2)
-  
+  QF_mod <- QFeatures::impute(prepared_qf_norm,
+    method = "QRILC",
+    i = "peptides_log_norm",
+    name = "peptides_log_impute_norm", MARGIN = 2
+  )
+  # reverse log2 transform the data
+  QF_mod <- QFeatures::aggregateFeatures(QF_mod,
+    i = "peptides_log_impute_norm",
[…]

### 2024-09-26 tests/testthat/test-calc_AA_count_of_motif.R

@@ -1,14 +1,14 @@
 test_that("works with standard number of AAs)", {
-  expect_no_error(mspms::calc_AA_count_of_motif("XXXXXXXX"))
+  expect_no_error(mspms:::calc_AA_count_of_motif("XXXXXXXX"))
 })
 
 
 
 test_that("works with non standard number of AAs (so long as even number)", {
-  expect_no_error(mspms::calc_AA_count_of_motif("XXXXXXXXXX"))
+  expect_no_error(mspms:::calc_AA_count_of_motif("XXXXXXXXXX"))
 })
 
 
 test_that("does not work with odd number of AAs", {
-  expect_error(mspms::calc_AA_count_of_motif("XXX"))
+  expect_error(mspms:::calc_AA_count_of_motif("XXX"))
 })

### 2024-09-26 tests/testthat/test-find_nd_peptides.R

@@ -1,5 +1,0 @@
-test_that("function works", {
-  expect_no_error(find_nd_peptides(mspms::peaks_prepared_data,
-                                   mspms::peptide_library,
-                                   mspms::design_matrix))
-})

### 2024-09-26 tests/testthat/test-polish.R

@@ -1,48 +1,0 @@
-test_that("all peptide cleavages are dropped", {
-  
-  cleavage_added_data = tibble::tibble(nterm = "AAAAAAAA",
-         cterm = "AAAAAAAA",
-         cterm_cleavage_pos = 13,
-         nterm_cleavage_pos = 1) 
-  
-  polished = mspms::polish(cleavage_added_data)
-  
-  expect_equal(nrow(polished),0)
-})
-
-test_that("cterm cleavages retained", {
-  
-  cleavage_added_data = tibble::tibble(nterm = NA,
-                                       cterm = "AAAAAAAA",
-                                       cterm_cleavage_pos = 13,
-                                       nterm_cleavage_pos = NA) 
-  
-  polished = mspms::polish(cleavage_added_data)
-  
-  expect_equal(nrow(polished),1)
-})
-
-test_that("nterm retained", {
-  
-  cleavage_added_data = tibble::tibble(nterm = "AAAAAAAA",
-                                       cterm = NA,
-                                       cterm_cleavage_pos = NA,
-                                       nterm_cleavage_pos = 1) 
-  
-  polished = mspms::polish(cleavage_added_data)
-  
-  expect_equal(nrow(polished),1)
-})
-
-test_that("full length peptides retained", {
-  
-  cleavage_added_data = tibble::tibble(nterm = NA,
-                                       cterm = NA,
-                                       cterm_cleavage_pos = NA,
-                                       nterm_cleavage_pos = NA) 
-  
-  polished = mspms::polish(cleavage_added_data)
-  
-  expect_equal(nrow(polished),1)
-})
-

### 2024-09-26 tests/testthat/test-prepare_pd.R

@@ -1,15 +1,0 @@
-test_that("there are no duplicate peptides after processing", {
-  expect_true({
-    dat <- mspms::prepare_pd("../testdata/proteome_discoverer_output.xlsx")
-
-    dat %>%
-      dplyr::pull(Peptide) %>%
-      dplyr::n_distinct() == nrow(dat)
-  })
-})
-
-test_that("error to find unexpected id columns works", {
-  expect_error({
-    mspms::prepare_pd(filepath = "../testdata/proteome_discoverer_output_bad_names.xlsx")
-  })
-})

### 2024-09-26 tests/testthat/test-preprocessing_helper_functions.R

@@ -1,0 +1,51 @@
+### check_file_is_valid_peaks()###
+test_that("check_file_is_valid_peaks() errors when PTM is provided.", {
+  peaks_ptm <- readr::read_csv(system.file(
+    "extdata/peaks_protein-peptides-lfq.csv",
+    package = "mspms"
+  )) %>%
+    dplyr::mutate(PTM = "PTM")
+  expect_error(mspms:::check_file_is_valid_peaks(peaks_ptm))
+})
+
+### check_file_is_valid_fragpipe()###
+test_that("check_file_is_valid_fragpipe() errors when peaks file is
+          provided.", {
+  peaks_ptm <- readr::read_csv("tests/testdata/
+                               check_file_is_valid_peaks_ptm.csv")
+  expect_error(mspms:::check_file_is_valid_fragpipe(peaks_ptm))
+})
+
+### consolidate_cleavages()###
+
+test_that("consolidate_cleavages() removes double cleaved peptides", {
+  test <- tibble::tibble(
+    peptide = "P_AAAAAAAAAAAAA",
+    nterm = "P_AAAAAAAA_P",
+    nterm_cleavage_pos = 5,
+    cterm = "P_AAAAAAA_P",
+    cterm_cleavage_pos = 5
+  )
+  nrow <- nrow(mspms:::consolidate_cleavages(test))
+  expect_equal(nrow, 0)
+})
+
+### prepared_to_qf ###
+test_that("prepared_to_qf() generates an error when supplied a QFeatures
+          object", {
+  expect_error(mspms:::prepared_to_qf(mspms::peaks_prepared_data,
+    colData = mspms::colData
+  ))
+})
+
+### load_colData() ###
+test_that("load_colData() generates expected expected error when colData is
+          off", {
+  expect_error(
+    mspms:::load_colData(
+      "tests/testdata/missing_time.csv",
+      "colData must have columns named \"quantCols\",\"group\",
+                       \"condition\", and \"time\" you are missing time"
+    )
+  )
+})

### 2024-09-26 tests/testthat/test-statistics_helper_functions.R

@@ -1,0 +1,69 @@
+### remaining_cd_names() ###
+test_that("remaining_cd_names() returns the expected missing name, when time is
+provided as reference", {
+  time <- mspms:::remaining_cd_names(mspms::processed_qf, "time")
+  expect_equal(time, "condition")
+})
+
+test_that("remaining_cd_names() returns the expected missing name, when condition
+is provided as reference", {
+  condition <- mspms:::remaining_cd_names(mspms::processed_qf, "condition")
+  expect_equal(condition, "time")
+})
+
+test_that("remaining_cd_names() throws error when column not in data is
+          attempted to be passed", {
+  expect_error(mspms:::remaining_cd_names(mspms::processed_qf, "not_here"))
+})
+
+### mspms_log2fc() ###
+test_that("mspms_log2fc() returns does not error when data is as expected", {
+  expect_no_error(mspms:::mspms_log2fc(
+    mspms::processed_qf,
+    "time",
+    0
+  ))
+})
+
+test_that("mspms_log2fc() returns does not error when time is supplied as a
+          character", {
+  expect_no_error(mspms:::mspms_log2fc(
+    mspms::processed_qf,
+    "time",
+    "0"
+  ))
+})
+
+test_that("mspms_log2fc() errors when a variable not in data is provided", {
+  expect_error(mspms:::mspms_log2fc(
+    mspms::processed_qf,
+    "not_here",
+    "0"
+  ))
+})
+
+### mspms_t_tests() ###
+test_that("mspms_t_tests() does not error when data is as expected", {
+  expect_no_error(mspms:::mspms_t_tests(
+    mspms::processed_qf,
+    "time",
+    "0"
+  ))
+})
+
+test_that("mspms_t_tests() does not error when numeric value of reference
+          variable is provided", {
+  expect_no_error(mspms:::mspms_t_tests(
+    mspms::processed_qf,
+    "time",
+    0
+  ))
+})
+
+test_that("mspms_t_tests() errors when nonexistant value is provided", {
+  expect_error(mspms:::mspms_t_tests(
+    mspms::processed_qf,
+    "not_here",
+    "not_here"
+  ))
+})

### 2024-09-26 README.Rmd
passed Rcmd check
rcmd check pass

@@
 [![Codecov test
-coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=bioconductor)
+coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
 
@@
 ```{r}
+set.seed(2)
 processed_qf <- process_qf(peaks_prepared_data)

### 2024-09-26 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd

--- 2024-07-03 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
@@
   prepared_data: NULL
-  design_matrix: NULL
   peptide_library: mspms::peptide_library
@@
 
-This report has been automatically generated by the mspms package. It contains standard results of the mspms analysis pipeline.
+This report has been automatically generated by the mspms package. 
+It contains standard results of the mspms analysis pipeline.
 
@@
 ```
-# Data Processing
+
+# Processed Data
 
@@
 # The standard mspms data processing pipeline is used here. 
-mspms_data = mspms::mspms(prepared_data,design_matrix)
+processed_qf = mspms::process_qf(prepared_data)
+
+mspms_data = mspms_tidy(processed_qf,"peptides_norm")
+
+DT::datatable(mspms_data)
 
@@
 
+**Table 1.** Imputed and normalized mspms data.
+
+# QC Checks
+
+```{r}
+mspms::plot_qc_check(processed_qf)
+
+```
+
+**Figure 1.** Here the results of the qc checks are displayed.
+
+```{r}
+mspms::plot_nd_peptides(processed_qf)
+```
+**Figure 2.** Here the library peptides not detected in each samples are shown.
+
 # Overview 
 ```{r}
-mspms::plot_pca(mspms_data)
+mspms_tidy_data <- mspms_tidy(processed_qf)
+mspms::plot_pca(mspms_tidy_data)
 ```
 
-**Figure 1.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.
+**Figure 3.** Here a PCA of the experiment is displayed. The color represents
+time, while shape and linetype of eclipse 
+(if there are enough points to calculate) represents condition.
 
 ```{r,fig.height = 8, fig.width = 10}
-mspms::plot_heatmap(mspms_data)
+mspms::plot_heatmap(mspms_tidy_data)
 ```
 
-**Figure 2.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.
+**Figure 4.** Here an interactive heatmap of the experiment is displayed. 
+Mouse over to see more information on each cell. Column side colors 
+represent condition and time.
 
@@
 ```{r}
-t1 = mspms::mspms_anova(mspms_data) %>% 
-  dplyr::arrange(p.adj)
-
-DT::datatable(t1)
-
-t1 %>%
-  downloadthis::download_this(
-    output_name = "mspms_anova",
-    output_extension = ".csv",
-    button_label = "Download anova",
-    button_type = "warning",
-    has_icon = TRUE,
-    icon = "fa fa-save",
-    csv2 = FALSE
-  )
-```
-
-**Table 1. ** Here anova results are displayed for each condition (testing to see if there is an effect on time).
-
-```{r}
-t2 = log2fc_t_test(mspms_data) %>% 
+t2 = mspms::log2fc_t_test(processed_qf) %>% 
   dplyr::arrange(p.adj)
@@
 
-**Table 2.** Here T-test results are displayed as compared to time 0 for each condition
+**Table 4.** Here T-test results are displayed as compared to time 0 for each 
+condition
 
 ```{r}
-plot_volcano(t2)
+mspms::plot_volcano(t2)
 ```
 
-**Figure 3.** Here the Log2fc and T tests are displayed as volcano plots relative to T0 for each condition.
+**Figure 5.** Here the Log2fc and T tests are displayed as volcano plots 
+relative to T0 for each condition.
 
@@
 
-```{r,fig.height = 10, fig.width = 6}
-print(mspms::plot_all_icelogos(mspms_data))
+```{r,fig.height = 14, fig.width = 4}
+sig_cleavage_data <- t2 %>% 
+  dplyr::filter(p.adj <= 0.05, log2fc >3)
+
+tryCatch({
+  print(mspms::plot_all_icelogos(sig_cleavage_data))
+}, error = function(e) {
+  print(e)
+})
 ```
-
-**Figure 4.** Here the cleavage motifs enriched in each condition are shown. 
+**Figure 6.** Here the cleavage motifs enriched in each condition are shown. 
 Enriched cleavage motifs are defined as having a p.adj <= 0.05 and a log2fc >3
@@
 ```{r}
-sig = t2 %>% 
-  filter(p.adj <= 0.05, log2fc >3)
 
-d = count_cleavages_per_pos(sig)
+d = mspms::plot_cleavages_per_pos(sig_cleavage_data)
 
-plot_cleavages_per_pos(d)
+tryCatch({
+  print(d)
+}, error = function(e) {
+  print(e)
+})
 ```
 
-**Figure 5.** Here the count of significant cleavages at each position of the 
+**Figure 7.** Here the count of significant cleavages at each position of the 
 peptide library are shown for each condition (facet) per time point (color). 

### 2024-09-27 README.Rmd
added support for proteome discoverer exports.

@@
 
-#### Proteome Discoverer
-
-Analysis using a .csv file exported from [proteome
-discoverer](https://www.thermofisher.com/us/en/home/industrial/mass-spectrometry/liquid-chromatography-mass-spectrometry-lc-ms/lc-ms-software/multi-omics-data-analysis/proteome-discoverer-software.html)
-is supported.
-
-To generate this .csv file:
-
-```{r}
-```
-
 #### Fragpipe
@@
 
+colData_filepath <- system.file("extdata/colData.csv", package = "mspms")
+
 fragpipe_prepared_data <- prepare_fragpipe(
@@
 fragpipe_prepared_data
+```
+#### Proteome Discoverer
+
+Analysis of a PeptideGroups.txt file exported from [proteome
+discoverer](https://www.thermofisher.com/us/en/home/industrial/mass-spectrometry/liquid-chromatography-mass-spectrometry-lc-ms/lc-ms-software/multi-omics-data-analysis/proteome-discoverer-software.html) is supported.
+
+To generate this .txt file:
+1. Ensure that non-normalized/ non-scaled abundances are included in your 
+PeptideGroup table.   
+2. File > export > To Text (tab delimited) > Items to be exported - check 
+the Peptide Groups box > Export.
+
+Note that the fileIDs of each sample in proteome discoverer (By default 
+F1 - Fn) must match the quantCols in colData.
+
+```{r}
+peptide_groups_filepath <- system.file(
+  "extdata/proteome_discoverer_PeptideGroups.txt",
+  package = "mspms"
+)
+
+colData_filepath <- system.file("extdata/proteome_discover_colData.csv",
+                                package = "mspms")
+
+prepared_pd_data = prepare_pd(peptide_groups_filepath,colData_filepath)
+
+prepared_pd_data
 ```

### 2024-09-27 R/prepare_pd.R

@@ -1,0 +1,62 @@
+#' Title
+#' Prepare a label free quantification file exported from Proteome Discoverer
+#' for subsequent mspms analysis.
+#' @param peptide_groups_filepath filepath to PeptideGroups.txt file exported
+#' from proteome discoverer.
+#' @param proteome_discover_id_map filepath to PeptideGroups.txt file exported
+#' from proteome discoverer. Must contain non normalized and non scaled
+#' abundances.
+#' @param colData_filepath file path to .csv file containing colData.
+#' Must have columns named "quantCols","group","condition",and "time".
+#' @param peptide_library peptide library used with experiment. Contains
+#' columns "library_id", "library_match_sequence", and "library_real_sequence".
+#' @param n_residues the number of amino acid residues before and after the
+#' cleavage site to generate a cleavage seq for.
+#' @return a QFeatures object containing a summarizedExperiment named "peptides"
+#' @export
+#' @examples
+#' peptide_groups_filepath <- system.file("extdata/proteome_discoverer_PeptideGroups.txt", package = "mspms")
+#' colData_filepath <- system.file("extdata/colData.csv", package = "mspms")
+prepare_pd <- function(peptide_groups_filepath,
+                       colData_filepath,
+                       peptide_library = mspms::peptide_library,
+                       n_residues = 4) {
+  # Reading in the label free quantification data
+  lfq <- readr::read_delim(peptide_groups_filepath)
+  # Check that this file appears to be a proper fragpipe file
+  check_file_is_valid_pd(lfq)
+  # Reformat the columns to be consistent with mspsms framework
+  lfq <- lfq %>%
+    dplyr::rename(
+      peptide = "Annotated Sequence",
+      library_id = "Master Protein Accessions"
+    )
+  # Converting the peptide notation from proteome discover to standard format.
+  lfq <- lfq %>%
+    dplyr::mutate(
+      peptide = gsub("\\[-\\]", "", .data$peptide),
+      peptide = gsub("^\\.", "", .data$peptide),
+      peptide = gsub("\\.$", "", .data$peptide),
+      peptide = gsub("\\].", "_", .data$peptide),
+      peptide = gsub("\\.\\[", "_", .data$peptide),
+      peptide = gsub("\\]", "", .data$peptide),
+      peptide = gsub("\\[", "", .data$peptide)
+    )
[…]

### 2024-09-27 tests/testthat/test-preprocessing_helper_functions.R

@@ -38,6 +38,32 @@ test_that("prepared_to_qf() generates an error when supplied a QFeatures
   ))
 })
 
+
+test_that("prepared_to_qf() fails with unexpected colData", {
+  
+  prepared_data = mspms::processed_qf %>% 
+    mspms_tidy("peptides") %>% 
+    dplyr::select(peptide,library_id,quantCols,peptides) %>% 
+    tidyr::pivot_wider(names_from = "quantCols", values_from = "peptides")
+ 
+   wrong_col_data = mspms::colData %>% 
+    mutate(quantCols = 1:length(quantCols))
+   
+   expect_error(mspms:::prepared_to_qf(prepared_data,wrong_col_data))
+    
+})
+
+test_that("prepared_to_qf() works with expected colData", {
+  
+  prepared_data = mspms::processed_qf %>% 
+    mspms_tidy("peptides") %>% 
+    dplyr::select(peptide,library_id,quantCols,peptides) %>% 
+    tidyr::pivot_wider(names_from = "quantCols", values_from = "peptides")
+
+  expect_no_error(mspms:::prepared_to_qf(prepared_data,mspms::colData))
+  
+})
+
 ### load_colData() ###
 test_that("load_colData() generates expected expected error when colData is
           off", {

### 2024-09-27 README.Rmd
styler

@@
 colData_filepath <- system.file("extdata/proteome_discover_colData.csv",
-                                package = "mspms")
+  package = "mspms"
+)
 
-prepared_pd_data = prepare_pd(peptide_groups_filepath,colData_filepath)
+prepared_pd_data <- prepare_pd(peptide_groups_filepath, colData_filepath)
 

### 2024-09-27 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
styler

@@
 ```
+
 **Figure 2.** Here the library peptides not detected in each samples are shown.
@@
 ```
+
 **Figure 6.** Here the cleavage motifs enriched in each condition are shown. 

### 2024-09-27 tests/testthat/test-preprocessing_helper_functions.R

@@ -40,28 +40,24 @@ test_that("prepared_to_qf() generates an error when supplied a QFeatures
 
 
 test_that("prepared_to_qf() fails with unexpected colData", {
-  
-  prepared_data = mspms::processed_qf %>% 
-    mspms_tidy("peptides") %>% 
-    dplyr::select(peptide,library_id,quantCols,peptides) %>% 
+  prepared_data <- mspms::processed_qf %>%
+    mspms_tidy("peptides") %>%
+    dplyr::select(peptide, library_id, quantCols, peptides) %>%
     tidyr::pivot_wider(names_from = "quantCols", values_from = "peptides")
- 
-   wrong_col_data = mspms::colData %>% 
+
+  wrong_col_data <- mspms::colData %>%
     mutate(quantCols = 1:length(quantCols))
-   
-   expect_error(mspms:::prepared_to_qf(prepared_data,wrong_col_data))
-    
+
+  expect_error(mspms:::prepared_to_qf(prepared_data, wrong_col_data))
 })
 
 test_that("prepared_to_qf() works with expected colData", {
-  
-  prepared_data = mspms::processed_qf %>% 
-    mspms_tidy("peptides") %>% 
-    dplyr::select(peptide,library_id,quantCols,peptides) %>% 
+  prepared_data <- mspms::processed_qf %>%
+    mspms_tidy("peptides") %>%
+    dplyr::select(peptide, library_id, quantCols, peptides) %>%
     tidyr::pivot_wider(names_from = "quantCols", values_from = "peptides")
 
-  expect_no_error(mspms:::prepared_to_qf(prepared_data,mspms::colData))
-  
+  expect_no_error(mspms:::prepared_to_qf(prepared_data, mspms::colData))
 })
 
 ### load_colData() ###

### 2024-09-28 R/generate_report.R
4 spaces for indents

@@ -22,20 +22,20 @@ generate_report <- function(prepared_data,
                             n_residues = 4,
                             outdir = getwd(),
                             output_file = paste0(
-                              Sys.Date(),
-                              "_mspms_report.html"
+                                Sys.Date(),
+                                "_mspms_report.html"
                             )) {
-  rmarkdown::render(
-    system.file("rmarkdown/templates/mspms_report/skeleton/skeleton.RMD",
-      package = "mspms"
-    ),
-    params = list(
-      prepared_data = prepared_data,
-      peptide_library = peptide_library,
-      n_residues = n_residues
-    ),
-    clean = TRUE,
-    output_dir = outdir,
-    output_file = output_file
-  )
+    rmarkdown::render(
+        system.file("rmarkdown/templates/mspms_report/skeleton/skeleton.RMD",
+            package = "mspms"
+        ),
+        params = list(
+            prepared_data = prepared_data,
+            peptide_library = peptide_library,
+            n_residues = n_residues
+        ),
+        clean = TRUE,
+        output_dir = outdir,
+        output_file = output_file
+    )
 }

### 2024-09-28 R/plot_cleavages_per_pos.R

@@ -11,21 +11,21 @@
 #' @examples
 #' # Defining the significant peptides
 #' sig_cleavage_data <- log2fc_t_test_data %>%
-#'   dplyr::filter(p.adj <= 0.05, log2fc > 3)
+#'     dplyr::filter(p.adj <= 0.05, log2fc > 3)
 #' # Plotting
 #' p1 <- mspms::plot_cleavages_per_pos(sig_cleavage_data)
 #' p1
 plot_cleavages_per_pos <- function(sig_cleavage_data,
                                    ncol = 1) {
-  count_cleavages_per_pos <- count_cleavages_per_pos(sig_cleavage_data)
-  p1 <- count_cleavages_per_pos %>%
-    ggplot2::ggplot(ggplot2::aes(
-      x = .data$cleavage_pos,
-      y = .data$n,
-      color = .data$time
-    )) +
-    ggplot2::facet_wrap(~ .data$condition, scales = "free_y") +
-    ggplot2::geom_point() +
-    ggplot2::geom_line()
-  return(p1)
+    count_cleavages_per_pos <- count_cleavages_per_pos(sig_cleavage_data)
+    p1 <- count_cleavages_per_pos %>%
+        ggplot2::ggplot(ggplot2::aes(
+            x = .data$cleavage_pos,
+            y = .data$n,
+            color = .data$time
+        )) +
+        ggplot2::facet_wrap(~ .data$condition, scales = "free_y") +
+        ggplot2::geom_point() +
+        ggplot2::geom_line()
+    return(p1)
 }

### 2024-09-28 R/plot_pca.R

@@ -19,18 +19,18 @@ plot_pca <- function(mspms_tidy_data,
                      value_colname = "peptides_norm",
                      color = "time",
                      shape = "condition") {
-  color <- dplyr::sym(color)
-  shape <- dplyr::sym(shape)
-  all <- prepare_for_PCA(mspms_tidy_data)
-  plot <- all[[1]] %>%
-    ggplot2::ggplot(ggplot2::aes(.data$PC1, .data$PC2,
-      color = !!color,
-      linetype = !!shape, shape = !!shape
-    )) +
-    ggplot2::geom_point() +
-    ggplot2::stat_ellipse(level = 0.95) +
-    ggplot2::xlab(paste0("PCA-1 ", round(all[[2]][1] * 100, 2), "%")) +
-    ggplot2::ylab(paste0("PCA-2 ", round(all[[2]][2] * 100, 2), "%"))
+    color <- dplyr::sym(color)
+    shape <- dplyr::sym(shape)
+    all <- prepare_for_PCA(mspms_tidy_data)
+    plot <- all[[1]] %>%
+        ggplot2::ggplot(ggplot2::aes(.data$PC1, .data$PC2,
+            color = !!color,
+            linetype = !!shape, shape = !!shape
+        )) +
+        ggplot2::geom_point() +
+        ggplot2::stat_ellipse(level = 0.95) +
+        ggplot2::xlab(paste0("PCA-1 ", round(all[[2]][1] * 100, 2), "%")) +
+        ggplot2::ylab(paste0("PCA-2 ", round(all[[2]][2] * 100, 2), "%"))
 
-  return(plot)
+    return(plot)
 }

### 2024-09-28 R/plotting_helper_functions.R

@@ -12,80 +12,80 @@
 
 calc_per_samples_library_nd <- function(processed_qf,
                                         peptide_library_ids =
-                                          mspms::peptide_library$library_id) {
-  # tidying the data so we can work with it
-  mspms_data <- mspms_tidy(processed_qf, "peptides")
-  # calculating n of samples
-  n_samples <- length(unique(mspms_data$quantCols))
-  # converting to wide format, easier to reason with this way
-  wide <- mspms_data %>%
-    dplyr::select(
-      "peptide", "library_id", "peptides",
-      "peptide_type", "quantCols"
+                                            mspms::peptide_library$library_id) {
+    # tidying the data so we can work with it
+    mspms_data <- mspms_tidy(processed_qf, "peptides")
+    # calculating n of samples
+    n_samples <- length(unique(mspms_data$quantCols))
+    # converting to wide format, easier to reason with this way
+    wide <- mspms_data %>%
+        dplyr::select(
+            "peptide", "library_id", "peptides",
+            "peptide_type", "quantCols"
+        )
+    # considering full length peptides only
+    full_length <- wide %>%
+        dplyr::filter(.data$peptide_type == "full_length") %>%
+        tidyr::pivot_wider(
+            names_from = "quantCols",
+            values_from = "peptides"
+        )
+    # What peptides are not detected at all
+    nd_full <- peptide_library_ids[peptide_library_ids %!in% unique(
+        full_length$library_id
+    )]
+    # Building a tibble of the peptides not detected
+    nd_full <- tibble::tibble(
+        library_id = nd_full,
+        n_samples = n_samples,
+        n_missing = n_samples
     )
-  # considering full length peptides only
-  full_length <- wide %>%
-    dplyr::filter(.data$peptide_type == "full_length") %>%
-    tidyr::pivot_wider(
-      names_from = "quantCols",
-      values_from = "peptides"
+    # Figuring the number of full length library ids missing per sample
+    n_missing_full_length <- tibble::tibble(
+        library_id = full_length$library_id,
+        n_samples = n_samples,
+        n_missing = rowSums(is.na(
+            full_length
+        ))
     )
-  # What peptides are not detected at all
[…]

### 2024-09-28 R/process_qf.R

@@ -10,30 +10,30 @@
 #' @examples
 #' processed_qf <- process_qf(mspms::peaks_prepared_data)
 process_qf <- function(prepared_qf) {
-  # Log2transform the data, add as assay
-  prepared_qf_norm <- QFeatures::addAssay(prepared_qf,
-    QFeatures::logTransform(
-      prepared_qf[["peptides"]]
-    ),
-    name = "peptides_log"
-  )
-  # Normalize the data, add as assay
-  prepared_qf_norm <- QFeatures::normalize(prepared_qf_norm,
-    i = "peptides_log",
-    name = "peptides_log_norm",
-    method = "center.median"
-  )
-  # Impute the data, add as assay
-  QF_mod <- QFeatures::impute(prepared_qf_norm,
-    method = "QRILC",
-    i = "peptides_log_norm",
-    name = "peptides_log_impute_norm", MARGIN = 2
-  )
-  # reverse log2 transform the data
-  QF_mod <- QFeatures::aggregateFeatures(QF_mod,
-    i = "peptides_log_impute_norm",
-    "peptide", name = "peptides_norm",
-    fun = rlog2
-  )
-  return(QF_mod)
+    # Log2transform the data, add as assay
+    prepared_qf_norm <- QFeatures::addAssay(prepared_qf,
+        QFeatures::logTransform(
+            prepared_qf[["peptides"]]
+        ),
+        name = "peptides_log"
+    )
+    # Normalize the data, add as assay
+    prepared_qf_norm <- QFeatures::normalize(prepared_qf_norm,
+        i = "peptides_log",
+        name = "peptides_log_norm",
+        method = "center.median"
+    )
+    # Impute the data, add as assay
+    QF_mod <- QFeatures::impute(prepared_qf_norm,
+        method = "QRILC",
+        i = "peptides_log_norm",
+        name = "peptides_log_impute_norm", MARGIN = 2
+    )
+    # reverse log2 transform the data
+    QF_mod <- QFeatures::aggregateFeatures(QF_mod,
+        i = "peptides_log_impute_norm",
+        "peptide", name = "peptides_norm",
+        fun = rlog2
+    )
+    return(QF_mod)
 }

### 2024-09-28 tests/testthat/test-cterm_cleavage.R

@@ -1,78 +1,78 @@
 # Testing for case where there is no cleavage site
 testthat::test_that("works with no cleavage site", {
-  testthat::expect_equal(
-    mspms:::cterm_cleavage(
-      "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",
-      "ABCDEFGHIJKLMN", 4
-    ),
-    tibble::tibble(
-      peptide = "ABCDEFGHIJKLMN",
-      cterm = NA,
-      cterm_cleavage_pos = NA
+    testthat::expect_equal(
+        mspms:::cterm_cleavage(
+            "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",
+            "ABCDEFGHIJKLMN", 4
+        ),
+        tibble::tibble(
+            peptide = "ABCDEFGHIJKLMN",
+            cterm = NA,
+            cterm_cleavage_pos = NA
+        )
     )
-  )
 })
 
 
 # Testing for case where there is no cleavage site
 testthat::test_that("works with one clevage site", {
-  testthat::expect_equal(
-    mspms:::cterm_cleavage(
-      "ABCDEFGHIJKLM_N", "ABCDEFGHIJKLMN",
-      "ABCDEFGHIJKLMN", 4
-    ),
-    tibble::tibble(
-      peptide = "ABCDEFGHIJKLM_N",
-      cterm = "JKLMNXXX",
-      cterm_cleavage_pos = 13
+    testthat::expect_equal(
+        mspms:::cterm_cleavage(
+            "ABCDEFGHIJKLM_N", "ABCDEFGHIJKLMN",
+            "ABCDEFGHIJKLMN", 4
+        ),
+        tibble::tibble(
+            peptide = "ABCDEFGHIJKLM_N",
+            cterm = "JKLMNXXX",
+            cterm_cleavage_pos = 13
+        )
     )
-  )
 })
 
 
 # Testing for case where there are two cleavage sites
 testthat::test_that("works with two clevage site", {
-  testthat::expect_equal(
-    mspms:::cterm_cleavage(
-      "A_BCDEFGHIJKLM_N", "ABCDEFGHIJKLMN",
-      "ABCDEFGHIJKLMN", 4
-    ),
-    tibble::tibble(
-      peptide = "A_BCDEFGHIJKLM_N",
-      cterm = "JKLMNXXX",
-      cterm_cleavage_pos = 13
[…]

### 2024-09-28 tests/testthat/test-mspms_tidy.R

@@ -1,21 +1,21 @@
 test_that("mspms_tidy() returns a tibble with the expected colnames", {
-  expected_colData_names <- c("quantCols", "time", "condition", "group")
-  expected_rowData_names <- c(
-    "peptide",
-    "peptide_type",
-    "library_match_sequence",
-    "library_real_sequence",
-    "library_id",
-    "cleavage_seq",
-    "cleavage_pos"
-  )
-  expected_value_name <- "peptides_norm"
-  expected_names <- c(
-    expected_colData_names, expected_rowData_names,
-    expected_value_name
-  )
-  tidy <- mspms_tidy(mspms::processed_qf)
-  colnames <- colnames(tidy)
-  n_missing <- sum(expected_names %!in% colnames)
-  expect_equal(n_missing, 0)
+    expected_colData_names <- c("quantCols", "time", "condition", "group")
+    expected_rowData_names <- c(
+        "peptide",
+        "peptide_type",
+        "library_match_sequence",
+        "library_real_sequence",
+        "library_id",
+        "cleavage_seq",
+        "cleavage_pos"
+    )
+    expected_value_name <- "peptides_norm"
+    expected_names <- c(
+        expected_colData_names, expected_rowData_names,
+        expected_value_name
+    )
+    tidy <- mspms_tidy(mspms::processed_qf)
+    colnames <- colnames(tidy)
+    n_missing <- sum(expected_names %!in% colnames)
+    expect_equal(n_missing, 0)
 })

### 2024-09-28 tests/testthat/test-plot_pca.R

@@ -1,3 +1,3 @@
 test_that("plot_pca does not generate an error when testing with expected data", {
-  expect_no_error(plot_pca(mspms::mspms_tidy_data))
+    expect_no_error(plot_pca(mspms::mspms_tidy_data))
 })

### 2024-09-28 tests/testthat/test-plot_qc_check.R

@@ -1,4 +1,4 @@
 test_that("plot_qc_check() does not give an error when expected data as
           expected is supplied", {
-  expect_no_error(plot_qc_check(mspms::processed_qf))
+    expect_no_error(plot_qc_check(mspms::processed_qf))
 })

### 2024-09-28 tests/testthat/test-process_qf.R

@@ -1,9 +1,9 @@
 test_that("process_qf() produces the expected assays", {
-  processed_qf <- process_qf(mspms::peaks_prepared_data)
-  expected_assays <- c(
-    "peptides", "peptides_log", "peptides_log_norm",
-    "peptides_log_impute_norm", "peptides_norm"
-  )
-  sum <- sum(names(processed_qf) %!in% expected_assays)
-  expect_equal(sum, 0)
+    processed_qf <- process_qf(mspms::peaks_prepared_data)
+    expected_assays <- c(
+        "peptides", "peptides_log", "peptides_log_norm",
+        "peptides_log_impute_norm", "peptides_norm"
+    )
+    sum <- sum(names(processed_qf) %!in% expected_assays)
+    expect_equal(sum, 0)
 })

### 2024-09-28 tests/testthat/test-statistics_helper_functions.R

@@ -1,69 +1,69 @@
 ### remaining_cd_names() ###
 test_that("remaining_cd_names() returns the expected missing name, when time is
 provided as reference", {
-  time <- mspms:::remaining_cd_names(mspms::processed_qf, "time")
-  expect_equal(time, "condition")
+    time <- mspms:::remaining_cd_names(mspms::processed_qf, "time")
+    expect_equal(time, "condition")
 })
 
 test_that("remaining_cd_names() returns the expected missing name, when condition
 is provided as reference", {
-  condition <- mspms:::remaining_cd_names(mspms::processed_qf, "condition")
-  expect_equal(condition, "time")
+    condition <- mspms:::remaining_cd_names(mspms::processed_qf, "condition")
+    expect_equal(condition, "time")
 })
 
 test_that("remaining_cd_names() throws error when column not in data is
           attempted to be passed", {
-  expect_error(mspms:::remaining_cd_names(mspms::processed_qf, "not_here"))
+    expect_error(mspms:::remaining_cd_names(mspms::processed_qf, "not_here"))
 })
 
 ### mspms_log2fc() ###
 test_that("mspms_log2fc() returns does not error when data is as expected", {
-  expect_no_error(mspms:::mspms_log2fc(
-    mspms::processed_qf,
-    "time",
-    0
-  ))
+    expect_no_error(mspms:::mspms_log2fc(
+        mspms::processed_qf,
+        "time",
+        0
+    ))
 })
 
 test_that("mspms_log2fc() returns does not error when time is supplied as a
           character", {
-  expect_no_error(mspms:::mspms_log2fc(
-    mspms::processed_qf,
-    "time",
-    "0"
-  ))
+    expect_no_error(mspms:::mspms_log2fc(
+        mspms::processed_qf,
+        "time",
+        "0"
+    ))
 })
 
 test_that("mspms_log2fc() errors when a variable not in data is provided", {
-  expect_error(mspms:::mspms_log2fc(
-    mspms::processed_qf,
-    "not_here",
-    "0"
-  ))
+    expect_error(mspms:::mspms_log2fc(
+        mspms::processed_qf,
+        "not_here",
+        "0"
+    ))
 })
 
 ### mspms_t_tests() ###
 test_that("mspms_t_tests() does not error when data is as expected", {
-  expect_no_error(mspms:::mspms_t_tests(
-    mspms::processed_qf,
-    "time",
-    "0"
-  ))
[…]

### 2024-09-30 README.Rmd
4 spaces for indents
Bioccheck changes

@@
 ```{r}
-colData <- readr::read_csv(system.file("extdata/colData.csv", package = "mspms"))
+colData <- readr::read_csv(system.file("extdata/colData.csv",
+    package = "mspms"
+))
 
@@
 combined_peptide_filepath <- system.file("extdata/fragpipe_combined_peptide.tsv",
-  package = "mspms"
-)
+                                        package = "mspms")
 
@@
 Analysis of a PeptideGroups.txt file exported from [proteome
-discoverer](https://www.thermofisher.com/us/en/home/industrial/mass-spectrometry/liquid-chromatography-mass-spectrometry-lc-ms/lc-ms-software/multi-omics-data-analysis/proteome-discoverer-software.html) is supported.
+discoverer](https://www.thermofisher.com/us/en/home/industrial/mass-spectrometry/liquid-chromatography-mass-spectrometry-lc-ms/lc-ms-software/multi-omics-data-analysis/proteome-discoverer-software.html)
+is supported.
 

### 2024-09-30 R/plot_qc_check.R

@@ -37,7 +37,9 @@ plot_qc_check <- function(processed_qf,
             binwidth = 0.1,
             alpha = 0.5
         ) +
-        ggplot2::geom_density(ggplot2::aes(.data$per_library_id_undetected), alpha = 0.5) +
+        ggplot2::geom_density(ggplot2::aes(.data$per_library_id_undetected),
+            alpha = 0.5
+        ) +
         ggplot2::geom_vline(
             xintercept = full_length_threshold, linetype = "dashed",
             color = "#00BFC4"

### 2024-09-30 R/prepare_peaks.R

@@ -34,7 +34,7 @@ prepare_peaks <- function(lfq_filepath,
     message <- " peptides were removed because they had a quality score < "
     percentage <- round(n_peptides_rm / length(lfq$Peptide) * 100, 0)
     # Printing how many peptides are filtered out
-    print(paste0(n_peptides_rm, message, quality_threshold, " (", percentage, "%)"))
+    message(n_peptides_rm, message, quality_threshold, " (", percentage, "%)")
     # Selecting columns containing data
     start <- which(names(lfq) == "Avg. Area") + 1
     end <- which(names(lfq) == "Sample Profile (Ratio)") - 1

### 2024-10-01 README.Rmd
0.99.1

@@
 
-You can install the stable release from Bioconductor:
-
-Alternatively, you can install the development version from github.
+You can install the development version from github.
 
 ```{r,eval = FALSE}
-devtools::install_github("baynec2/mspms",
-    ref = "develop"
-)
+devtools::install_github("baynec2/mspms")
 ```

### 2024-10-01 README.Rmd
0.99.2

@@
 Lastly, mspms takes advantage of iceLogos ([Colaert, N. et al. Nature Methods 6,
-786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to viusalize
-overrepresented amino acid motifs relative to a background set by implementing
+786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to visualize
+over-represented amino acid motifs relative to a background set by implementing
 components of the Java software in R.
@@
 requires data that has been prepared for mspms data analysis by a converter
-function, and a design matrix. For more information see subsequent sections.
+function. For more information see subsequent sections.
 

### 2024-10-01 R/prepare_pd.R
0.99.2

@@ -1,4 +1,4 @@
-#' Title
+#' prepare_pd 
 #' Prepare a label free quantification file exported from Proteome Discoverer
 #' for subsequent mspms analysis.
 #' @param peptide_groups_filepath filepath to PeptideGroups.txt file exported

### 2024-10-26 R/calculate_all_cleavages.R
styler changes

@@ -10,52 +10,52 @@
 #' @export
 #' @examples
 #' calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
-#'     n_AA_after_cleavage = 4
+#'   n_AA_after_cleavage = 4
 #' )
 calculate_all_cleavages <- function(peptide_library_seqs,
                                     n_AA_after_cleavage = 4) {
-    # calculate the length of the peptides in library.
-    length <- unique(nchar(peptide_library_seqs))
-    # The current code assumes all the sequences are the same length
-    if (length(length) > 1) {
-        stop("Peptide library sequences are not of the same length")
+  # calculate the length of the peptides in library.
+  length <- unique(nchar(peptide_library_seqs))
+  # The current code assumes all the sequences are the same length
+  if (length(length) > 1) {
+    stop("Peptide library sequences are not of the same length")
+  }
+  # first cut at every single position
+  peptide_sequences <- c()
+  # for each position go through and cut into two pieces
+  for (i in seq_len(length - 1)) {
+    ## Doing this for the first fragment, left side of cleavage
+    first_fragment <- substr(peptide_library_seqs, 1, i)
+    n_char <- unique(nchar(first_fragment))
+    if (n_AA_after_cleavage - n_char > 0) {
+      n_x <- paste0(rep("X", n_AA_after_cleavage - n_char), collapse = "")
+    } else {
+      n_x <- ""
     }
-    # first cut at every single position
-    peptide_sequences <- c()
-    # for each position go through and cut into two pieces
-    for (i in seq_len(length - 1)) {
-        ## Doing this for the first fragment, left side of cleavage
-        first_fragment <- substr(peptide_library_seqs, 1, i)
-        n_char <- unique(nchar(first_fragment))
-        if (n_AA_after_cleavage - n_char > 0) {
-            n_x <- paste0(rep("X", n_AA_after_cleavage - n_char), collapse = "")
-        } else {
-            n_x <- ""
-        }
-        first_fragment_x <- paste0(n_x, first_fragment)
-        first_fragment_x <- substr(first_fragment_x,
-            stop = unique(nchar(first_fragment_x)),
-            start = unique(nchar(first_fragment_x)) -
-                n_AA_after_cleavage + 1
-        )
-        ## Doing this for the second fragment, right side of cleavage
-        second_fragment <- substr(
-            peptide_library_seqs, i + 1,
[…]

### 2024-10-26 R/log2fc_t_test.R

@@ -15,27 +15,27 @@
 log2fc_t_test <- function(processed_qf,
                           reference_variable = "time",
                           reference_value = 0) {
-    # Calculating the log2fc
-    log2fc <- mspms_log2fc(
-        processed_qf,
-        reference_variable,
-        reference_value
-    )
+  # Calculating the log2fc
+  log2fc <- mspms_log2fc(
+    processed_qf,
+    reference_variable,
+    reference_value
+  )
 
-    # Calculating the stats
-    stat <- mspms_t_tests(processed_qf,
-        reference_variable,
-        reference_value = as.character(reference_value)
-    )
+  # Calculating the stats
+  stat <- mspms_t_tests(processed_qf,
+    reference_variable,
+    reference_value = as.character(reference_value)
+  )
 
-    # Combining them together
-    log2fc_stats <- dplyr::inner_join(log2fc, stat, by = c(
-        "peptide",
-        "comparison"
-    )) %>%
-        tibble::as_tibble() %>%
-        dplyr::rename("condition" = "condition.x") %>%
-        dplyr::select(-"condition.y")
+  # Combining them together
+  log2fc_stats <- dplyr::inner_join(log2fc, stat, by = c(
+    "peptide",
+    "comparison"
+  )) %>%
+    tibble::as_tibble() %>%
+    dplyr::rename("condition" = "condition.x") %>%
+    dplyr::select(-"condition.y")
 
-    return(log2fc_stats)
+  return(log2fc_stats)
 }

### 2024-10-26 R/plot_cleavages_per_pos.R

@@ -11,21 +11,21 @@
 #' @examples
 #' # Defining the significant peptides
 #' sig_cleavage_data <- log2fc_t_test_data %>%
-#'     dplyr::filter(p.adj <= 0.05, log2fc > 3)
+#'   dplyr::filter(p.adj <= 0.05, log2fc > 3)
 #' # Plotting
 #' p1 <- mspms::plot_cleavages_per_pos(sig_cleavage_data)
 #' p1
 plot_cleavages_per_pos <- function(sig_cleavage_data,
                                    ncol = NULL) {
-    count_cleavages_per_pos <- count_cleavages_per_pos(sig_cleavage_data)
-    p1 <- count_cleavages_per_pos %>%
-        ggplot2::ggplot(ggplot2::aes(
-            x = .data$cleavage_pos,
-            y = .data$n,
-            color = .data$time
-        )) +
-        ggplot2::facet_wrap(~ .data$condition, scales = "free_y",ncol = ncol) +
-        ggplot2::geom_point() +
-        ggplot2::geom_line()
-    return(p1)
+  count_cleavages_per_pos <- count_cleavages_per_pos(sig_cleavage_data)
+  p1 <- count_cleavages_per_pos %>%
+    ggplot2::ggplot(ggplot2::aes(
+      x = .data$cleavage_pos,
+      y = .data$n,
+      color = .data$time
+    )) +
+    ggplot2::facet_wrap(~ .data$condition, scales = "free_y", ncol = ncol) +
+    ggplot2::geom_point() +
+    ggplot2::geom_line()
+  return(p1)
 }

### 2024-10-26 R/plot_volcano.R

@@ -22,38 +22,38 @@ plot_volcano <- function(log2fc_t_test_data,
                          padj_threshold = 0.05,
                          facets = "grid",
                          ncol = 1) {
-    ncol <- length(unique(log2fc_t_test_data$time))
-    p1 <- log2fc_t_test_data %>%
-        dplyr::mutate(`log_10_p.adj` = -log10(.data$p.adj)) %>%
-        ggplot2::ggplot(ggplot2::aes(
-            x = .data$log2fc,
-            y = .data$log_10_p.adj
-        )) +
-        ggplot2::geom_point(size = 0.5) +
-        ggplot2::geom_hline(
-            yintercept = -log10(padj_threshold),
-            linetype = "dashed", color = "red"
-        ) +
-        ggplot2::geom_vline(
-            xintercept = log2fc_threshold,
-            linetype = "dashed", color = "red"
-        ) +
-        ggplot2::geom_vline(
-            xintercept = -log2fc_threshold,
-            linetype = "dashed", color = "red"
-        ) +
-        ggplot2::labs(x = "Log2 (Time TX/T0)", y = "-log10(p.adj value)")
-    if (facets == "grid") {
-        p2 <- p1 + ggplot2::facet_grid(
-            rows = dplyr::vars(.data$condition),
-            cols = dplyr::vars(.data$time)
-        )
-    } else if (facets == "wrap") {
-        p2 <- p1 + ggplot2::facet_wrap(~ .data$condition + .data$time,
-            ncol = ncol
-        )
-    } else {
-        stop("facets must be either grid or wrap")
-    }
-    return(p2)
+  ncol <- length(unique(log2fc_t_test_data$time))
+  p1 <- log2fc_t_test_data %>%
+    dplyr::mutate(`log_10_p.adj` = -log10(.data$p.adj)) %>%
+    ggplot2::ggplot(ggplot2::aes(
+      x = .data$log2fc,
+      y = .data$log_10_p.adj
+    )) +
+    ggplot2::geom_point(size = 0.5) +
+    ggplot2::geom_hline(
+      yintercept = -log10(padj_threshold),
+      linetype = "dashed", color = "red"
+    ) +
+    ggplot2::geom_vline(
+      xintercept = log2fc_threshold,
+      linetype = "dashed", color = "red"
+    ) +
+    ggplot2::geom_vline(
+      xintercept = -log2fc_threshold,
[…]

### 2024-10-26 R/prepare_fragpipe.R

@@ -24,46 +24,46 @@ prepare_fragpipe <- function(combined_peptide_filepath,
                              colData_filepath,
                              peptide_library = mspms::peptide_library,
                              n_residues = 4) {
-    # Reading in the label free quantification data
-    lfq <- readr::read_tsv(combined_peptide_filepath, na = c("-", 0, "", "NA"))
-    # Check that this file appears to be a proper fragpipe file
-    check_file_is_valid_fragpipe(lfq)
-    # Reformat the columns to be consistent with mspsms framework
-    lfq_mod <- lfq %>%
-        dplyr::mutate(`Prev AA` = dplyr::case_when(
-            is.na(.data$`Prev AA`) ~ "",
-            TRUE ~ .data$`Prev AA`
-        )) %>%
-        dplyr::mutate(`Next AA` = dplyr::case_when(
-            is.na(.data$`Next AA`) ~ "",
-            TRUE ~ .data$`Next AA`
-        )) %>%
-        dplyr::mutate(
-            peptide = paste0(
-                .data$`Prev AA`, "_",
-                .data$`Peptide Sequence`, "_", .data$`Next AA`
-            ),
-            .before = 1
-        ) %>%
-        dplyr::mutate(
-            peptide = gsub("^_", "", .data$peptide),
-            peptide = gsub("_$", "", .data$peptide)
-        ) %>%
-        dplyr::select("peptide",
-            library_id = "Protein",
-            dplyr::contains("MaxLFQ Intensity")
-        ) %>%
-        dplyr::rename_with(
-            ~ stringr::str_replace_all(., " MaxLFQ Intensity", "")
-        )
-    # Converting 0 to NA
-    lfq_mod[lfq_mod == 0] <- NA
-    # reading in colData
-    colData <- load_colData(colData_filepath)
-    # converting into QF object
-    fragpipe_prepared_qf <- prepared_to_qf(
-        lfq_mod, colData,
-        peptide_library, n_residues
+  # Reading in the label free quantification data
+  lfq <- readr::read_tsv(combined_peptide_filepath, na = c("-", 0, "", "NA"))
+  # Check that this file appears to be a proper fragpipe file
+  check_file_is_valid_fragpipe(lfq)
[…]

### 2024-10-26 tests/testthat/test-cterm_cleavage.R

@@ -1,83 +1,82 @@
 # Testing for case where there is no cleavage site
 testthat::test_that("works with no cleavage site", {
-    testthat::expect_equal(
-        mspms:::cterm_cleavage(
-            "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",
-            "ABCDEFGHIJKLMN", 4
-        ),
-        tibble::tibble(
-            peptide = "ABCDEFGHIJKLMN",
-            cterm = NA,
-            cterm_cleavage_pos = NA
-        )
+  testthat::expect_equal(
+    mspms:::cterm_cleavage(
+      "ABCDEFGHIJKLMN", "ABCDEFGHIJKLMN",
+      "ABCDEFGHIJKLMN", 4
+    ),
+    tibble::tibble(
+      peptide = "ABCDEFGHIJKLMN",
+      cterm = NA,
+      cterm_cleavage_pos = NA
     )
+  )
 })
 
 
 # Testing for case where there is no cleavage site
 testthat::test_that("works with one clevage site", {
-    testthat::expect_equal(
-        mspms:::cterm_cleavage(
-            "ABCDEFGHIJKLM_N", "ABCDEFGHIJKLMN",
-            "ABCDEFGHIJKLMN", 4
-        ),
-        tibble::tibble(
-            peptide = "ABCDEFGHIJKLM_N",
-            cterm = "JKLMNXXX",
-            cterm_cleavage_pos = 13
-        )
+  testthat::expect_equal(
+    mspms:::cterm_cleavage(
+      "ABCDEFGHIJKLM_N", "ABCDEFGHIJKLMN",
+      "ABCDEFGHIJKLMN", 4
+    ),
+    tibble::tibble(
+      peptide = "ABCDEFGHIJKLM_N",
+      cterm = "JKLMNXXX",
+      cterm_cleavage_pos = 13
     )
+  )
 })
 
 
 # Testing for case where there are two cleavage sites
 testthat::test_that("works with two clevage site", {
-  
-  peptide_sequence = "A_BCDEFGHIJKLM_N"
-  library_match_sequence = "ABCDEFGHIJKLMN"
-  library_real_sequence = "ABCDEFGHIJKLMN"
-  n_residues = 4
-    testthat::expect_equal(
-        mspms:::cterm_cleavage(
[…]

### 2024-10-26 tests/testthat/test-log2fc_t_test.R

@@ -1,10 +1,10 @@
 test_that("log2fc_t_test() works using time as reference ", {
-    expect_no_error(log2fc_t_test(mspms::processed_qf,
-        "time",
-        reference_value = 0
-    ))
+  expect_no_error(log2fc_t_test(mspms::processed_qf,
+    "time",
+    reference_value = 0
+  ))
 })
-# 
+#
 # test_that("log2fc_t_test() thows an error when a value is not in data ", {
 #     expect_error(log2fc_t_test(mspms::processed_qf,
 #         "condition",

### 2024-10-26 tests/testthat/test-plot_all_icelogos.R

@@ -1,35 +1,35 @@
 test_that("plot_all_icelogos() produces ggarrange s3 class", {
-    sig <- mspms::log2fc_t_test_data %>%
-        dplyr::filter(p.adj <= 0.05, log2fc > 3)
-    t <- plot_all_icelogos(sig)
-    expect_s3_class(t, "ggarrange")
+  sig <- mspms::log2fc_t_test_data %>%
+    dplyr::filter(p.adj <= 0.05, log2fc > 3)
+  t <- plot_all_icelogos(sig)
+  expect_s3_class(t, "ggarrange")
 })
 
 test_that("plot_all_icelogos() works with nresidues = 4", {
   sig <- mspms::log2fc_t_test_data %>%
     dplyr::filter(p.adj <= 0.05, log2fc > 3)
-    expect_no_error(plot_all_icelogos(sig))
+  expect_no_error(plot_all_icelogos(sig))
 })
 
 
 test_that("plot_all_icelogos() works with nresidues = 8", {
-    
-  all_16mers = calculate_all_cleavages(
-    mspms::peptide_library$library_real_sequence,n_AA_after_cleavage = 8
-    )
-  
-  peaks_prepared_data = mspms::prepare_peaks(
+  all_16mers <- calculate_all_cleavages(
+    mspms::peptide_library$library_real_sequence,
+    n_AA_after_cleavage = 8
+  )
+
+  peaks_prepared_data <- mspms::prepare_peaks(
     system.file("extdata/peaks_protein-peptides-lfq.csv",
-                package = "mspms"
+      package = "mspms"
     ),
     system.file("extdata/colData.csv", package = "mspms"),
     n_residues = 8
-  ) %>% 
-    process_qf() %>% 
+  ) %>%
+    process_qf() %>%
     log2fc_t_test()
-  
-   sig <- peaks_prepared_data %>%
+
+  sig <- peaks_prepared_data %>%
     dplyr::filter(p.adj <= 0.05, log2fc > 3)
-   
-  expect_no_error(plot_all_icelogos(sig,background_universe = all_16mers))
+
+  expect_no_error(plot_all_icelogos(sig, background_universe = all_16mers))
 })

### 2024-10-26 tests/testthat/test-plot_heatmap.R

@@ -1,7 +1,7 @@
 test_that("plot_heatmap() produces no error with expected data", {
-    expect_no_error(plot_heatmap(mspms::mspms_tidy_data))
+  expect_no_error(plot_heatmap(mspms::mspms_tidy_data))
 })
 
 test_that("plot_heatmap() produces error with bad argument", {
-    expect_error(plot_heatmap(mspms::mspms_tidy_data, scale = "wrong_arg"))
+  expect_error(plot_heatmap(mspms::mspms_tidy_data, scale = "wrong_arg"))
 })

### 2024-10-26 tests/testthat/test-plot_time_course.R

@@ -1,4 +1,4 @@
 test_that("plot_time_course() does not generate an error when data is as
           expected", {
-    expect_no_error(plot_time_course(mspms::mspms_tidy_data))
+  expect_no_error(plot_time_course(mspms::mspms_tidy_data))
 })

### 2024-10-26 tests/testthat/test-preprocessing_helper_functions.R

@@ -1,534 +1,534 @@
 ### check_file_is_valid_peaks()###
 test_that("check_file_is_valid_peaks() errors when PTM is provided.", {
-    peaks_ptm <- readr::read_csv(system.file(
-        "extdata/peaks_protein-peptides-lfq.csv",
-        package = "mspms"
-    )) %>%
-        dplyr::mutate(PTM = "PTM")
-    expect_error(mspms:::check_file_is_valid_peaks(peaks_ptm))
+  peaks_ptm <- readr::read_csv(system.file(
+    "extdata/peaks_protein-peptides-lfq.csv",
+    package = "mspms"
+  )) %>%
+    dplyr::mutate(PTM = "PTM")
+  expect_error(mspms:::check_file_is_valid_peaks(peaks_ptm))
 })
 
 ### check_file_is_valid_fragpipe()###
 test_that("check_file_is_valid_fragpipe() errors when peaks file is
           provided.", {
-    peaks_ptm <- readr::read_csv("tests/testdata/
+  peaks_ptm <- readr::read_csv("tests/testdata/
                                check_file_is_valid_peaks_ptm.csv")
-    expect_error(mspms:::check_file_is_valid_fragpipe(peaks_ptm))
+  expect_error(mspms:::check_file_is_valid_fragpipe(peaks_ptm))
 })
 
 
 ### Testing nterm cleavages ###
 
 test_that("nterm_cleavages() works as expected with n_residue = 1", {
-  peptide_sequence = "G_HIDHL"
-  library_match_sequence = "LVATVYEFGHIDHL"
-  library_real_sequence =   "LVATVYEFGHIDHL"
-  
-  out = mspms:::nterm_cleavage(peptide_sequence,
-                               library_match_sequence,
-                               library_real_sequence,
-                               n_residues = 1)
-  
-  expect_equal(out$nterm,"GH")
-  expect_equal(out$nterm_cleavage_pos,9)
-  
+  peptide_sequence <- "G_HIDHL"
+  library_match_sequence <- "LVATVYEFGHIDHL"
+  library_real_sequence <- "LVATVYEFGHIDHL"
+
+  out <- mspms:::nterm_cleavage(peptide_sequence,
+    library_match_sequence,
+    library_real_sequence,
+    n_residues = 1
+  )
+
+  expect_equal(out$nterm, "GH")
+  expect_equal(out$nterm_cleavage_pos, 9)
 })
 
[…]

### 2024-10-26 tests/testthat/test-process_qf.R

@@ -1,9 +1,9 @@
 test_that("process_qf() produces the expected assays", {
-    processed_qf <- process_qf(mspms::peaks_prepared_data)
-    expected_assays <- c(
-        "peptides", "peptides_log", "peptides_log_norm",
-        "peptides_log_impute_norm", "peptides_norm"
-    )
-    sum <- sum(names(processed_qf) %!in% expected_assays)
-    expect_equal(sum, 0)
+  processed_qf <- process_qf(mspms::peaks_prepared_data)
+  expected_assays <- c(
+    "peptides", "peptides_log", "peptides_log_norm",
+    "peptides_log_impute_norm", "peptides_norm"
+  )
+  sum <- sum(names(processed_qf) %!in% expected_assays)
+  expect_equal(sum, 0)
 })

### 2024-10-26 tests/testthat/test-processing_helper_functions.R

@@ -1,4 +1,4 @@
 ### rlog2() ###
 test_that("rlog2() works", {
-    expect_equal(mspms:::rlog2(2), 4)
+  expect_equal(mspms:::rlog2(2), 4)
 })

### 2024-10-28 README.Rmd
styler changes
removed cathepsin B-D from example datasets

@@
 head(colData)
+
 ```
@@
 peaks_prepared_data
+
 ```
@@
 combined_peptide_filepath <- system.file("extdata/fragpipe_combined_peptide.tsv",
-                                        package = "mspms")
+  package = "mspms"
+)
 
@@
 log2fc_t_test_data <- mspms::log2fc_t_test(processed_qf = processed_qf)
+
 ```

### 2024-11-21 README.Rmd
readme and vignette updates
made suggested final changes from Bioconductor review

--- 2024-10-01 README.Rmd
@@
 
-# mspms
+# `r BiocStyle::Biocpkg("mspms")`
 
@@
 coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
-
+[![Bioc release status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)
+[![Bioc devel status](http://www.bioconductor.org/shields/build/devel/bioc/mspms.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/mspms)
+[![Bioc downloads rank](https://bioconductor.org/shields/downloads/release/mspms.svg)](http://bioconductor.org/packages/stats/bioc/mspms/)
+[![Bioc support](https://bioconductor.org/shields/posts/mspms.svg)](https://support.bioconductor.org/tag/mspms)
+[![Bioc history](https://bioconductor.org/shields/years-in-bioc/mspms.svg)](https://bioconductor.org/packages/release/bioc/html/mspms.html#since)
+[![Bioc last commit](https://bioconductor.org/shields/lastcommit/devel/bioc/mspms.svg)](http://bioconductor.org/checkResults/devel/bioc-LATEST/mspms/)
+[![Bioc dependencies](https://bioconductor.org/shields/dependencies/release/mspms.svg)](https://bioconductor.org/packages/release/bioc/html/mspms.html#since)
 <!-- badges: end -->
@@
 
-The goal of mspms is provide a R Package that can be used to easily and
+The goal of `r BiocStyle::Biocpkg("mspms")` is provide a R Package that can be used to easily and
 reproducibly analyze data resulting from the [Multiplex Substrate Profiling by
@@
 Additionally, we provide a [graphical user interface powered by shiny
-apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to
+apps](https://github.com/baynec2/mspms-shiny) that allows for a user to
 utilize the method without requiring any R coding knowledge.
@@
 This involves unique preprocessing steps and data visualizations tailored to
-this specific assay
+this specific assay.
 
@@
 
--   QFeatures: management and processing of peptide quantitative features and
+-   `r BiocStyle::Biocpkg("QFeatures")`: management and processing of peptide quantitative features and
     sample metadata.
 
--   MsCoreUtils: log2 transformation, imputation (QRILC), and normalization
+-   `r BiocStyle::Biocpkg("MsCoreUtils")`: log2 transformation, imputation (QRILC), and normalization
     (center.median).
@@
 
--   ggplot2: data visualizations
+-   *ggplot2*: data visualizations
 
--   heatmaply: interactive heatmaps
+-   *heatmaply*: interactive heatmaps
 
--   rstatix: conduct statistics
+-   *rstatix*: conduct statistics
 
--   ggseqlogo: iceLogo motif visualization.
+-   *ggseqlogo*: iceLogo motif visualization.
 
--   downloadthis: gives users the ability to download data from standard .html
+-   *downloadthis*: gives users the ability to download data from standard .html
     report generated by mspms::generate_report().
 
-Lastly, mspms takes advantage of iceLogos ([Colaert, N. et al. Nature Methods 6,
+Lastly, *mspms* takes advantage of iceLogos ([Colaert, N. et al. Nature Methods 6,
 786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to viusalize
@@
 
-2.  Data processing. These functions allow the user to process the proteomics
-    data (imputation, normalization, etc).
+2.  Data processing. These functions allow the user to process the proteomics data (imputation, normalization, etc).
 
-3.  Statistics. These methods allow the user to perform basic statistics on the
-    normalized/processed data.
+3.  Statistics. These methods allow the user to perform basic statistics on the normalized/processed data.
 
@@
 
-**Note that only a subset of functions are exported to the user.** Helper
-functions can be found in the \_helper_functions.R files.
+**Note that only a subset of functions are exported to the user.** Helper functions can be found in the
+[preprocessing_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/preprocessing_helper_functions.R),
+[processing_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/processing_helper_functions.R),[icelogo_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/iceLogo_helper_fuctions.R), [statistics_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/statistics_helper_functions.R), and [iceLogo_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/iceLogo_helper_fuctions.R) files.
+
+These functions were not intended to be directly used by the user and are not exported in an effort to make the package API cleaner and more intuitive.
 
@@
 
-#### A note on colData
+### A note on colData
 
@@
 
-It is very important that this contains at least the following columns
--"quantCols","group","condition",and "time".
+**It is very important that this contains at least the following columns: quantCols","group","condition",and "time".**
 
@@
 
-#### PEAKS
+### PEAKS
 
@@
 ```{r}
+
 library(dplyr)
@@
   colData_filepath,
-    0.3,
+  quality_threshold = 0.3,
   n_residues = 4
@@
 peaks_prepared_data
+
 ```
 
-#### Fragpipe
+### Fragpipe
 
@@
 ```
-#### Proteome Discoverer
+### Proteome Discoverer
 
@@
 
-This makes the data easy to manipulate prior to ggplot2 or plotly based plotting 
-functions.
+This makes the data easy to manipulate prior to ggplot2 or plotly based plotting functions.
 
@@
 
-```{r}
+```{r,eval = FALSE}
 plot_heatmap(mspms_tidy_data)
 ```
+
+note that the interactive plot is not returned in this vignette 
+due to size constraints.
 
@@
 
-```{r,fig.height = 10, fig.width = 4}
+```{r}
 sig_cleavage_data <- log2fc_t_test_data %>%

### 2025-01-28 README.Rmd
Added limma based statistics.

@@
 
--   `r BiocStyle::Biocpkg("QFeatures")`: management and processing of peptide quantitative features and
-    sample metadata.
+-   `r BiocStyle::Biocpkg("QFeatures")`: management and processing of peptide 
+quantitative features and sample metadata.
 
--   `r BiocStyle::Biocpkg("MsCoreUtils")`: log2 transformation, imputation (QRILC), and normalization
-    (center.median).
+-   `r BiocStyle::Biocpkg("limma")`: Powerful linear models and 
+empirical Bayesian methods for omics data.
+
+-   `r BiocStyle::Biocpkg("MsCoreUtils")`: log2 transformation, 
+imputation (QRILC), and normalization (center.median).
 
@@
 
-Lastly, *mspms* takes advantage of iceLogos ([Colaert, N. et al. Nature Methods 6,
-786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to viusalize
-overrepresented amino acid motifs relative to a background set by implementing
+Lastly, *mspms* takes advantage of iceLogos ([Colaert, N. et al. Nature Methods
+6,786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to visualize
+over represented amino acid motifs relative to a background set by implementing
 components of the Java software in R.
@@
 
-2.  Data processing. These functions allow the user to process the proteomics data (imputation, normalization, etc).
+2.  Data processing. These functions allow the user to process the proteomics 
+data (imputation, normalization, etc).
 
-3.  Statistics. These methods allow the user to perform basic statistics on the normalized/processed data.
+3.  Statistics. These methods allow the user to perform basic statistics on the 
+normalized/processed data.
 
@@
 
-These functions were not intended to be directly used by the user and are not exported in an effort to make the package API cleaner and more intuitive.
+These functions were not intended to be directly used by the user and are not 
+exported in an effort to make the package API cleaner and more intuitive.
 
@@
 
-We provide a user-friendly implementation of this statistical test as follows:
+We recommend that the user now uses a Limma based approach that uses a linear 
+modeling and empirical Bayesian approach. This is a approach well suited to 
+proteomics data that is powerful, flexible, and fast.
+
+This limma based approach (with stats calculated relative to time 0 for each 
+condition) can be utilized as follows:
+
+```{r}
+limma_stats <- mspms::limma_stats(processed_qf = processed_qf)
+```
+
+Alternatively, we provide a user-friendly implementation of the traditional
+t-test approach as follows:
 
@@
 ```
-
-Please note that statistical analyses are one of the strengths of the
-Bioconductor project and several packages provide alternative statistical
-approaches that are more flexible than t-tests and could be used with the
-QFeatures objects produced by mspms.
-
-This includes packages using empirical Bayesian approaches (see
-[limma](https://bioconductor.org/packages/release/bioc/html/limma.html)) and
-mixed modelling (see
-[msqrob2](https://bioconductor.org/packages/release/bioc/html/msqrob2.html)),
-that are suited for proteomics data.
-
-Users are encouraged to use these excellent packages.
 
@@
 
-This makes the data easy to manipulate prior to ggplot2 or plotly based plotting functions.
+This makes the data easy to manipulate prior to ggplot2 or plotly based plotting
+functions.
 

### 2025-01-28 R/preprocessing_helper_functions.R

@@ -372,7 +372,7 @@ prepared_to_qf <- function(prepared_data,
 
 #' load_colData
 #'
-#' load a .csv file containing sample colData.
+#' load a .csv file containing sample colData. Check for errors
 #'
 #' @param colData_filepath filepath to .csv file containing colData.
 #'
@@ -381,20 +381,43 @@ prepared_to_qf <- function(prepared_data,
 
 load_colData <- function(colData_filepath) {
   colData <- readr::read_csv(colData_filepath)
-  colData_names <- names(colData)
+  # Expected names
+  `%!in%` <- Negate(`%in%`)
   expected_names <- c("quantCols", "group", "condition", "time")
+  colData_names <- names(colData)
   unexpected_names <- expected_names %!in% colData_names
   n_unexpected_names <- sum(unexpected_names)
   if (n_unexpected_names > 0) {
     missing_names <- expected_names[unexpected_names]
     stop(paste0(c("colData must have columns named \"quantCols\",\"group\",
-    \"condition\", and \"time\" ", "you are missing ", missing_names)),
-      collapse = ""
+    \"condition\", and \"time\" ", "you are missing ", paste(missing_names,
+                                                             collapse = ", "))),
+         collapse = ""
     )
   }
+  # Check if the column types match
+  expected_types <- c("character", "character", "character", "numeric")
+  colData_types <- sapply(colData, class)
+  unexpected_types <- (colData_types != expected_types[1:4])
+  n_unexpected_types <- sum(unexpected_types)
+  
+  if (n_unexpected_types > 0) {
+    missing_types <- names(colData)[unexpected_types]
+    stop(paste0("colData columns must be as follows: quantCols = character,
+    group = character, condition = character, time = numeric.",
+                "You have the wrong type for the following column(s): ",
+                paste(missing_types, collapse = ", ")))
+  }
+  # Check to make sure none of the columns have spaces in them
+  n_spaces <- sum(sapply(colData, function(x){grepl(".* .*",x)}))
+  if(n_spaces > 0){
+    stop("Your colData has spaces in it (ie condition = Cathepsin A). Please
+    change it to not contain any spaces to be compatible with limma (ie
+    condition  = cathepsin_A)")
+    }
   return(colData)
 }
-
+  
 #' check_peptide_library
 #'
 #' @param peptide_library

### 2025-05-12 README.Rmd
added prepare_diann() to support DIA-NN output.

@@
 ```
+### DIA-NN
+
+We are currently investigating how well the method works with DIA data. 
+To allow for this we have created a parser to make DIA-NN output compatible with
+mspms.
+
+Note that this could be used with DIA-NN alone or with Fragpipe (MSFragger-DIA) 
+uses DIA-NN for quantification. Here we are using the file output from Fragpipe.
+
+Use caution when using this approach, we haven't thoroughly investigated how 
+well DIA data works with mspms using ground truth data yet.
+
+```{r}
+precursor_filepath = system.file(
+  "extdata/diann_report.pr_matrix.tsv",
+  package = "mspms"
+)
+colData_filepath = system.file(
+  "extdata/diann_colData.csv",
+  package = "mspms"
+)
+
+prepared_diann_data <- prepare_diann(precursor_filepath,colData_filepath)
+
+prepared_diann_data
+```
 

### 2025-05-12 R/preprocessing_helper_functions.R

@@ -88,6 +88,31 @@ check_file_is_valid_pd <- function(pd_data) {
     )
   }
 }
+#' check_file_is_valid_diann
+#' Check to make sure the input data looks like the expected DIA-NN output
+#' file.
+#' @param pd_data PeptideGroups.txt file generated by DIA-NN and read
+#' into R.
+#'
+#' @return a stop command with a informative message if file looks unexpected.
+#' otherwise, nothing.
+check_file_is_valid_diann <- function(diann_data){
+  expected_names <- c("Protein.Group","Protein.Ids","Protein.Names","Genes",
+                      "First.Protein.Description","Proteotypic",
+                      "Stripped.Sequence", "Modified.Sequence",
+                      "Precursor.Charge","Precursor.Id","All Mapped Proteins",
+                      "All Mapped Genes" )
+  
+  names_in_data <- names(diann_data)
+  # What names are missing
+  missing_names <- expected_names[expected_names %!in% names_in_data]
+    if (length(missing_names) > 1) {
+    stop(
+      paste(c("This doesn't look like the expected PEAKS file. The following
+          columns are missing:", missing_names), collapse = "")
+    )
+  }
+}
 
 #' cterm_cleavage
 #'

### 2025-11-12 README.Rmd

@@
 ```{r}
-
 library(dplyr)
@@
 peaks_prepared_data
-
 ```
@@
 ```
+### Sage
+
+Analysis of a lfq.tsv file exported from [sage](https://github.com/lazear/sage) is supported.
+We found this software a little harder to use than other solutions (No-GUI), however it is completly open-source, free, and by far the fastest!
+
+```{r}
+sage_lfq_filepath <- system.file(
+  "extdata/sage_lfq.tsv",
+  package = "mspms"
+)
+
+colData_filepath <- system.file("extdata/colData.csv",
+  package = "mspms"
+)
+
+prepare_sage(sage_lfq_filepath, colData_filepath)
+```
+
 ### DIA-NN
@@
 ```{r}
-precursor_filepath = system.file(
+precursor_filepath <- system.file(
   "extdata/diann_report.pr_matrix.tsv",
@@
 )
-colData_filepath = system.file(
+colData_filepath <- system.file(
   "extdata/diann_colData.csv",
@@
 
-prepared_diann_data <- prepare_diann(precursor_filepath,colData_filepath)
+prepared_diann_data <- prepare_diann(precursor_filepath, colData_filepath)
 
